### 2021-02-26 latex/MPP21/paper.tex
added style files for MPP conference

\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{cleveref}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }

\newcommand{\GB}[1]{\textcolor{red}{\textbf{GB}: #1}}

\begin{document}

\title{Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms
%\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{Grey Ballard}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Wake Forest University}\\
Winston Salem, NC, USA \\
\email{ballard@wfu.edu}}
\and
\IEEEauthorblockN{Jack Weissenberger}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.) \GB{?}} \\
\textit{Capital One \GB{?}}\\
Richmond, VA \\
email address or ORCID}
\and
\IEEEauthorblockN{Luoping\GB{?} Zhang}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Wake Forest University}\\
Winston Salem, NC, USA \\
\email{zhanl317@wfu.edu}}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}

\GB{reference this paper: \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=1675476}}

This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2021-02-27 latex/overleaf-version/main.tex
added overleaf to repo

\documentclass{article}
\usepackage[utf8]{inputenc}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{cite}

\newtheorem{theorem}[]{Theorem}

\title{APA-FMM Algorithms}
\author{}
\date{}

\begin{document}

\maketitle

\section{Introduction}

\section{Divide and Conquer}
For a matrix multiplication, we have $A\cdot B = C$, where $A$ is a $n$ by $m$ matrix, and $B$ is a $m$ by $k$ matrix, then C is to be $n$ by $k$. Thus, there should be $nm + mk + nk$ entries.\\
Before we use any fast matrix multiplication algorithm, we know that the if we implement the classic algorithm into Divide and Conquer form, we will have $nmk$ recursive calls in each iteration by Master Theorem. We want to find recursive algorithms that have less than $nmk$ recursive calls each iteration.\\

\section{Strassen's Algorithm}
Strassen's Algorithm deals with 2 by 2 matrix multiplies 2 by 2 matrix, denoted as $(2,2,2)$. By Master Theorem, we want the recursive calls to be less than $2*2*2 = 8$. Strassen provides us an algorithm of 7 recurisve calls.\\
We can think of the recursive calls as a mapping from a table of parameter to $r$ recursive calls with different entries from the matrices and their parameters. In each recursive call, we only select a few entries from the original matrices and add a parameter on it (in this case, the parameter is only 1 or -1). If an entry does not appear on the call, we can think its parameter is 0. So we can construct a table with parameter for each entry of the matrices and each recursive call. (UVW Notation)\\
For Strassen's Algorithm, we know that there are $nm+mk+nk = 4+4+4 = 12$ entries and 7 recursive calls, so the table should be 12 by 7.
\begin{eqnarray*}
M_1 &=& (A_{11} + A_{22}) \cdot (B_{11} + B_{22}) \\
M_2 &=& (A_{21} + A_{22})\cdot  B_{11} \\
M_3 &=& A_{11}\cdot(B_{12} - B_{22}) \\
M_4 &=&        A_{22}\cdot (B_{21} -B_{11}) \\
M_5 &=& (A_{11}+ A_{12}) \cdot B_{22} \\
M_6 &=& (A_{21} - A_{11}) \cdot (B_{11} + B_{12}) \\
M_7 &=& (A_{12}- A_{22}) \cdot (B_{21} + B_{22}) \\\\
C_{11} &=& M_1 + M_4 - M_5 + M_7 \\
C_{12} &=& M_3 + M_5 \\
C_{21} &=& M_2 + M_4 \\
C_{22} &=& M_1 - M_2 + M_3 + M_6\\
\end{eqnarray*}
By Master Theory, the time complexity of Strassen's Algorithm is $O(n^{\log\limits_27}) = O(n^{2.807355})$.\\

\section{APA Algorithm}
In Strassen's Algorithm, all the parameter in the table are integers. To get a faster algorithm, we want to reduce the precision of the calculation small enough so that the speedup is much more significant.\\
We will add a variable $\varepsilon$ (denoted as $e$ below in the recursive equation), we want to choose an appropriate value for $\varepsilon$ so that the error of the APA algorithm will the smallest.\\
But here's a trade-off of the value of $\varepsilon$. Theoretically, when $\varepsilon \to 0$, error of the algorithm $\to 0$. Thus, we want $\varepsilon$ to be small.\\
However, practically, when computers are making floating point calculation, they have precision of certain amount of bits. If $\varepsilon$ is chosen too small, computers cannot calculate the correct answer, so we want $\varepsilon$ to be big enough.

\section{UVW Notation}
In the section of Strassen's Algorithm, we provide a table that could store parameters for any matrix mutiplication algorithm,  and now we want to formalize the table so that it is easy to understand. The UVW Notation divides the rows into 3 parts: the row for matrix $A$, the row for matrix $B$, and the row for matrix $C$, where $A\cdot B = C$. So the original table $nm+mk+nk$ by $r$ becomes three tables with same number of columns: $U \rightarrow (nm,r), V\rightarrow(mk,r), W\rightarrow(nk,r)$.\\
To retrieve recursive calls, we can just do a UVW table lookup. For the Divide part, each column represents a recursive call, and the 2 variables of the recursive call $i$ are the linear combination of the entries in matrix $A$ for corresponding column $r_i$ in U table, and the linear combination of the entries in matrix $B$ for corresponding column $r_i$ in V table. For the Conquer part, we take each row of table W, the linear combination between each recursive call and their parameters represents the corresponding entries.

\section{Bini's Algorithm \cite{BCRL79}}

\begin{eqnarray*}
M_1 &=& (A_{11} + A_{22}) \cdot (eB_{11} + B_{22}) \\
M_2 &=& A_{22}\cdot  (-B_{21} - B_{22}) \\
M_3 &=& A_{11}\cdot B_{22} \\
M_4 &=& (eA_{12} + A_{22})\cdot (-eB_{11} + B_{21}) \\
M_5 &=& (A_{11} + eA_{12}) \cdot (eB_{12} + B_{22}) \\
M_6 &=& (A_{21} + A_{32}) \cdot (B_{11} + eB_{22}) \\
M_7 &=& A_{21} \cdot (-B_{11} - B_{12}) \\
M_8 &=& A_{32} \cdot B_{11}\\
M_9 &=& (A_{21} + eA_{31}) \cdot (B_{12} - eB_{22})\\
M_{10} &=& (eA_{31} + A_{32}) \cdot (B_{12} - eB_{22})\\\\
C_{11} &=& (M_1 + M_2 - M_3 + M_4)/e \\
C_{12} &=& (-M_3 + M_5)/e \\
C_{21} &=& M_4 + M_6 - M_{10}\\
C_{22} &=& M_1 - M_5 + M_9\\
C_{31} &=& (-M_8 + M_{10})/e \\
C_{32} &=& (M_6 + M_7 - M_8 + M_9)/e\\
\end{eqnarray*}
By Master Theorem, the time complexity of Bini's Algorithm is $O(n^{\log\limits_{3*2*2}10^3}) = O(n^{\log\limits_{12}1000}) = O(n^{2.779885})$.
\section{Schonhage's Algorithm \cite{Schonhage81}}

\begin{eqnarray*}
M_{11} &=& (A_{11} + e^2A_{12})\cdot (e^2B_{11} + B_{21})\\
M_{22} &=& (A_{21} + e^2A_{22})\cdot (e^2B_{12} + B_{22})\\ 
M_{33} &=& (A_{31} + e^2A_{32})\cdot (e^2B_{13} + B_{23})\\
N_{11} &=& (A_{11} + e^2A_{13})\cdot B_{31}\\
N_{22} &=& (A_{21} + e^2A_{23})\cdot B_{32}\\
N_{33} &=& (A_{31} + e^2A_{33})\cdot B_{33}\\
O_1 &=& A_{11}\cdot (B_{21} + B_{31})\\
O_2 &=& A_{21}\cdot (B_{22} + B_{32})\\
O_3 &=& A_{31}\cdot (B_{23} + B_{33})\\
M_{12} &=& (A_{11} + e^2A_{22})\cdot (B_{21} - eB_{12})\\
M_{13} &=& (A_{11} + e^2A_{32})\cdot (B_{21} - eB_{13})\\
M_{23} &=& (A_{21} + e^2A_{32})\cdot (B_{22} - eB_{13})\\
M_{21} &=& (A_{21} + e^2A_{12})\cdot (B_{22} - eB_{11})\\
M_{31} &=& (A_{31} + e^2A_{12})\cdot (B_{23} - eB_{11})\\
M_{32} &=& (A_{31} + e^2A_{22})\cdot (B_{23} - eB_{12})\\
N_{12} &=& (A_{11} + e^2A_{23})\cdot (B_{31} + eB_{12})\\
N_{13} &=& (A_{11} + e^2A_{33})\cdot (B_{31} + eB_{13})\\
N_{23} &=& (A_{21} + e^2A_{33})\cdot (B_{32} + eB_{13})\\
N_{21} &=& (A_{21} + e^2A_{13})\cdot (B_{32} + eB_{11})\\
N_{31} &=& (A_{31} + e^2A_{13})\cdot (B_{33} + eB_{11})\\
N_{32} &=& (A_{31} + e^2A_{23})\cdot (B_{33} + eB_{12})\\\\
C'_{11} &=& (M_{11} + N_{11} - O_1)/e^2\\
C'_{22} &=& (M_{22} + N_{22} - O_2)/e^2\\
C'_{33} &=& (M_{33} + N_{33} - O_3)/e^2\\
C'_{12} &=& (M_{21} + N_{21} - O_2)/e^2 + (N_{12} - N_{11})/e\\
C'_{13} &=& (M_{31} + N_{31} - O_3)/e^2 + (N_{13} - N_{11})/e\\
C'_{23} &=& (M_{32} + N_{32} - O_3)/e^2 + (N_{23} - N_{22})/e\\
C'_{21} &=& (M_{12} + N_{12} - O_1)/e^2 + (N_{21} - N_{22})/e\\
C'_{31} &=& (M_{13} + N_{13} - O_1)/e^2 + (N_{31} - N_{33})/e\\
C'_{32} &=& (M_{23} + N_{23} - O_2)/e^2 + (N_{32} - N_{33})/e\\
\end{eqnarray*}
By Master Theorem, the time complexity of Schonhage's Algorithm is $O(n^{\log\limits_321}) = O(n^{2.77124})$.

\section{Smirnov's 323 Algorithm \cite{Smirnov13}}

\begin{eqnarray*}
M_1 &=& (e^{-1}A_{12} + eA_{21}) \cdot e^{-1}B_{12}\\
M_2 &=& (e^{-1}A_{12} + e^2A_{31} - A_{32}) \cdot e^{-1}B_{11}\\
M_3 &=& (-e^{-1}A_{11} + (1+e^2)A_{31}) \cdot (e^{-1}B_{11} + eB_{13} - eB_{23})\\
M_4 &=& A_{21} \cdot (B_{11} + B_{12} + eB_{13})\\
M_5 &=& (-e^{-1}A_{11} - e^{-1}A_{12} + A_{31} + A_{32}) \cdot (-e^{-1}B_{11} + eB_{23})\\
M_6 &=& (-e^{-1}A_{12} - eA_{21} - e^2A_{31} + A_{32}) \cdot (-e^{-1}B_{11} + B_{22})\\
M_7 &=& (-e^{-1}A_{12} + e^3A_{21} + A_{32}) \cdot (e^{-1}B_{11} + eB_{21})\\
M_8 &=& -e^{-1}A_{12} \cdot (-e^{-1}B_{12} + e^{-1}B_{21})\\
M_9 &=& (e^{-1}A_{12} + eA_{21} - A_{22} + e^2A_{31} - A_{32}) \cdot B_{22}\\
M_{10} &=& e^{-1}A_{11} \cdot (e^{-1}B_{12} + eB_{13} - eB_{23})\\
M_{11} &=& (e^{-1}A_{11} + e^{-1}A_{12} + eA_{21}) \cdot (-e^{-1}B_{12} + eB_{23})\\
M_{12} &=& (e^{-1}A_{12} + eA_{22}) \cdot e^{-1}B_{21}\\
M_{13} &=& (e^{-1}A_{12} + (e+e^2)A_{21} + e^2A_{31}) \cdot (e^{-1}B_{12} + eB_{22})\\
M_{14} &=& A_{22} \cdot (-B_{21} - B_{22} + eB_{23})\\\\
C_{11} &=& eM_2 - e^2M_3 + eM_7 + (e^2+e^3)M_{12}\\
C_{12} &=& -M_1 + e^2M_{10} + M_{13}\\
C_{13} &=& M_1 - e^2M_2 + e^2M_3 + e^2M_5 + M_{10} + M_{11}\\
C_{21} &=& -M_1 + M_4 + M_8 + M_{12}\\
C_{22} &=& M_1 - M_8 - M_{12} - M_{14}\\
C_{23} &=& -e^{-1}M_1 + e^{-1}M_2 + e^{-1}M_4 - e^{-1}M_6 + e^{-1}M_8 - e^{-1}M_9 + e^{-1}M_{12} + e^{-1}M_{14}\\
C_{31} &=& eM_1 + e^{-1}M_2 - eM_4 + e^{-1}M_7 - eM_8\\
C_{32} &=& -e^{-1}M_1 - M_2 - M_4 + M_6 + e^{-1}M_{13}\\
C_{33} &=& e^{-1}M_1 - e^{-1}M_2 + e^{-1}M_3 + e^{-1}M_5 + e^{-1}M_{10} + e^{-1}M_{11}\\
\end{eqnarray*}
By Master Theorem, the time complexity of Smirnov's 323 Algorithm is $O(n^{\log\limits_{3*2*3}14^3}) = O(n^{\log\limits_{18}2744}) = O(n^{2.73915})$.

\section{Smirnov's 333 Algorithm \cite{Smirnov13}}
\begin{eqnarray*}
M_1 &=& (eA_{11}-e^3A_{13}+e^{-1}A_{31}+e^{-2}A_{33})\cdot(B_{11}+eB_{12}+e^{-1}B_{}21)+e^{-2}B_{31}+2e^3B_{33}\\
M_2 &=& (e^3A_{13}+e^{-2}A_{21}-e^{-1}A_{22}-eA_{33})\cdot(e^3B_{23}+e^{-1}B_{33})\\
M_3 &=& (-e^{-1}A_{21}+A_{22}-e^{-2}A_{32})\cdot(e^{-1}B_{21}+B_{22}-e^2B_{23})\\
M_4 &=& (-A_{11}+eA_{12}-e^{-2}A_{31}+e^{-1}A_{32})\cdot(eB_{11}+B_{21}+e^{-1}B_{31})\\
M_5 &=& (-e^3A_{13}+e^{-2}A_{21}+e^{-2}A_{33})\cdot(B_{11}+eB_{12}+e^{-1}B_{21}e^{-2}B_{31}+e^{-1}B_{32}+e^3B_{33})\\
M_6 &=& (e^{-2}A_{31}-e^{-1}A_{32})\cdot(e^2B_{13}+e^{-2}B_{21})\\
M_7 &=& (eA_{11}-2e^2A_{12}+e^{-2}A_{31}-e^{-1}A_{32})\cdot(eB_{11}+e^3B_{13}+e^{-1}B_{21})\\
M_8 &=& (eA_{11}-e^3A_{13}-e^{-2}A_{21}+A_{23}+e^{-1}A_{31})\cdot()eB_{13}+e^{-2}B_{31}\\
M_9 &=& (-e_{-2}A_{21}+e^{-1}A_{22})\cdot(eB_{22}+e^{-1}B_{33})\\
M_{10} &=& e^{-2}\cdot(e^3B_{13}+e^{-1}B_{21}+B_{22})\\
M_{11} &=& (e^{-2}A_{21}-A_{23})\cdot(eB_{13}+B_{23}+B_{32})\\
M_{12} &=& (e^2A_{12}-e^{-2}A_{21}+A_{23}+A_{32})\cdot(-B_{23}+e^{-2}B_{31})\\
M_{13} &=& (-A_{21}+eA_{22}+e^{-2}A_{31}-e^{-1}A_{32})\cdot e^{-2}B_{21}\\
M_{14} &=& (-A_{11}+eA_{12}+e^2A_{13}-e^{-2}A_{31}+e^{-1}A_{32})\cdot e^{-1}B_{31}\\
M_{15} &=& e^{-2}A_{33}\cdot(e^{-2}B_{32}-e^2B_{33})\\
M_{16} &=& (-e^2A_{12}+e^{-2}A_{21}-A_{32})\cdot(B_{11}+e^{-1}B_{21}-B_{22}+e^{-2}B_{31})\\
M_{17} &=& e^{-1}A_{31}\cdot(-eB_{12}+e^3B_{13}+e^{-1}B_{21}-2e^{3}B_{33})\\
M_{18} &=& (-e^{-1}A_{21}+A_{22}+eA_{23})\cdot(B_{23}+e^{-1}B_{33})\\
M_{19} &=& (-A_{21}+e^2A_{23}+e^{-2}A_{33})\cdot e^{-2}B_{32}\\
M_{20} &=& e^{-2}A_{21}\cdot(eB_{12}+B_{22}+e^{-1}B_{32}+e^3B_{33})\\\\
C_{11} &=& -eM_2+eM_3-e^{-1}M_4+M_6-e^{-1}M_7-eM_9-eM_{13}+e^{-1}M_{14}\\
C_{12} &=& e^{-2}M_1+e^{-2}M_4-e^{-2}M_5-M_6-M_{10}+e^{-1}M_{15}+e^{-2}M_{16}+e^{-2}M_{17}+e^{-2}M_{20}\\
C_{13} &=& e^{-2}M_2-e^{-2}M_3-e^{-2}M_6+e^{-2}M_8+e^{-2}M_9-e^{-2}M_{10}+e^{-2}M_{11}-e^{-2}M_{12}+e^{-2}M_{13}e^{-2}M_{14}-e^{-2}M_{15}+e^{-2}M_{16}+e^{-2}M_{17}+e^{-2}M_{20}\\
C_{21} &=& -e^2M_1-eM_2+(e-e^4)M_3+e^2M_5+(-e+e^2)M_8-eM_9+eM_{10}-eM_{11}+eM_{12}+e^4M_{13}-eM_{14}-e^3m_{15}\\
C_{22} &=& M_2+(1-e^3)M_9-M_{15}+e^2M_{18}+m_{19}+eM_{20}\\
C_{23} &=& eM_2+eM_{11}-eM_{15}+eM_{19}\\
C_{31} &=& e^4M_1+e^4M_4+(-e^2-e^3)M_6+eM_7+eM_{17}\\
C_{32} &=& e^2M_6+e^2M_{10}+e^4M_{15}-M_{17}\\
C_{33} &=& -M_2+M_3+M_6-M_9+M_{10}-M_{13}\\
\end{eqnarray*}
By Master Theorem, the time complexity of Smirnov's 333 Algorithm is $O(n^{\log\limits_320}) = O(n^{2.726833})$.

\section{Error Analysis \cite{BLR80}}

\begin{theorem}[Bini]
The error produced by an APA-algorithm, is $O(2^{-d/\omega})$, where $\omega = 1 + \varphi/\sigma$ is a stability parameter depending on the border basis.
Where $O(\varepsilon^{-\varphi})$ is the largest infinite triads of the basis, i.e., $$\varphi = max\{z|u^{(r)}_i(\varepsilon)v^{(r)}_j(\varepsilon)w^{(r)}_s(\varepsilon)=O(\varepsilon^{-z}),i=1,2,...,m,j=1,2,...,n,s=1,2,...,p,r=1,2,...,t_B\}$$
and $O(\varepsilon^\sigma)$ is the slowest infinitesimal in $E(\varepsilon)$, i.e., $$\sigma = min\{z|e^{(s)}_{ij}=O(\varepsilon^z),i=1,2,...,m,j=1,2,...,n,s=1,2,...,p\}$$
The order of the error is minimized as above when $\varepsilon (d) = 2^{-d/(\varphi + \sigma)}$.
\end{theorem}
$d$ here represents the precision of the computer architecture. Single precision gives 23 bits precision and double gives 52 bits (without the hidden bit).

\section{Numerical Accuracy Calculation}
verify2.py is a file that could generate parameters $\sigma, \varphi$ which helps us to determinate the accuracy of each algorithm. \\
It will return 2 numbers sigma and phi, corresponding to $\sigma$ and $\varphi$ above. We can get $\sigma$ by searching all the error entries, and pick the minimum number of the exponent of the most dominate $\epsilon$ term from each entry of the error tensor $e_{ij}^{(s)}$. Notice that since $\epsilon < 1$, then the less the exponent of $\epsilon$ is, the more dominate it is. Thus, $\sigma \geq 1$ since negative term are cancelled. And $\varphi$ is the dominant term of $\epsilon$ in each slice of the tensor, which equals to the number of recursive calls in each iteration, in the UVW notation it is the $r$ dimension. Since $\epsilon < 1$, the most negative exponential term is the dominant one, which has the maximum absolute value.\\
Therefore, to get a better precision, we want the error to be as small as possible, which means we want $O(2^{-d/\omega})$ to be small. Since $d$ is decided by computer architecture, we want $\omega = 1+\varphi / \sigma$ to be as small as possible. Then we want $\varphi$ to be small, $\sigma$ to be large.\\
In Bini's paper, he didn't mention the case that if we apply the algorithm for multiple steps. Here we have the conclusion that $\sigma$ for any number of steps are the same, but the $\varphi$ for $s$ steps $\varphi_s = s * \varphi$. Thus, the error is $O(2^{-d\sigma/(s\varphi + \sigma)})$, and $\varepsilon (d,s) = O(2^{-d/(s\varphi + \sigma)})$.\\
Notice that in the table below, $-d/\omega$ equals the bits of error, and assuming single precision. 
\begin{center}
 \begin{tabular}{||c c c c c c c c||} 
 \hline
 Algorithm & $\sigma$ & $\varphi$ & \textit{error} bits & \textit{error} digits & $\varepsilon$ digits & $\varepsilon$ digits & speedup \\
  & & & (base 2) & (base 10) & (1 step) & (2 steps) & $(m*n*k)/r$\\ [0.5ex] 
 \hline\hline
 Bini & 1 & 1 & -11.5 & -3.46 & -3.46 & -2.31 & 1.2\\
 \hline
 Schonhage & 1 & 2 & -7.67 & -2.31 & -2.31 & -1.38 & 1.286\\
 \hline
 Smirnov 224 & 1 & 2 & -7.67 & -2.31 & -2.31 & -1.38 & 1.231\\
 \hline
 Smirnov 225 & 1 & 3 & -5.75 & -1.73 & -1.73 & -0.99 & 1.25\\
 \hline
 Smirnov 323 & 1 & 3 & -5.75 & -1.73 & -1.73 & -0.99 & 1.286\\
 \hline
 Smirnov 334 & 1 & 3 & -5.75 & -1.73 & -1.73 & -0.99 & 1.333\\
 \hline
 Smirnov 442 & 1 & 3 & -5.75 & -1.73 & -1.73 & -0.99 & 1.333\\
 \hline
 Smirnov 444 & 1 & 3 & -5.75 & -1.73 & -1.73 & -0.99 & 1.391\\
 \hline
 Smirnov 552 & 1 & 3 & -5.75 & -1.73 & -1.73 & -0.99 & 1.351\\
 \hline
 Smirnov 555 & 1 & 3 & -5.75 & -1.73 & -1.73 & -0.99 & 1.389\\
 \hline
 Smirnov 272 & 1 & 5 & -3.83 & -1.15 & -1.15 & -0.63 & 1.273\\
 \hline
 Smirnov 333 & 1 & 6 & -3.29 & -0.99 & -0.99 & -0.53 & 1.35\\ [1ex] 
 \hline
\end{tabular}
\end{center}
Also, this file will generate the indexes of corresponding entries from each matrix.

\bibliographystyle{alpha}

\bibliography{refs}

\end{document}

### 2021-02-27 latex/MPP21/paper.tex
added structure to paper

@@
 \newcommand{\GB}[1]{\textcolor{red}{\textbf{GB}: #1}}
+\newcommand{\JW}[1]{\textcolor{blue}{\textbf{JW}: #1}}
+\newcommand{\RZ}[1]{\textcolor{purple}{\textbf{RZ}: #1}}
 
@@
 
-\GB{reference this paper: \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=1675476}}
+\GB{motivate importance of matmul in NN}
 
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+\GB{emphasize unseen potential of APA algorithms}
 
-\section{Ease of Use}
+\GB{find references for low-precision arithmetic inside NN}
 
-\subsection{Maintaining the Integrity of the Specifications}
+\GB{reference this paper: \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=1675476} as example use of APA algorithms in practice}
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
+\section{Arbitrary Precision Approximating Algorithms}
 
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
+\begin{enumerate}
+	\item complexity of matmul discussion
+	\item describe nature of approximation
+	\item give explicit algorithm (Bini?)
+	\item table of 12 algorithms
+	\item accuracy plot?
+\end{enumerate}
 
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
+\section{Parallel Fast Matrix Multiplication}
 
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
+\GB{reference PPoPP paper: \cite{BB15}}
 
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
+\begin{enumerate}
+	\item describe code generation and hybrid parallelization
+	\item give performance plots (sequential and parallel)
+\end{enumerate}
 
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
+\section{Neural Network Accuracy}
 
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
+\begin{enumerate}
+	\item describe network (MNIST?) and implementation details
+	\item accuracy results in table or fig
+\end{enumerate}
 
-\subsection{\LaTeX-Specific Advice}
+\section{Neural Network Performance}
 
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
+\begin{enumerate}
+	\item describe synthetic network and implementation details
+	\item perf results for growing dim (maybe for serial and 12-threads separately)
+	\item parallel speedup results
+\end{enumerate}
 
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
+\section{Conclusion}
 
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
+\begin{enumerate}
+	\item promising results for synthetic network, goal is to target costly networks bottlenecked by matmul
+	\item can tune for matmul dimensions (not just square)
+	\item GPU implementation (cite Jianyu Huang's work)
+	\item mixed precision fast matmul (APA or otherwise)
+\end{enumerate}
 
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
+\bibliographystyle{IEEEtran}
+\bibliography{refs}
 

### 2021-02-27 latex/MPP21/paper.tex
added table of algs and started to clean up references
working on first plots

@@
 \usepackage{cleveref}
+
+% for subfigures
+\usepackage[caption=false,font=footnotesize]{subfig}
+
+% for scientific notation like 1e-8
+\usepackage{siunitx}
+\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} } 
+\newcommand{\dims}[1]{\langle #1 \rangle}
+
+\newcommand{\datafile}{}
 
@@
 \IEEEauthorblockN{Jack Weissenberger}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.) \GB{?}} \\
-\textit{Capital One \GB{?}}\\
-Richmond, VA \\
-email address or ORCID}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Wake Forest University}\\
+Winston Salem, NC, USA \\
+\email{jack.weissenberger@gmail.com}}
 \and
@@
 
+\begin{table}
+\centering
+\caption{Properties of Arbitrary Precision Approximating Algorithms  \\ Speedup and Error are computed assuming 1 recursive step}
+\label{tab:algs}
+\begin{tabular}{| c | c c c | c c c |} 
+\hline
+\textbf{Ref} & \textbf{Dims} & \textbf{Rank} & \textbf{Speedup} & $\mathbf{\sigma}$ & $\mathbf{\varphi}$ & \textbf{Error} \\
+\hline
+- & $\dims{2,2,2}$ & 8 & - & 1 & 0 & \num{1.2e-7} \\
+\hline
+\cite{BCRL79} & $\dims{3,2,2}$ & 10 & $20\%$ & 1 & 1 & \num{3.5e-4} \\
+\cite{Smirnov13} & $\dims{4,2,2}$ & 13 & $23\%$ & 1 & 2 & \num{4.9e-3} \\
+\cite{Smirnov13} & $\dims{3,3,2}$ & 14 & $29\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov13} & $\dims{5,2,2}$ & 16 & $25\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov13} & $\dims{3,3,3}$ & 20 & $35\%$ & 1 & 6 & \num{1.0e-1} \\
+\cite{Schonhage81} & $\dims{3,3,3}$ & 21 & $29\%$ & 1 & 2 & \num{4.9e-3} \\
+\cite{Smirnov15} & $\dims{7,2,2}$ & 22 & $27\%$ & 1 & 5 & \num{7.0e-2} \\
+\cite{Smirnov13} & $\dims{4,4,2}$ & 24 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov16} & $\dims{4,3,3}$ & 27 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov18} & $\dims{5,5,2}$ & 37 & $35\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov14} & $\dims{4,4,4}$ & 46 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov18} & $\dims{5,5,5}$ & 90 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
+\hline
+\end{tabular}
+\end{table}
+
+\GB{specify formula for speedup and error values...}
+
 \section{Parallel Fast Matrix Multiplication}
@@
 
+\begin{figure}
+\subfloat[One thread]{
+\renewcommand{\datafile}{data/NNtime_1t.dat}
+\input{plots/reltime}
+\label{fig:NNreltime_seq}}
+\hfill
+\subfloat[Twelve threads]{
+csv files should have columns n,DGEMM,Bini,... and row for each dim
+\label{fig:NNreltime_par}}
+\caption{Network training time relative to using classical matrix multiplication}
+\label{fig:NNreltime}
+\end{figure}
+
 \section{Conclusion}

### 2021-02-27 TF_op/nn_test.sh
update nn_test for 3 threads

@@ -1,16 +1,14 @@
 
 for size in 512 1024 2048 4096 8192
 do
-#for name in bini322 dgemm classic schonhage333 smirnov224 smirnov225 smirnov272 smirnov323 smirnov333 smirnov334 smirnov442 smirnov444 smirnov552 smirnov555
-#do
-for thread in 1 12
+for name in bini322 dgemm classic schonhage333 smirnov224 smirnov225 smirnov272 smirnov323 smirnov333 smirnov334 smirnov442 smirnov444 smirnov552 smirnov555
 do
 
-python tensorboard_test.py --layers 4 --nodes ${size} --epochs 5 --bs ${size} --mm dgemm --logdir FINAL_${thread}_thread_dgemm${size} --num_threads ${thread}
+
+python tensorboard_test.py --layers 4 --nodes ${size} --epochs 5 --bs ${size} --mm ${name} --logdir PAPER_1_thread_${name}${size} --num_threads 1
 
 done
 done
-#done
 
 
 #python tensorboard_test.py --layers 4 --nodes 512 --epochs 5 --bs 512 --mm smirnov442 --logdir test

### 2021-02-28 latex/MPP21/paper.tex
cleaned up references so far
cleaned up refs a bit more
made progress on introduction and added a bunch of refs

@@
 \usepackage{cleveref}
+\usepackage{tikz,pgfplots}
 
@@
 
-\GB{motivate importance of matmul in NN}
+\emph{Fast} matrix multiplication algorithms are those that perform fewer than the $2n^3+O(n^2)$ flops performed by the classical algorithm.
+For example, Strassen's original fast matrix multiplication algorithm performs $O(n^{\log_2 7})$ flops, where $\log_27 \approx 2.81$ \cite{Strassen69}.
+The true complexity of matrix multiplication, typically measured as the exponent $\omega$ for complexity $O(n^\omega)$ is an open question, but the current tightest upper bound is 2.37286 \cite{AW21}.
+Upper bounds in this range correspond to theoretical algorithms that are not expected to be practical.
 
-\GB{emphasize unseen potential of APA algorithms}
+One of the reasons such algorithms are impractical is that they are based on so-called \emph{Arbitrary Precision Approximating} (APA) algorithms, which we describe in detail in \cref{sec:APA}.
+In exact arithmetic, these algorithms compute an approximation of the correct result, where the error is polynomial in a nonzero parameter of the algorithm \cite{BLR80}.
+That is, in exact arithmetic, the error can be made arbitrarily small.
+In floating point arithmetic, however, there is a lower bound on the approximation error that depends on the working precision and properties of the algorithm.
+In theory, APA algorithms can be converted to exact algorithms at the cost of an extra logarithmic factor of $n$, which is typically hidden by an arbitrarily small increase in the exponent \cite{Bini80}.
+In practice, however, $n$ is not large enough to ignore the logarithmic factor, and floating point error plagues the conversion method, so APA algorithms are considered approximate algorithms with insufficient accuracy for most applications.
+Thus, APA algorithms are largely overlooked as practical tools despite their performance potential \cite{BB15}.
 
-\GB{find references for low-precision arithmetic inside NN}
+Our goal in this paper is to demonstrate that APA algorithms can offer practical performance improvements for applications that are tolerant to error in matrix multiplications, notably the training phase of neural networks.
+Training large neural networks is computationally expensive and the cost has spurred a surge of research into more efficient hardware, better algorithms, and techniques for trading off accuracy for performance.
+For instance, low-precision arithmetic has been shown to decrease running time with little to no effect on the ultimate learning task \cite{GAGN15,HCSEB17}, and new floating point formats have been developed and supported in hardware to implement highly efficient low- and mixed-precision computation \cite{KM+19,YWC20}.
 
-\GB{reference this paper: \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=1675476} as example use of APA algorithms in practice}
+Matrix multiplication in particular is a bottleneck computation for many neural networks... \GB{continue here, make the case that matmul is bottleneck}
+
+\GB{reference this paper \cite{KAA20} as using Strassen-Winograd in a single-layer feed-forward network on a GPU}
+
+\GB{reference this paper \cite{NPOV15} for tensorizing large fully-connected layers, including layers with dimensions 25K x 4K and 4K x 4K}
+
+\GB{list of contributions:}
+
+\begin{enumerate}
+	\item curating a collection of recently discovered practical APA algorithms
+	\item extending the framework of \cite{BB15} to generate efficient multithreaded code for a more general set of APA algorithms
+	\item demonstrating the robustness of learning accuracy to approximate matrix multiplications
+	\item presenting multithreaded performance improvements of a synthetic MLP of up to $Y\times$
+\end{enumerate}
 
 \section{Arbitrary Precision Approximating Algorithms}
+\label{sec:APA}
 
@@
 \end{enumerate}
+
+\GB{cite \cite{Bini80} for conversion of APA algorithm to exact algorithm}
 
@@
 \cite{BCRL79} & $\dims{3,2,2}$ & 10 & $20\%$ & 1 & 1 & \num{3.5e-4} \\
-\cite{Smirnov13} & $\dims{4,2,2}$ & 13 & $23\%$ & 1 & 2 & \num{4.9e-3} \\
+\cite{AS13} & $\dims{4,2,2}$ & 13 & $23\%$ & 1 & 2 & \num{4.9e-3} \\
 \cite{Smirnov13} & $\dims{3,3,2}$ & 14 & $29\%$ & 1 & 3 & \num{1.9e-2} \\
@@
 \cite{Smirnov15} & $\dims{7,2,2}$ & 22 & $27\%$ & 1 & 5 & \num{7.0e-2} \\
-\cite{Smirnov13} & $\dims{4,4,2}$ & 24 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
-\cite{Smirnov16} & $\dims{4,3,3}$ & 27 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
-\cite{Smirnov18} & $\dims{5,5,2}$ & 37 & $35\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov16b} & $\dims{4,4,2}$ & 24 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov16a} & $\dims{4,3,3}$ & 27 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov16b} & $\dims{5,5,2}$ & 37 & $35\%$ & 1 & 3 & \num{1.9e-2} \\
 \cite{Smirnov14} & $\dims{4,4,4}$ & 46 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
@@
 \subfloat[One thread]{
+\centering
 \renewcommand{\datafile}{data/NNtime_1t.dat}
@@
 \label{fig:NNreltime_seq}}
-\hfill
+\\
 \subfloat[Twelve threads]{
-csv files should have columns n,DGEMM,Bini,... and row for each dim
+\centering
+\renewcommand{\datafile}{data/NNtime_12t.dat}
+\input{plots/reltime}
 \label{fig:NNreltime_par}}

### 2021-02-28 latex/MPP21/paper.tex
finished draft of introduction

@@
 In practice, however, $n$ is not large enough to ignore the logarithmic factor, and floating point error plagues the conversion method, so APA algorithms are considered approximate algorithms with insufficient accuracy for most applications.
-Thus, APA algorithms are largely overlooked as practical tools despite their performance potential \cite{BB15}.
+Thus, APA algorithms are largely overlooked as practical tools despite their performance potential and ability to outperform their exact counterparts \cite{BB15}.
 
@@
 
-Matrix multiplication in particular is a bottleneck computation for many neural networks... \GB{continue here, make the case that matmul is bottleneck}
+Matrix multiplication in particular is a bottleneck computation for many neural networks.
+Forward and backward propagation in training the weights of fully connected layers is a matrix multiplication with dimensions given by the sizes of the layers and the size of the batch.
+Training convolutional and other types of layers can also be cast as matrix multiplication, either via monolithic multiplications or batches of smaller multiplications \cite{CW+14,GB+19}.
+In this paper we focus on Multi-Layer Perceptron (MLP) networks that rely on a sequence of fully connected layers \cite{HSW89}.
+Because the sizes of the layers in MLP networks continue to grow, various techniques have been used to reduce the computational demands of the training phase.
+For example, low-rank tensor approximation of the weights can reduce both memory and computation \cite{NPOV15}, and fast matrix multiplication (the Strassen-Winograd algorithm) has been applied to the bottleneck matrix multiplications \cite{KAA20}.
 
-\GB{reference this paper \cite{KAA20} as using Strassen-Winograd in a single-layer feed-forward network on a GPU}
+Our contribution is the use of APA matrix multiplication algorithms to address this problem and accelerate training.
+In particular we
+\begin{enumerate}
+	\item curate a collection of well-established and recently discovered practical APA algorithms;
+	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them;
+	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications; and
+	\item present multithreaded performance improvements of a synthetic MLP of up to \GB{$Y\times$}.
+\end{enumerate}
 
-\GB{reference this paper \cite{NPOV15} for tensorizing large fully-connected layers, including layers with dimensions 25K x 4K and 4K x 4K}
-
-\GB{list of contributions:}
-
-\begin{enumerate}
-	\item curating a collection of recently discovered practical APA algorithms
-	\item extending the framework of \cite{BB15} to generate efficient multithreaded code for a more general set of APA algorithms
-	\item demonstrating the robustness of learning accuracy to approximate matrix multiplications
-	\item presenting multithreaded performance improvements of a synthetic MLP of up to $Y\times$
-\end{enumerate}
+\GB{\cite{NPOV15} uses a network including layers with dimensions 25K x 4K and 4K x 4K}
 

### 2021-02-28 latex/MPP21/paper.tex
progress on section II (APA algs)

@@
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+\GB{TBD}
 \end{abstract}
-
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
 
@@
 In floating point arithmetic, however, there is a lower bound on the approximation error that depends on the working precision and properties of the algorithm.
-In theory, APA algorithms can be converted to exact algorithms at the cost of an extra logarithmic factor of $n$, which is typically hidden by an arbitrarily small increase in the exponent \cite{Bini80}.
-In practice, however, $n$ is not large enough to ignore the logarithmic factor, and floating point error plagues the conversion method, so APA algorithms are considered approximate algorithms with insufficient accuracy for most applications.
-Thus, APA algorithms are largely overlooked as practical tools despite their performance potential and ability to outperform their exact counterparts \cite{BB15}.
+Thus, APA algorithms are often considered to have insufficient accuracy for most applications and have largely been overlooked as practical tools despite their performance potential and ability to outperform their exact counterparts \cite{BB15}.
 
@@
 
-\section{Arbitrary Precision Approximating Algorithms}
+\section{Practical Arbitrary Precision Approximating Matrix Multiplication Algorithms}
 \label{sec:APA}
@@
 
-\GB{cite \cite{Bini80} for conversion of APA algorithm to exact algorithm}
+\subsection{Fast Matrix Multiplication}
+
+Nearly all fast matrix multiplication algorithms are based a rule for multiplying matrices of fixed size, and the reduction in asymptotic complexity stems from using the rule recursively on general matrices.
+For example, Strassen's algorithm is specified by a rule for multiplying two $2\times2$ matrices (denoted $\dims{2,2,2}$) using 7 multiplications instead of the classical algorithm's 8 multiplications.
+Applying the rule to $n\times n$ matrices, we split each matrix into quadrants, and the 7 multiplications are multiplications of $(n/2)\times (n/2)$ matrices.
+For even better efficiency, we can consider larger fixed sizes and find rules that require a lower percentage of multiplications compared to the classical rule.
+The number of multiplications in a rule is known as the \emph{rank}, so Strassen's is a rank-7 algorithm.
+Algorithms have been derived both analytically and computationally, and there exists a vast set of improvements leading to the current world record \cite{Pan84,CW87,Williams12,AW21}.
+
+\subsection{APA Algorithms}
+
+A key characteristic of matrix multiplication, first demonstrated by Bini et al.~\cite{BCRL79}, is that it can be approximated to arbitrary accuracy more efficiently than being computed exactly.
+An APA algorithm is one that takes as input $A$, $B$, and a parameter $0<\lambda<1$ and computes 
+\begin{equation}
+\label{eq:APAapprox}
+\hat C = A\cdot B + \lambda E + O(\lambda^2).
+\end{equation}
+Thus, in exact arithmetic, letting $\lambda \rightarrow 0$ achieves arbitrarily small approximation error.
+In floating point arithmetic, choosing too small a value for $\lambda$ leads to accumulation of roundoff error that exceeds the approximation error.
+We discuss optimizing $\lambda$ in \cref{sec:APAerr}.
+
+In theory, APA algorithms can be converted to exact algorithms at the cost of an extra logarithmic factor of $n$, which is typically hidden by an arbitrarily small increase in the exponent \cite{Bini80}.
+In practice, $n$ is not large enough to ignore the logarithmic factor, so we consider each APA algorithm as is.
+
+For a concrete example, we reproduce the rule for the algorithm developed by Bini et al.~\cite{BCRL79} for the $\dims{3,2,2}$ case (multiplying a $3\times 2$ matrix $A$ by a $2\times 2$ matrix $B$), where we use the following notation for input and output matrices:
+$$\begin{bmatrix} C_{11} & C_{12} \\ C_{21} & C_{22} \\ C_{31} & C_{32} \end{bmatrix} = 
+\begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \\ A_{31} & A_{32}  \end{bmatrix} \cdot
+\begin{bmatrix} B_{11} & B_{12} \\ B_{21} & B_{22} \end{bmatrix}.$$
+Bini's rule is given by
+\begin{align*}
+M_1 =& (A_{11} + A_{22}) \cdot (\lambda B_{11} + B_{22}) \\
+M_2 =& A_{22}\cdot  (-B_{21} - B_{22}) \\
+M_3 =& A_{11}\cdot B_{22} \\
+M_4 =& (\lambda A_{12} + A_{22})\cdot (-\lambda B_{11} + B_{21}) \\
+M_5 =& (A_{11} + \lambda A_{12}) \cdot (\lambda B_{12} + B_{22}) \\
+M_6 =& (A_{21} + A_{32}) \cdot (B_{11} + \lambda B_{22}) \\
+M_7 =& A_{21} \cdot (-B_{11} - B_{12}) \\
+M_8 =& A_{32} \cdot B_{11}\\
+M_9 =& (A_{21} + \lambda A_{31}) \cdot (B_{12} - \lambda B_{22})\\
+M_{10} =& (\lambda A_{31} + A_{32}) \cdot (B_{12} - \lambda B_{22})\\
+\end{align*}
+\begin{align*}
+C_{11} =& \lambda^{-1} (M_1 + M_2 - M_3 + M_4) \\
+C_{12} =& \lambda^{-1} (-M_3 + M_5) \\
+C_{21} =& M_4 + M_6 - M_{10}\\
+C_{22} =& M_1 - M_5 + M_9\\
+C_{31} =& \lambda^{-1} (-M_8 + M_{10}) \\
+C_{32} =& \lambda^{-1} (M_6 + M_7 - M_8 + M_9).\\
+\end{align*}
+
+We highlight several properties of the rule which are common across all APA algorithms we consider.
+First, the rule requires fewer multiplications than the classical one (rank 10 instead of 12 for $\dims{3,2,2}$ here).
+Next, each multiplication is between a linear combination of entries of $A$ and a linear combination of entries of $B$, each output entry is computed as a linear combination of the outputs of the multiplications, and each coefficient in the linear combinations is a (Laurent) polynomial in $\lambda$.
+The coefficients include both positive and negative powers of $\lambda$, which explains why small values of $\lambda$ can lead to significant roundoff error.
+For Bini's algorithm all coefficients are monomial with degree between $-1$ and $1$.
+
+Because of this general pattern, we can encode APA and other fast algorithms succinctly by their linear combination coefficients.
+For example, encoding the first multiplication $M_1$ in Bini's algorithm can be done using a triplet of matrices:
+\begin{equation}
+\label{eq:triplet}
+\begin{bmatrix} 1 & 0 \\ 0 & 1 \\ 0 & 0 \end{bmatrix}, \hfill
+\begin{bmatrix} \lambda & 0 \\ 0 & 1 \end{bmatrix}, \hfill
+\begin{bmatrix} \lambda^{-1} & 0 \\ 0 & 1 \\ 0 & 0 \end{bmatrix}.
+\end{equation}
+The first two matrices specify the linear combinations taken of entries of $A$ and $B$, and the third matrix specifies the contributions of $M_1$ to the entries of $C$.
+Ten such triplets completely specify Bini's algorithm.
+
+\subsection{Numerical Error of APA Algorithms}
+\label{sec:APAerr}
+
+\GB{TBD...}
+\begin{enumerate}
+	\item define working precision $2^{-d}$
+	\item define $\varphi$ and $\sigma$ and give intuition
+	\item define optimal $\lambda=O(2^{-d/(\varphi+\sigma)})$
+	\item define error given optimal $\lambda$: $O(2^{-d\sigma/(\varphi+\sigma)})$
+\end{enumerate}
+
+\subsection{Practical Algorithms}
+
+We are particularly interested in algorithms with rules for \emph{small} fixed sizes because they have more promise for practical performance.
+This is because larger fixed sizes result in multiplications of small submatrices, and matrix multiplication performance degrades for smaller dimensions.
+For instance, consider a rule for dimensions $\dims{4,4,4}$ applied to matrices of reasonable size, less than dimension $10{,}000$.
+After one recursive call, the submatrices are of size less than 2500, and after two recursive calls, the dimension is less than 625.
+At this size, the reduction in number of flops is offset by a reduction in performance, which may result in longer running time.
+
+Instead of focusing on the exponent of the asymptotic complexity of fast algorithms, for practical algorithms we are more interested in the constant reduction in flops of a single recursive level (a single use of the rule of the algorithm).
+This is because in practice, for reasonable matrix dimensions (no more than 10,000, say), only 1 or 2 recursive levels will yield performance improvement \cite{BB15}.
+We also prefer algorithms with fewer nonzero coefficients in the linear combinations, because while less costly than multiplications, the matrix additions are less efficient (they are memory bandwidth bound) and prevent achieving the ideal speedup given by the reduction in multiplications.
+For dimensions $\dims{m,n,k}$ and rank $r$, the ideal speedup for a single recursive step is given by $mnk/r$, and two recursive steps would enable a possible speedup of $(mnk/r)^2$.
+These speedups are typically not attained because of degradation in performance for smaller matrix dimensions and the overhead of matrix additions.
+
+\subsection{APA Algorithm Properties}
 
@@
 \end{table}
-
-\GB{specify formula for speedup and error values...}
 

### 2021-03-01 latex/MPP21/paper.tex
finished draft of APA section

@@
 
-\GB{TBD...}
-\begin{enumerate}
-	\item define working precision $2^{-d}$
-	\item define $\varphi$ and $\sigma$ and give intuition
-	\item define optimal $\lambda=O(2^{-d/(\varphi+\sigma)})$
-	\item define error given optimal $\lambda$: $O(2^{-d\sigma/(\varphi+\sigma)})$
-\end{enumerate}
+In floating point arithmetic, the lower bound on the numerical error of APA algorithms depends on the working precision and two parameters of the algorithm \cite{BLR80}.
+The working precision, also referred to as machine precision, is the upper bound on relative error incurred by basic operations in a given floating point format and depends on the number of fractional bits used in the format.
+We use the notation $2^{-d}$ for working precision, where $d=52$ for double and $d=23$ for single precision (note that $2^{-52}\approx 10^{-16}$ and $2^{-23}\approx 10^{-7}$).
+
+The two parameters of the APA algorithm specify the contribution of the approximation and roundoff errors, respectively.
+The first parameter, $\sigma$, is the smallest positive exponent of the error polynomial and represents the approximation error.
+\Cref{eq:APAapprox} shows the error as a polynomial of $\lambda$ whose leading term is linear in $\lambda$.
+If an algorithm satisfies \cref{eq:APAapprox} with $E\neq 0$, then $\sigma=1$.
+However, if $E=0$, then $\sigma>1$ is the degree of the leading monomial.
+Larger $\sigma$ implies smaller error due to the algorithm, though all of the APA algorithms we consider have $\sigma=1$.
+
+The second parameter, $\varphi$, is the largest (in absolute value) negative exponent of the algorithm, computed as the largest sum of negative exponents across all triplets of matrices.
+This parameter represents the effect of roundoff error caused by floating point arithmetic involving the largest intermediate values computed by the algorithm.
+For example, the triplet given in \cref{eq:triplet} yields a sum of negative exponents of $0+0+1=1$, and in the case of Bini's algorithm, no other triplet has a larger sum, so $\varphi=1$ for that algorithm.
+Smaller $\varphi$ implies smaller error due to roundoff, and the APA algorithms we consider exhibit a range of values.
+
+Given these two contributions to the numerical error, $\lambda$ can be optimized to balance the effects based on parameters $\sigma$ and $\varphi$ (and $d$).
+As shown by Bini, Lotti, and Romani \cite{BLR80}, the optimal $\lambda$ should be set to $\Theta(2^{-d/(\sigma+\varphi)})$.
+Using this value of $\lambda$, the numerical error incurred by the algorithm will be bounded by $O(2^{-d\sigma/(\sigma+\varphi)})$.
+Taking Bini's algorithm as an example, we have $\sigma=\varphi=1$, so the error is $O(2^{-d/2})$, or the square root of working precision.
+Note that if multiple recursive steps are used, then $\varphi$ increases proportional to the number of steps, so straightforward optimization of $\lambda$ results in error of $O(2^{-d\sigma/(\sigma+s\varphi)})$ for $s$ recursive steps.
 
@@
 For dimensions $\dims{m,n,k}$ and rank $r$, the ideal speedup for a single recursive step is given by $mnk/r$, and two recursive steps would enable a possible speedup of $(mnk/r)^2$.
-These speedups are typically not attained because of degradation in performance for smaller matrix dimensions and the overhead of matrix additions.
+These speedups are typically not fully attained because of degradation in performance for smaller matrix dimensions and the overhead of matrix additions.
+In the experimental results of this work, we use only 1 recursive step for all algorithms.
 
 \subsection{APA Algorithm Properties}
+
+\Cref{tab:algs} shows the key performance and accuracy properties of the APA algorithms we consider.
+Each row corresponds to an algorithm, and the first row includes the classical algorithm for comparison.
+For all algorithms, we assume only 1 recursive step is used, though the speedup and error for more steps can be readily calculated from the parameters.
+The first column gives the reference where the algorithm was first specified.
+The second block column demonstrates the possible performance improvement, where speedup is calculated as $(mnk/r - 1)\cdot100\%$ for $\dims{m,n,k}$ and rank $r$.
+The third block column shows error parameters, with error calculated as $2^{-d\cdot\sigma/(\sigma+\varphi)}$ with $d=23$, corresponding to single precision.
 

### 2021-03-01 latex/MPP21/paper.tex
minor update to comments in latex
added text to section III, still need to add fig, performanc results, and write up analysis

@@
 
-\begin{enumerate}
-	\item complexity of matmul discussion
-	\item describe nature of approximation
-	\item give explicit algorithm (Bini?)
-	\item table of 12 algorithms
-	\item accuracy plot?
-\end{enumerate}
-
 \subsection{Fast Matrix Multiplication}
@@
 \subsection{APA Algorithms}
+\label{sec:APAalgs}
 
@@
 \label{sec:APAerr}
+
+\GB{should we add an accuracy plot in this section?}
 
@@
 
-\GB{reference PPoPP paper: \cite{BB15}}
+We exploit the common algorithmic structure across APA (and exact) algorithms to develop a unified strategy of high performance multithreaded implementation using C++/OpenMP.
+We build upon the work of Benson and Ballard \cite{BB15}, using code generation to apply the strategy to each algorithm.
 
-\begin{enumerate}
-	\item describe code generation and hybrid parallelization
-	\item give performance plots (sequential and parallel)
-\end{enumerate}
+\subsection{Hybrid Parallelization Strategy}
+
+As described in \cref{sec:APAalgs}, each algorithm can be encoded by a set of triplets of coefficient matrices.
+From this representation, we generate recursive code that computes the linear combinations for the inputs to each multiplication and combines the outputs.
+The linear combinations are computed using a ``write-once'' strategy that was found to be most efficient in terms of memory bandwidth and performance.
+The multiplications are either computed recursively or via a call to \texttt{gemm}, the interface to a highly efficient BLAS implementation of the classical algorithm.
+Because we use only 1 recursive step in our experiments, there are no recursive calls, and every multiplication is performed by \texttt{gemm}.
+We note that the original code generation tool was designed primarily for exact algorithms, and we generalized it to apply to all of the APA algorithms we consider here.
+
+We adopt the ``hybrid'' parallelization strategy proposed by Benson and Ballard \cite{BB15}.
+All linear combinations are parallelized in a straightforward way, in order to maximize the memory bandwidth of the machine.
+The multiplications are parallelized as follows: given $r$ multiplications and $p$ threads, with $r=p\cdot q + \ell$ for integers $q$ and $\ell<r$, we assign each thread $q$ multiplications to be performed independently, and the remaining $\ell$ multiplications are performed by all threads using the multithreaded implementation of \texttt{gemm}.
+\Cref{fig:hybrid} shows the hybrid strategy for $r=10$ (Bini's algorithm) and $p=4$.
+
+\begin{figure}
+\input{figs/hybrid}
+\caption{Illustration of hybrid parallelization strategy for $r=10$ and 4 threads.}
+\label{fig:hybrid}
+\end{figure}
+
+The hybrid strategy is efficient because each thread can achieve close to the peak performance of a core when computing independent matrix multiplications, even for relatively small problems.
+The alternative strategy of employing multithreaded \texttt{gemm} for each of the $r$ multiplications (known as ``DFS'') suffers performance degradation for small problems, where the parallel implementation attains a smaller fraction of peak.
+The hybrid strategy also perfectly load balances the computation across threads, as opposed to the alternate strategy of assigning the $\ell$ remainder multiplications to $\ell$ different threads (known as ``BFS''), leaving the other $p-\ell$ threads idle.
+
+\subsection{Sequential Performance}
+
+\GB{add sequential performance results for all 12 algorithms, compared to \texttt{gemm}}
+
+\subsection{Parallel Performance}
+
+\GB{add parallel performance results for all 12 algorithms, compared to \texttt{gemm}}
 

### 2021-03-01 latex/MPP21/paper.tex
added hybrid figure

@@
 \usepackage{tikz,pgfplots}
+\usetikzlibrary{positioning}
 
@@
 \begin{figure}
+\centering
 \input{figs/hybrid}
-\caption{Illustration of hybrid parallelization strategy for $r=10$ and 4 threads.}
+\caption{Illustration of hybrid parallelization strategy for $r=10$ and 4 threads.
+Each thread is assigned two multiplications to compute using single-threaded \texttt{gemm} and the two remaining multiplications are performed using multithreaded \texttt{gemm}.}
 \label{fig:hybrid}

### 2021-03-01 latex/MPP21/paper.tex
added dummy accuracy plot and data file

@@
 \and
-\IEEEauthorblockN{Luoping\GB{?} Zhang}
+\IEEEauthorblockN{Luoping Zhang}
 \IEEEauthorblockA{\textit{Department of Computer Science} \\
@@
 Note that if multiple recursive steps are used, then $\varphi$ increases proportional to the number of steps, so straightforward optimization of $\lambda$ results in error of $O(2^{-d\sigma/(\sigma+s\varphi)})$ for $s$ recursive steps.
+The parameters and minimum error for the algorithms we consider are presented in \cref{sec:APAprops}.
+We show empirical error results for uniform random inputs of varying dimension in \cref{fig:accuracy}, as compared to the classical algorithm.
+\GB{add more details on experiment here...}
+
+\begin{figure}
+\centering
+\renewcommand{\datafile}{data/accuracy.dat}
+\input{plots/accuracy}
+\caption{Relative Frobenius norm error for APA algorithms on random inputs.}
+\label{fig:accuracy}
+\end{figure}
 
@@
 \subsection{APA Algorithm Properties}
+\label{sec:APAprops}
 

### 2021-03-01 latex/MPP21/paper.tex
added abstract and dummy plots for matmul performance

@@
 \usepackage{xcolor}
+\usepackage{ifthen}
+
+% cleveref
 \usepackage{hyperref}
-\usepackage{cleveref}
+\usepackage[nameinlink]{cleveref}
+\crefname{section}{\S}{Sections}
+\crefname{figure}{Fig.}{Figs.}
+\crefname{table}{Table}{Tables}
+
+% tikz
 \usepackage{tikz,pgfplots}
@@
 
+% macros for plots
 \newcommand{\datafile}{}
+\newcommand{\threads}{}
 
@@
 \begin{abstract}
-\GB{TBD}
+Matrix multiplication is one of the bottleneck computations for training the weights within deep neural networks.
+To speed up the training phase, we propose to use faster algorithms for matrix multiplication known as Arbitrary Precision Approximating (APA) algorithms.
+APA algorithms perform asymptotically fewer arithmetic operations than the classical algorithm, but they compute an approximate result with an error that can be made arbitrarily small in exact arithmetic.
+Practical APA algorithms provide significant reduction in computation time and still provide enough accuracy for many applications like neural network training.
+We demonstrate that APA algorithms can be efficiently implemented and parallelized for shared-memory platforms to obtain up to \GB{Z\%} speedups over the fastest implementation of the classical algorithm.
+Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no signification reduction in the training or testing error.
+Our performance results on a large MLP network show overall performance improvements of up to \GB{W\%}.
 \end{abstract}
@@
 Matrix multiplication in particular is a bottleneck computation for many neural networks.
-Forward and backward propagation in training the weights of fully connected layers is a matrix multiplication with dimensions given by the sizes of the layers and the size of the batch.
+Forward and backward propagation in training the weights of fully connected layers requires matrix multiplication with dimensions given by the sizes of the layers and number of batch samples.
 Training convolutional and other types of layers can also be cast as matrix multiplication, either via monolithic multiplications or batches of smaller multiplications \cite{CW+14,GB+19}.
@@
 	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications; and
-	\item present multithreaded performance improvements of a synthetic MLP of up to \GB{$Y\times$}.
+	\item present multithreaded performance improvements of a synthetic MLP of up to \GB{$Y\times$} over the use of classical matrix multiplication.
 \end{enumerate}
@@
 
-\GB{should we add an accuracy plot in this section?}
-
 In floating point arithmetic, the lower bound on the numerical error of APA algorithms depends on the working precision and two parameters of the algorithm \cite{BLR80}.
@@
 The second block column demonstrates the possible performance improvement, where speedup is calculated as $(mnk/r - 1)\cdot100\%$ for $\dims{m,n,k}$ and rank $r$.
-The third block column shows error parameters, with error calculated as $2^{-d\cdot\sigma/(\sigma+\varphi)}$ with $d=23$, corresponding to single precision.
+The third block column shows error parameters, and the error is calculated as $2^{-d\cdot\sigma/(\sigma+\varphi)}$ with $d=23$, corresponding to single precision.
 
@@
 
+\begin{figure}
+\centering
+\renewcommand{\datafile}{data/matmul_seq.dat}
+\renewcommand{\threads}{1}
+\input{plots/matmul_perf}
+\caption{Single-threaded matrix multiplication performance}
+\label{fig:matmul_seq}
+\end{figure}
+
 \subsection{Parallel Performance}
@@
 \GB{add parallel performance results for all 12 algorithms, compared to \texttt{gemm}}
+
+\begin{figure}
+\centering
+\renewcommand{\datafile}{data/matmul_par.dat}
+\renewcommand{\threads}{12}
+\input{plots/matmul_perf}
+\caption{Multithreaded matrix multiplication performance}
+\label{fig:matmul_seq}
+\end{figure}
 

### 2021-03-03 latex/MPP21/paper.tex

@@
 
+\begin{enumerate}
+	\item confirmed that optimal lambda values are with a factor of 4 of theoretical (for these random inputs)
+	\item some errors were lower than expected (555 and 7222), we think this is because of fractions that appear in the coefficients of the largest negative powers of lambda
+	\item legend is ordered according to theoretical error
+	\item not much fluctuation over dimensions
+\end{enumerate}
+
 \begin{figure}

### 2021-03-03 latex/MPP21/paper.tex
added analysis of matmul accuracy data

--- 2021-03-01 latex/MPP21/paper.tex
@@
 The parameters and minimum error for the algorithms we consider are presented in \cref{sec:APAprops}.
+
 We show empirical error results for uniform random inputs of varying dimension in \cref{fig:accuracy}, as compared to the classical algorithm.
-\GB{add more details on experiment here...}
+We measure the relative Frobenius norm error, or $\|C-\hat C\|_F/\|C\|$, where $\hat C$ is computed by each algorithm and $C$ is computed using the classical algorithm in double precision.
+In order to choose the optimal $\lambda$ value for each algorithm, we tested the 5 powers of 2 closest to the theoretical optimal value and chose the best.
 
@@
 \end{figure}
+
+Overall, we see little fluctuation of the error over matrix dimension, and the theoretical error bound is an upper bound on all empirical errors.
+Note that the legend is ordered according to the error parameters, and the empirical error generally follows this ordering.
+The two most accurate APA algorithms are the least recently discovered: Bini's $\dims{3,2,2}$ and Schonhage's $\dims{3,3,3;21}$.
+Algorithms that offer more potential speedup tend to be less accurate.
+Two algorithms with smaller error than expected are $\dims{5,5,5}$ and ${7,2,2}$, which is explained by the coefficients of those algorithms including fraction pre-factors.
+While $\phi$ is computed using the exponents of the largest intermediate term of $\lambda$, the leading term of $\dims{5,5,5}$ has a constant of $1/4$, lessening its magnitude. 
+We study the effects of the matrix multiplication error on neural network training and test accuracy in \cref{sec:NNacc}.
 
@@
 \section{Neural Network Accuracy}
+\label{sec:NNacc}
 
@@
 \section{Neural Network Performance}
+\label{sec:NNperf}
 

### 2021-03-03 TF_op/accuracy_test.py
update for accuracy test

@@ -1,0 +1,239 @@
+import tensorflow as tf
+from tensorflow.keras import layers
+from tensorflow.keras import Model
+from tensorflow.python.ops import array_ops, gen_math_ops, math_ops
+import time
+import argparse
+from openmpext import controlOMP
+
+
+# To change OMP num threads
+#print(controlOMP(12))
+
+# to change MKL's threads at runtime
+import ctypes
+mkl_rt = ctypes.CDLL('libmkl_rt.so')
+mkl_set_num_threads = mkl_rt.MKL_Set_Num_Threads
+mkl_get_max_threads = mkl_rt.MKL_Get_Max_Threads
+
+print( "MKL num threads default: ", mkl_get_max_threads() )
+mkl_set_num_threads(1)
+print( "MKL num threads set to: ", mkl_get_max_threads() )
+
+# To change TensorFlow's threads at runtime
+tf.config.threading.set_intra_op_parallelism_threads(1)
+tf.config.threading.set_inter_op_parallelism_threads(1)
+
+
+parser = argparse.ArgumentParser()
+parser.add_argument("--layers", type=int)
+parser.add_argument("--nodes", type=int)
+parser.add_argument("--bs", type=int)
+parser.add_argument("--epochs", type=int)
+parser.add_argument("--mm", type=str)  # name of the matrix multiplication algorithm
+parser.add_argument("--num_threads", type=int)
+
+args = parser.parse_args()
+batch_size = args.bs
+EPOCHS = args.epochs
+nodes = args.nodes
+num_layers = args.layers
+mm_algo = args.mm
+num_threads = args.num_threads
+
+epsilon_values = {
+    'bini322': 2**-11,
+    'schonhage333': 2**-5,
+    'smirnov224': 2**-7,
+    'smirnov225': 2**-5,
+    'smirnov272': 2**-3,
+    'smirnov323': 2**-5,
+    'smirnov333': 2**-3,
+    'smirnov334': 2**-5,
+    'smirnov442': 2**-5,
+    'smirnov444': 2**-5,
+    'smirnov552': 2**-5,
+    'smirnov555': 2**-5
+}
+
+
+@tf.RegisterGradient("FastMatMul")
+def _Fast_MatMul_grad(op, grad):
+    bt = array_ops.transpose(op.inputs[1])
+    at = array_ops.transpose(op.inputs[0])
[…]

### 2021-03-03 TF_op/accuracy_test.sh

@@ -1,0 +1,8 @@
+
+for name in bini322 dgemm schonhage333 smirnov224 smirnov225 smirnov272 smirnov323 smirnov333 smirnov334 smirnov442 smirnov444 smirnov552 smirnov555
+do
+
+
+python tensorboard_test.py --layers 4 --nodes 300 --epochs 50 --bs 300 --mm ${name} --num_threads 12
+
+done

### 2021-03-03 TF_op/accuracy_test.sh
fix typo

@@ -3,6 +3,6 @@ for name in bini322 dgemm schonhage333 smirnov224 smirnov225 smirnov272 smirnov3
 do
 
 
-python tensorboard_test.py --layers 4 --nodes 300 --epochs 50 --bs 300 --mm ${name} --num_threads 12
+python accuracy_test.py --layers 4 --nodes 300 --epochs 50 --bs 300 --mm ${name} --num_threads 12
 
 done

### 2021-03-03 TF_op/accuracy_test.sh
fix typo

@@ -3,6 +3,6 @@ for name in bini322 dgemm schonhage333 smirnov224 smirnov225 smirnov272 smirnov3
 do
 
 
-python accuracy_test.py --layers 4 --nodes 300 --epochs 50 --bs 300 --mm ${name} --num_threads 12
+python accuracy_test.py --layers 2 --nodes 300 --epochs 50 --bs 300 --mm ${name} --num_threads 12
 
 done

### 2021-03-04 TF_op/accuracy_matmul_test.sh
add matmul_accuracy and matmul_time benchmark python scripts

@@ -1,0 +1,7 @@
+for name in dgemm bini322 schonhage333 smirnov224 smirnov225 smirnov272 smirnov323 smirnov333 smirnov334 smirnov442 smirnov444 smirnov552 smirnov555
+do
+
+
+python matmul_accuracy_test.py --mm ${name}
+
+done

### 2021-03-04 TF_op/matmul_accuracy_test.py

@@ -1,0 +1,102 @@
+#!/usr/bin/python
+
+import tensorflow as tf
+import time
+import argparse
+import numpy as np
+#from openmpext import controlOMP
+
+
+parser = argparse.ArgumentParser()
+#parser.add_argument("--tf", type=int)
+#parser.add_argument("--omp", type=int)
+#parser.add_argument("--mkl", type=int)
+parser.add_argument("--mm", type=str)
+
+args = parser.parse_args()
+#ompNum = args.ompNumThread
+
+# to change TensorFlow's threads at runtime
+tf.config.threading.set_intra_op_parallelism_threads(12)
+tf.config.threading.set_inter_op_parallelism_threads(1)
+'''
+#print(controlOMP(args.omp))
+# to change MKL's threads at runtime
+import ctypes
+mkl_rt = ctypes.CDLL('libmkl_rt.so')
+mkl_set_num_threads = mkl_rt.MKL_Set_Num_Threads
+mkl_get_max_threads = mkl_rt.MKL_Get_Max_Threads
+
+mkl_set_num_threads(args.mkl)
+print( mkl_get_max_threads() )
+'''
+algo_name = args.mm
+epsilon_ = 1e-3
+step_ = 1
+fast_mm_module = tf.load_op_library('obj/%s_mat_mul.so'%algo_name)
+
+epsilon_values = {
+    'bini322': -11,
+    'schonhage333': -5,
+    'smirnov224': -7,
+    'smirnov225': -5,
+    'smirnov323': -5,
+    'smirnov334': -5,
+    'smirnov442': -5,
+    'smirnov444': -5,
+    'smirnov552': -5,
+    'smirnov555': -5,
+    'smirnov272': -3,
+    'smirnov333': -3,
+    'dgemm': -3 # placeholder
+}
+
+epsilon_ = epsilon_values.get(algo_name)+2
+
+
+
+
+diff = 0
+#dims = [4096] #[512, 1024, 2048, 4096, 8192]
+err = np.zeros((5, 16))
+k=0
+for dim in range(512, 8192+512, 512):
+    a = tf.Variable(tf.random.uniform(shape=(dim, dim)), dtype=tf.float32)
+    b = tf.Variable(tf.random.uniform(shape=(dim, dim)), dtype=tf.float32)
+    regular = tf.matmul(a, b)
+    loops = 1
+    
+    for j in range(5):
+        diff = 0
+        ep = 2**(epsilon_-j)
+        for i in range(loops):
+
[…]

### 2021-03-04 TF_op/matmul_time_test.py

@@ -1,0 +1,104 @@
+#!/usr/bin/python
+
+import tensorflow as tf
+import time
+import argparse
+import numpy as np
+#from openmpext import controlOMP
+
+
+parser = argparse.ArgumentParser()
+parser.add_argument("--tf", type=int)
+parser.add_argument("--omp", type=int)
+#parser.add_argument("--mkl", type=int)
+parser.add_argument("--mm", type=str)
+
+args = parser.parse_args()
+#ompNum = args.ompNumThread
+
+# to change TensorFlow's threads at runtime
+tf.config.threading.set_intra_op_parallelism_threads(args.tf)
+tf.config.threading.set_inter_op_parallelism_threads(1)
+
+'''
+#print(controlOMP(args.omp))
+# to change MKL's threads at runtime
+import ctypes
+mkl_rt = ctypes.CDLL('libmkl_rt.so')
+mkl_set_num_threads = mkl_rt.MKL_Set_Num_Threads
+mkl_get_max_threads = mkl_rt.MKL_Get_Max_Threads
+
+mkl_set_num_threads(args.mkl)
+print( mkl_get_max_threads() )
+'''
+algo_name = args.mm
+step_ = 1
+fast_mm_module = tf.load_op_library('obj/%s_mat_mul.so'%algo_name)
+
+
+epsilon_values = {
+    'bini322': -11,
+    'schonhage333': -5,
+    'smirnov224': -7,
+    'smirnov225': -5,
+    'smirnov323': -5,
+    'smirnov334': -5,
+    'smirnov442': -5,
+    'smirnov444': -5,
+    'smirnov552': -5,
+    'smirnov555': -5,
+    'smirnov272': -3,
+    'smirnov333': -3,
+    'dgemm': -3 # placeholder
+}
+
+epsilon_ = epsilon_values.get(algo_name)
+
+
+loops = 5
+times = np.zeros((loops, 16))
+k=0
+
+for dim in range(512, 8192+512, 512):
+    
+    '''
+    a = tf.Variable(tf.random.uniform(shape=(dim, dim)), dtype=tf.float32)
+    b = tf.Variable(tf.random.uniform(shape=(dim, dim)), dtype=tf.float32)
+    t3 = time.time()
+    regular = tf.matmul(a, b)
+    t4 = time.time()
+    regular_time = t4-t3
+    '''
+    
+    for i in range(loops):
+
+        a = tf.Variable(tf.random.uniform(shape=(dim, dim)), dtype=tf.float32)
[…]

### 2021-03-04 TF_op/mm_test.sh

@@ -1,0 +1,11 @@
+for sq in 1 12
+do
+
+for name in dgemm bini322 schonhage333 smirnov224 smirnov225 smirnov272 smirnov323 smirnov333 smirnov334 smirnov442 smirnov444 smirnov552 smirnov555
+do
+
+python matmul_time_test.py --tf ${sq} --omp ${sq} --mm ${name}
+
+done
+
+done

### 2021-03-04 TF_op/pull_tensorboard_data.py
add accuracy plots

@@ -19,11 +19,11 @@ if __name__ == '__main__':
                        ]
     matrix_sizes = [512, 1024, 2048, 4096, 8192]
 
-    tensorboard_folders_prefix = 'FINAL_12_thread_'
+    tensorboard_folders_prefix = 'PAPER_1_thread_'
 
-    total_time_log_file_name = "nohup.out"
+    total_time_log_file_name = "oneThreadTimes.log"
 
-    output_file_name = '12_threads_time.csv'
+    output_file_name = '1_thread_time.csv'
 
     output = []
 

### 2021-03-04 TF_op/nn_test.sh
update nn_test

@@ -1,12 +1,15 @@
 
+for threads in 6 12
+do
 for size in 512 1024 2048 4096 8192
 do
 for name in bini322 dgemm classic schonhage333 smirnov224 smirnov225 smirnov272 smirnov323 smirnov333 smirnov334 smirnov442 smirnov444 smirnov552 smirnov555
 do
 
 
-python tensorboard_test.py --layers 4 --nodes ${size} --epochs 5 --bs ${size} --mm ${name} --logdir PAPER_1_thread_${name}${size} --num_threads 1
+python tensorboard_test.py --layers 4 --nodes ${size} --epochs 5 --bs ${size} --mm ${name} --logdir PAPER_${threads}_threads_${name}${size} --num_threads ${threads}
 
+done
 done
 done
 

### 2021-03-04 TF_op/tensorboard_test.py

@@ -125,16 +125,18 @@ def train_step(images, labels):
 
 
 
-@tf.function
-def test_step(images, labels):
-  predictions = model(images)
-  t_loss = loss_object(labels, predictions)
-
-  #test_loss(t_loss)
-  #test_accuracy(labels, predictions)
+# @tf.function
+# def test_step(images, labels):
+#   predictions = model(images)
+#   t_loss = loss_object(labels, predictions)
+#
+#   #test_loss(t_loss)
+#   #test_accuracy(labels, predictions)
 
 
 if __name__ == '__main__':
+
+    start = time.time()
 
     if mm_algo != 'regular':
         fast_mm_module = tf.load_op_library(f'obj/{mm_algo}_mat_mul.so')
@@ -157,8 +159,8 @@ if __name__ == '__main__':
     train_loss = tf.keras.metrics.Mean(name='train_loss')
     train_accuracy = tf.keras.metrics.SparseCategoricalAccuracy(name='train_accuracy')
 
-    test_loss = tf.keras.metrics.Mean(name='test_loss')
-    test_accuracy = tf.keras.metrics.SparseCategoricalAccuracy(name='test_accuracy')
+    #test_loss = tf.keras.metrics.Mean(name='test_loss')
+    #test_accuracy = tf.keras.metrics.SparseCategoricalAccuracy(name='test_accuracy')
 
     train_accuracy_list = []
     train_loss_list = []
@@ -175,8 +177,8 @@ if __name__ == '__main__':
         # Reset the metrics at the start of the next epoch
         train_loss.reset_states()
         train_accuracy.reset_states()
-        test_loss.reset_states()
-        test_accuracy.reset_states()
+        #test_loss.reset_states()
+        #test_accuracy.reset_states()
 
         total_batch_time = 0
         batches = 0
@@ -191,15 +193,15 @@ if __name__ == '__main__':
             total_batch_time += y - x
             batches += 1
 
-        test_step(x_test, y_test)
+        #test_step(x_test, y_test)
         b = time.time()
-        print(f'Epoch {epoch + 1}, Loss: {train_loss.result()}, Train Accuracy: {train_accuracy.result() * 100},'
-              f'Test Loss: {test_loss.result()}, Test Accuracy: {test_accuracy.result() * 100}')
-
-        train_accuracy_list.append(train_accuracy.result())
-        train_loss_list.append(train_loss.result())
-        test_accuracy_list.append(test_accuracy.result())
-        test_loss_list.append(test_loss.result())
[…]

### 2021-03-04 TF_op/pull_tensorboard_data.py
add NN_acc

@@ -21,7 +21,7 @@ if __name__ == '__main__':
 
     tensorboard_folders_prefix = 'PAPER_1_thread_'
 
-    total_time_log_file_name = "oneThreadTimes.log"
+    total_time_log_file_name = "oneTheadTimes.log"
 
     output_file_name = '1_thread_time.csv'
 

### 2021-03-04 latex/MPP21/paper.tex
updated performance plots, now displaying performance data
added 6-thread matmul data to paper, renamed a couple data files
added NN accuracy plots to paper
drafted analysis of matmul performance

@@
 \crefname{figure}{Fig.}{Figs.}
+\Crefname{figure}{Figure}{Figures}
 \crefname{table}{Table}{Tables}
@@
 \newcommand{\threads}{}
+\newcommand{\set}{}
 
@@
 \textit{Wake Forest University}\\
-Winston Salem, NC, USA \\
+Winston-Salem, NC, USA \\
 \email{ballard@wfu.edu}}
@@
 \textit{Wake Forest University}\\
-Winston Salem, NC, USA \\
+Winston-Salem, NC, USA \\
 \email{jack.weissenberger@gmail.com}}
@@
 \textit{Wake Forest University}\\
-Winston Salem, NC, USA \\
+Winston-Salem, NC, USA \\
 \email{zhanl317@wfu.edu}}
@@
 \centering
-\renewcommand{\datafile}{data/accuracy.dat}
+\renewcommand{\datafile}{data/matmul_acc.dat}
 \input{plots/accuracy}
@@
 Because we use only 1 recursive step in our experiments, there are no recursive calls, and every multiplication is performed by \texttt{gemm}.
-We note that the original code generation tool was designed primarily for exact algorithms, and we generalized it to apply to all of the APA algorithms we consider here.
+We note that the original code generation tool was designed primarily for exact algorithms, and we generalized it to apply to all of the APA algorithms we consider here, some of which have more complicated coefficients than previously considered.
 
@@
 
+\subsection{Experimental Platform}
+
+\GB{add paragraph describing householder, listing versions of MKL, peak GFLOPS}
+
 \subsection{Sequential Performance}
 
-\GB{add sequential performance results for all 12 algorithms, compared to \texttt{gemm}}
+\Cref{fig:matmul_seq} reports the sequential performance of the APA algorithms in comparison to the most efficient implementation of the classical algorithm, MKL's \texttt{sgemm}.
+The y-axis is the \emph{effective} GFLOPS, which is measured as $\num{1e-9} \cdot 2n^3 / \text{ time}$.
+That is, the GFLOPS reported for APA algorithms is not true performance, as they perform fewer flops than the classical algorithm.
+We use this metric to be able to compare relative times across algorithms performing different amounts of computation.
+The machine peak for a classical algorithm is given by the horizontal dotted line.
+
+We vary the dimension from 512 up to 8192 and see that all algorithms outperform classical for dimensions larger than 2000 or so.
+The highest performing algorithm is $\dims{4,4,4}$, and at dimension 8192, it is 28\% faster than \texttt{sgemm}.
+Ignoring the cost of the matrix additions, $\dims{4,4,4}$ performs 39\% fewer flops than the classical algorithm; the drop to 28\% achieved improvement is because of the overhead of matrix additions and the reduced performance of \texttt{sgemm} on smaller matrices.
+We note that $\dims{4,4,2}$ is also high performing, achieving a 25\% observed improvement (out of a theoretical 33\%), along with $\dims{3,3,3;20}$ and $\dims{5,5,5}$.
 
@@
 
-\GB{add parallel performance results for all 12 algorithms, compared to \texttt{gemm}}
+We report parallel performance in \cref{fig:matmul_par_6,fig:matmul_par_12} for running with 6 threads (one socket) and 12 threads (both sockets).
+Overall, speedups of APA algorithms over classical are reduced in the parallel case.
+Again, the theoretical speedup is based on a reduction in the multiplications, and the overhead of additions is the biggest impediment to realizing that speedup.
+In the parallel case, the additions can become an even larger bottleneck because the additions are memory bandwidth bound, and the memory bandwidth does not scale with the number of cores \cite{BB15}.
+While we can expect close to linear speedup with cores on the multiplications (which are perfectly load balanced), linear scaling is impossible to achieve with the additions.
+We also note that when the hybrid method uses all threads on remainder multiplications, the smaller dimensions make it harder for \texttt{sgemm} to scale as well as it can on the original matrix dimensions.
+
+In the performance results for 6 threads (\cref{fig:matmul_par_6}), we see that many of the algorithms start to outperform classical around dimension 2000, though some poor performance is observed for algorithms and particular matrix dimensions.
+The fastest algorithms (e.g., $\dims{4,4,4}$ and $\dims{4,4,2}$) achieve a speedup of up to 23\% over \texttt{gemm} and exceed the machine peak for classical algorithms.
+We note that the $\dims{4,4,2}$ algorithm has 24 subproblems, which is a multiple of 6, so there are no remainder subproblems that require all threads.
+
+\Cref{fig:matmul_par_12} shows the results for 12 threads.
+Here we see a majority of the algorithms are slower than the classical algorithm, even for large matrices.
+We attribute the poor performance to the effect of lower \texttt{sgemm} performance for smaller matrices, as well as a lack of NUMA-aware optimization.
+The ``ramp-up'' range of \texttt{sgemm} performance is much shallower for 12 threads than for 6 threads, not achieving the plateau performance until dimension 4000 or so.
+This implies that the dimension of the submultiplications does not fall on the plateau for any APA algorithm, so the remainder multiplications suffer from poor parallel performance.
+The algorithm with no remainder multiplications, $\dims{4,2,2}$, does not suffer this problem and maintains higher performance.
+It exceeds \texttt{sgemm}'s effective performance at dimension 4000, exceeds the peak parallel performance of any classical algorithm for larger dimensions, and achieves a 21\% speedup over \texttt{sgemm} at dimension 8192 for an effective rate of 389 GFLOPS.
 
@@
 \centering
-\renewcommand{\datafile}{data/matmul_par.dat}
+\renewcommand{\datafile}{data/matmul_par_6.dat}
+\renewcommand{\threads}{6}
+\input{plots/matmul_perf}
+\caption{Multithreaded matrix multiplication performance (6 threads)}
+\label{fig:matmul_par_6}
+\end{figure}
+
+\begin{figure}
+\centering
+\renewcommand{\datafile}{data/matmul_par_12.dat}
 \renewcommand{\threads}{12}
 \input{plots/matmul_perf}
-\caption{Multithreaded matrix multiplication performance}
-\label{fig:matmul_seq}
+\caption{Multithreaded matrix multiplication performance (12 threads)}
+\label{fig:matmul_par_12}
 \end{figure}
@@
 \begin{enumerate}
-	\item describe network (MNIST?) and implementation details
+	\item describe network (MLP for MNIST) and implementation details (including TensorFlow with version number)
 	\item accuracy results in table or fig
 \end{enumerate}
+
+\begin{figure}
+\centering
+\renewcommand{\datafile}{data/NN_acc.dat}
+\renewcommand{\set}{train}
+\input{plots/NN_acc}
+\caption{MLP training accuracy for MNIST data set}
+\label{fig:matmul_seq}
+\end{figure}
+
+\begin{figure}
+\centering
+\renewcommand{\datafile}{data/NN_acc.dat}
+\renewcommand{\set}{test}
+\input{plots/NN_acc}
+\caption{MLP test accuracy for MNIST data set}
+\label{fig:matmul_seq}
+\end{figure}
 

### 2021-03-05 latex/MPP21/paper.tex
add draft of network accuracy

@@
 
-\begin{enumerate}
-	\item describe network (MLP for MNIST) and implementation details (including TensorFlow with version number)
-	\item accuracy results in table or fig
-\end{enumerate}
+DRAFT: To measure the effect of the error introduced from the APA algorithms, we trained a simple feed forward network on the MNIST dataset and measured its performance over the course of training.
+MNIST is a common machine learning dataset of 70,000 images of hand-written numerical digits.
+Each of the images are composed of 28 by 28 grayscale pixels.
+To feed the images into the network we flatten each of the images into a 784 dimensional vector.
+The dataset was split into 60,000 examples for the training set and 10,000 examples for the test set.
+The network is composed of 4 fully-connected layers of shapes 784, 300, 300, and 10.
+The input layer takes in the values of all of the pixels which is why it 784 nodes and the output layer is 10 nodes, one for each digit 0-9.
+The network was created using Tensorflow version 2.2.0.
+APA algorithms were used for the middle multiplication of the network (300x300).
+The input and output layers used the classical matrix multiplication algorithm.
+One network was trained for 50 epochs for each of the APA algorithms and one was trained using the classical matrix multiplication algorithm.
+Figure 6 plots the accuracy on the training data over each epoch and figure 7 plots the test accuracy over each epoch.
+The error introduced from these algorithms did not have much of an effect on these models and they were able to reach the same network performance as the classical algorithm.
 

### 2021-03-05 latex/MPP21/paper.tex
set up NN overall reltime plots
added description of householder, inserted speedup numbers in abstract and intro

@@
 Practical APA algorithms provide significant reduction in computation time and still provide enough accuracy for many applications like neural network training.
-We demonstrate that APA algorithms can be efficiently implemented and parallelized for shared-memory platforms to obtain up to \GB{Z\%} speedups over the fastest implementation of the classical algorithm.
+We demonstrate that APA algorithms can be efficiently implemented and parallelized for shared-memory platforms to obtain up to 28\% and 21\% speedups over the fastest implementation of the classical algorithm using one core and 12 cores, respectively.
 Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no signification reduction in the training or testing error.
@@
 	\item curate a collection of well-established and recently discovered practical APA algorithms;
-	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them;
+	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them, achieving up to 21\% performance improvement over the best parallel classical implementation,
 	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications; and
@@
 \end{enumerate}
-
-\GB{\cite{NPOV15} uses a network including layers with dimensions 25K x 4K and 4K x 4K}
 
@@
 A key characteristic of matrix multiplication, first demonstrated by Bini et al.~\cite{BCRL79}, is that it can be approximated to arbitrary accuracy more efficiently than being computed exactly.
-An APA algorithm is one that takes as input $A$, $B$, and a parameter $0<\lambda<1$ and computes 
+An APA algorithm is one that takes as input $A$, $B$, and a scalar parameter $0<\lambda<1$ and computes 
 \begin{equation}
@@
 For a concrete example, we reproduce the rule for the algorithm developed by Bini et al.~\cite{BCRL79} for the $\dims{3,2,2}$ case (multiplying a $3\times 2$ matrix $A$ by a $2\times 2$ matrix $B$), where we use the following notation for input and output matrices:
-$$\begin{bmatrix} C_{11} & C_{12} \\ C_{21} & C_{22} \\ C_{31} & C_{32} \end{bmatrix} = 
+$$
 \begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \\ A_{31} & A_{32}  \end{bmatrix} \cdot
-\begin{bmatrix} B_{11} & B_{12} \\ B_{21} & B_{22} \end{bmatrix}.$$
+\begin{bmatrix} B_{11} & B_{12} \\ B_{21} & B_{22} \end{bmatrix}=
+\begin{bmatrix} C_{11} & C_{12} \\ C_{21} & C_{22} \\ C_{31} & C_{32} \end{bmatrix} .$$
 Bini's rule is given by
@@
 
+\subsection{Experimental Platform}
+
+All experiments are performed on a dual-socket Intel Xeon E5-2620 (Sandy Bridge) server with 2 sockets each with 6 cores. 
+Each socket has a 15 MB L3 cache, and each core has a 256 KB L2 cache and 32 KB L1 data cache. 
+Each core has a clock rate of 2.00 GHz (with turbo boost disabled) and peak single precision flop rate of 32 GFLOPS. 
+Our code is compiled with GCC version 7.5.0, 
+and we use Intel's Math Kernel Library (MKL) version 2019.4.243.
+
 \subsection{Hybrid Parallelization Strategy}
@@
 The hybrid strategy also perfectly load balances the computation across threads, as opposed to the alternate strategy of assigning the $\ell$ remainder multiplications to $\ell$ different threads (known as ``BFS''), leaving the other $p-\ell$ threads idle.
-
-\subsection{Experimental Platform}
-
-\GB{add paragraph describing householder, listing versions of MKL, peak GFLOPS}
 
@@
 \centering
-\renewcommand{\datafile}{data/NNtime_1t.dat}
-\input{plots/reltime}
+\renewcommand{\datafile}{data/NN_seq.dat}
+\input{plots/NN_reltime}
 \label{fig:NNreltime_seq}}
@@
 \renewcommand{\datafile}{data/NNtime_12t.dat}
-\input{plots/reltime}
+%\input{plots/NN_reltime}
 \label{fig:NNreltime_par}}
@@
 \begin{enumerate}
-	\item promising results for synthetic network, goal is to target costly networks bottlenecked by matmul
+	\item promising results for synthetic network, goal is to target costly networks bottlenecked by matmul (\GB{\cite{NPOV15} uses a network including layers with dimensions 25K x 4K and 4K x 4K})
 	\item can tune for matmul dimensions (not just square)

### 2021-03-05 latex/MPP21/paper.tex
made editing pass over section IV (NN accuracy)

@@
 We demonstrate that APA algorithms can be efficiently implemented and parallelized for shared-memory platforms to obtain up to 28\% and 21\% speedups over the fastest implementation of the classical algorithm using one core and 12 cores, respectively.
-Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no signification reduction in the training or testing error.
-Our performance results on a large MLP network show overall performance improvements of up to \GB{W\%}.
+Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no significant reduction in the training or testing error.
+Our performance results on a large MLP network show overall multithreaded performance improvements of up to \GB{W\%}.
 \end{abstract}
@@
 
-DRAFT: To measure the effect of the error introduced from the APA algorithms, we trained a simple feed forward network on the MNIST dataset and measured its performance over the course of training.
+To measure the effect of the matrix multiplication error introduced from the APA algorithms on neural network accuracy, we trained a multi-layer perceptron (MLP) network on the MNIST dataset \cite{LBBH98} and measured its accuracy over the course of training.
+The network was created using Tensorflow version 2.2.0.
+
 MNIST is a common machine learning dataset of 70,000 images of hand-written numerical digits. 
-Each of the images are composed of 28 by 28 grayscale pixels.
-To feed the images into the network we flatten each of the images into a 784 dimensional vector.
-The dataset was split into 60,000 examples for the training set and 10,000 examples for the test set.
-The network is composed of 4 fully-connected layers of shapes 784, 300, 300, and 10.
-The input layer takes in the values of all of the pixels which is why it 784 nodes and the output layer is 10 nodes, one for each digit 0-9.
-The network was created using Tensorflow version 2.2.0.
-APA algorithms were used for the middle multiplication of the network (300x300).
-The input and output layers used the classical matrix multiplication algorithm.
+Each of the images are composed of 28 by 28 grayscale pixels, which are flattened into a vector of length 784.
+The dataset is split into 60,000 examples for the training set and 10,000 examples for the test set.
+
+The MLP network is composed of 4 fully-connected layers with 784, 300, 300, and 10 nodes, respectively.  
+The output layer has one node for each digit 0-9. 
+Training is performed using batched stochastic gradient descent with a batch size of \GB{ZZZ}.
+APA algorithms were used for the middle multiplication of the network (yielding matrix multiplication dimenions $300\times \GB{ZZZ}\times300$), while the input and output layers use the classical matrix multiplication algorithm.
 One network was trained for 50 epochs for each of the APA algorithms and one was trained using the classical matrix multiplication algorithm.
-Figure 6 plots the accuracy on the training data over each epoch and figure 7 plots the test accuracy over each epoch.
-The error introduced from these algorithms did not have much of an effect on these models and they were able to reach the same network performance as the classical algorithm.
+
+\Cref{fig:NN_acc_train} plots the accuracy on the training data over each epoch, and \cref{fig:NN_acc_test} shows the test accuracy over epochs. 
+The matrix multiplication error introduced from these algorithms had minimal effect on the training error; all algorithms yielded convergence to nearly full accuracy after 20 epochs or so.
+We see more variability in the test error due to the sensitivity inherent in generalizability of the model, though all algorithms achieve between 97\% and 99\% test accuracy.
 
@@
 \caption{MLP training accuracy for MNIST data set}
-\label{fig:matmul_seq}
+\label{fig:NN_acc_train}
 \end{figure}
@@
 \caption{MLP test accuracy for MNIST data set}
-\label{fig:matmul_seq}
+\label{fig:NN_acc_test}
 \end{figure}

### 2021-03-05 latex/MPP21/paper.tex
drafted conclusion section

@@
 \\
+\subfloat[Six threads]{
+\centering
+\renewcommand{\datafile}{data/NN_par_6.dat}
+%\input{plots/NN_reltime}
+\label{fig:NNreltime_par_6}}
+\\
 \subfloat[Twelve threads]{
 \centering
-\renewcommand{\datafile}{data/NNtime_12t.dat}
+\renewcommand{\datafile}{data/NN_par_12.dat}
 %\input{plots/NN_reltime}
-\label{fig:NNreltime_par}}
+\label{fig:NNreltime_par_12}}
 \caption{Network training time relative to using classical matrix multiplication}
@@
 
-\begin{enumerate}
-	\item promising results for synthetic network, goal is to target costly networks bottlenecked by matmul (\GB{\cite{NPOV15} uses a network including layers with dimensions 25K x 4K and 4K x 4K})
-	\item can tune for matmul dimensions (not just square)
-	\item GPU implementation (cite Jianyu Huang's work)
-	\item mixed precision fast matmul (APA or otherwise)
-\end{enumerate}
+Arbitrary Precision Approximating algorithms offer significant practical performance improvements over exact fast matrix multiplication algorithms.
+They outperform classical matrix multiplication at smaller matrix dimensions and yield greater benefits as matrix dimensions increase.
+APA algorithms sacrifice accuracy, achieving error of only a fractional root of the working precision, but they can be practical for applications that do not demand high accuracy.
+The results presented here show that neural network training, particularly for Multi-Layer Perceptron networks, is robust to such matrix multiplication error, and APA algorithms are a useful tool for accelerating the costly training computations.
+We observe that for APA algorithms offering a theoretical speedup of up to 39\%, we can achieve up to 28\% and 21\% improvements for single-threaded and multi-threaded implementations of individual matrix multiplications, respectively, which translate to up to \GB{W\%} speedup in overall training time on a multicore CPU.
+
+While the accuracy and performance results are promising for MLP networks and the MNIST dataset, we believe that APA algorithms can also be effective in state-of-the-art networks and more difficult machine learning problems.
+For example, the VGG-19 network for large-scale visual recognition \cite{SZ15} is based on a deep convolutional network and is bottlenecked by large fully connected layers, including a layer with dimensions $25088\times 4096$, where APA algorithms can offer significant performance improvement.
+
+As illustrated by Benson and Ballard \cite{BB15}, the highest performing fast algorithms for rectangular matrix multiplications often have dimensions that match the aspect ratio of the problem.
+We consider in this paper only square matrix multiplications, and we see that square (or nearly square) dimensions of $\dims{4,4,4}$ and $\dims{4,4,2}$ are fastest.
+Algorithms with more skewed aspect ratios will likely perform better for problems with matching skewed matrix multiplication dimensions.
+We note that an algorithm for dimensions $\dims{m,n,k}$ can be translated into an algorithm for $\dims{n,m,k}$ and any other reordering of the dimensions \cite{BB15}.
+For very large matrices, it may also be beneficial to consider more than one recursive step of the same algorithm or to use a combination of two or three different algorithms across recursive steps (i.e., uniform, non-stationary algorithms \cite{BBDLS16}).
+
+Finally, we would like to extend the code generation techniques to other hardware platforms, including GPUs.
+Because GPUs can offer higher efficiency for matrix multiplications, we wish to demonstrate the potential of APA algorithms in that environment.
+Techniques used to accelerate Strassen's algorithm on the GPU \cite{HYvdG20,LAES13} can be generalized to other fast algorithms as done in work for multicore CPUs \cite{BB15,HRMvdG17}.
 

### 2021-03-05 latex/MPP21/paper.tex
cleaning up bib and adding batch size for NN acc section

@@
 The output layer has one node for each digit 0-9. 
-Training is performed using batched stochastic gradient descent with a batch size of \GB{ZZZ}.
-APA algorithms were used for the middle multiplication of the network (yielding matrix multiplication dimenions $300\times \GB{ZZZ}\times300$), while the input and output layers use the classical matrix multiplication algorithm.
+Training is performed using batched stochastic gradient descent with a batch size of 300.
+APA algorithms were used for the middle multiplication of the network (yielding matrix multiplication dimensions $300\times 300\times300$), while the input and output layers use the classical matrix multiplication algorithm.
 One network was trained for 50 epochs for each of the APA algorithms and one was trained using the classical matrix multiplication algorithm.

### 2021-03-06 latex/MPP21/paper.tex
re-formatted plots to group appropriately
moving figs and tables around
setting up NN plots
Add neural network figure
add draft of section 5 to paper

@@
 
+\begin{table}
+\centering
+\caption{Properties of Arbitrary Precision Approximating Algorithms  \\ Speedup and Error are computed assuming 1 recursive step}
+\label{tab:algs}
+\begin{tabular}{| c | c c c | c c c |} 
+\hline
+\textbf{Ref} & \textbf{Dims} & \textbf{Rank} & \textbf{Speedup} & $\mathbf{\sigma}$ & $\mathbf{\varphi}$ & \textbf{Error} \\
+\hline
+- & $\dims{2,2,2}$ & 8 & - & 1 & 0 & \num{1.2e-7} \\
+\hline
+\cite{BCRL79} & $\dims{3,2,2}$ & 10 & $20\%$ & 1 & 1 & \num{3.5e-4} \\
+\cite{AS13} & $\dims{4,2,2}$ & 13 & $23\%$ & 1 & 2 & \num{4.9e-3} \\
+\cite{Smirnov13} & $\dims{3,3,2}$ & 14 & $29\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov13} & $\dims{5,2,2}$ & 16 & $25\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov13} & $\dims{3,3,3}$ & 20 & $35\%$ & 1 & 6 & \num{1.0e-1} \\
+\cite{Schonhage81} & $\dims{3,3,3}$ & 21 & $29\%$ & 1 & 2 & \num{4.9e-3} \\
+\cite{Smirnov15} & $\dims{7,2,2}$ & 22 & $27\%$ & 1 & 5 & \num{7.0e-2} \\
+\cite{Smirnov16b} & $\dims{4,4,2}$ & 24 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov16a} & $\dims{4,3,3}$ & 27 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov16b} & $\dims{5,5,2}$ & 37 & $35\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov14} & $\dims{4,4,4}$ & 46 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
+\cite{Smirnov18} & $\dims{5,5,5}$ & 90 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
+\hline
+\end{tabular}
+\end{table}
+
 \subsection{Numerical Error of APA Algorithms}
@@
 
-\begin{table}
-\centering
-\caption{Properties of Arbitrary Precision Approximating Algorithms  \\ Speedup and Error are computed assuming 1 recursive step}
-\label{tab:algs}
-\begin{tabular}{| c | c c c | c c c |} 
-\hline
-\textbf{Ref} & \textbf{Dims} & \textbf{Rank} & \textbf{Speedup} & $\mathbf{\sigma}$ & $\mathbf{\varphi}$ & \textbf{Error} \\
-\hline
-- & $\dims{2,2,2}$ & 8 & - & 1 & 0 & \num{1.2e-7} \\
-\hline
-\cite{BCRL79} & $\dims{3,2,2}$ & 10 & $20\%$ & 1 & 1 & \num{3.5e-4} \\
-\cite{AS13} & $\dims{4,2,2}$ & 13 & $23\%$ & 1 & 2 & \num{4.9e-3} \\
-\cite{Smirnov13} & $\dims{3,3,2}$ & 14 & $29\%$ & 1 & 3 & \num{1.9e-2} \\
-\cite{Smirnov13} & $\dims{5,2,2}$ & 16 & $25\%$ & 1 & 3 & \num{1.9e-2} \\
-\cite{Smirnov13} & $\dims{3,3,3}$ & 20 & $35\%$ & 1 & 6 & \num{1.0e-1} \\
-\cite{Schonhage81} & $\dims{3,3,3}$ & 21 & $29\%$ & 1 & 2 & \num{4.9e-3} \\
-\cite{Smirnov15} & $\dims{7,2,2}$ & 22 & $27\%$ & 1 & 5 & \num{7.0e-2} \\
-\cite{Smirnov16b} & $\dims{4,4,2}$ & 24 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
-\cite{Smirnov16a} & $\dims{4,3,3}$ & 27 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
-\cite{Smirnov16b} & $\dims{5,5,2}$ & 37 & $35\%$ & 1 & 3 & \num{1.9e-2} \\
-\cite{Smirnov14} & $\dims{4,4,4}$ & 46 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
-\cite{Smirnov18} & $\dims{5,5,5}$ & 90 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
-\hline
-\end{tabular}
-\end{table}
-
 \section{Parallel Fast Matrix Multiplication}
@@
 
+\begin{figure*}[t!]
+%
+\subfloat[One thread]{
+\centering
+\renewcommand{\datafile}{data/matmul_seq.dat}
+\renewcommand{\threads}{1}
+\input{plots/matmul_perf}
+%\caption{Single-threaded matrix multiplication performance}
+\label{fig:matmul_seq}}
+%
+\hfill
+\subfloat[Six threads]{
+\centering
+\renewcommand{\datafile}{data/matmul_par_6.dat}
+\renewcommand{\threads}{6}
+\input{plots/matmul_perf}
+%\caption{Multithreaded matrix multiplication performance (6 threads)}
+\label{fig:matmul_par_6}}
+%
+\hfill
+\subfloat[Twelve threads]{
+\centering
+\renewcommand{\datafile}{data/matmul_par_12.dat}
+\renewcommand{\threads}{12}
+\input{plots/matmul_perf}
+%\caption{Multithreaded matrix multiplication performance (6 threads)}
+\label{fig:matmul_par_12}}
+%
+\caption{Single- and multi-threaded square matrix multiplication performance, relative to $2n^3$ operations}
+\label{fig:matmul_perf}
+\end{figure*}
+
 \subsection{Sequential Performance}
@@
 We note that $\dims{4,4,2}$ is also high performing, achieving a 25\% observed improvement (out of a theoretical 33\%), along with $\dims{3,3,3;20}$ and $\dims{5,5,5}$.
-
-\begin{figure}
-\centering
-\renewcommand{\datafile}{data/matmul_seq.dat}
-\renewcommand{\threads}{1}
-\input{plots/matmul_perf}
-\caption{Single-threaded matrix multiplication performance}
-\label{fig:matmul_seq}
-\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\renewcommand{\datafile}{data/matmul_par_6.dat}
-\renewcommand{\threads}{6}
-\input{plots/matmul_perf}
-\caption{Multithreaded matrix multiplication performance (6 threads)}
-\label{fig:matmul_par_6}
-\end{figure}
-
-\begin{figure}
-\centering
-\renewcommand{\datafile}{data/matmul_par_12.dat}
-\renewcommand{\threads}{12}
-\input{plots/matmul_perf}
-\caption{Multithreaded matrix multiplication performance (12 threads)}
-\label{fig:matmul_par_12}
-\end{figure}
-
 \section{Neural Network Accuracy}
 \label{sec:NNacc}
+
+\begin{figure*}[t!]
+%
+\subfloat[Training accuracy]{
+\centering
+\renewcommand{\datafile}{data/NN_acc.dat}
+\renewcommand{\set}{train}
+\input{plots/NN_acc}
+\label{fig:NN_acc_train}}
+%
+\hfill
+\subfloat[Test accuracy]{
+\centering
+\renewcommand{\datafile}{data/NN_acc.dat}
+\renewcommand{\set}{test}
+\input{plots/NN_acc}
+\label{fig:NN_acc_test}}
+%
+\caption{Multi-Layer Perceptron network accuracy for MNIST data set}
+\label{fig:NN_acc}
+\end{figure*}
 
@@
 APA algorithms were used for the middle multiplication of the network (yielding matrix multiplication dimensions $300\times 300\times300$), while the input and output layers use the classical matrix multiplication algorithm.
+The APA algorithms were also used during back propagation in the corresponding multiplications in the center of the network.
 One network was trained for 50 epochs for each of the APA algorithms and one was trained using the classical matrix multiplication algorithm.
@@
 
-\begin{figure}
+\section{Neural Network Performance}
+\label{sec:NNperf}
+
+The neural networks used for performance benchmarking were 6 layer (4 hidden layers), fully connected networks and were trained on the MNIST dataset.
+The purpose of these experiments was to measure the speed up in training time provided by the APA algorithms and not to measure the accuracy performance of the networks.
+The APA algorithms were used only in the multiplications in the hidden layers during forwards and backward propagation.
+
+Similar to other benchmarks [https://github.com/Emma926/paradnn] on fully connected networks, we measured the performance of each algorithm with the number of nodes in each hidden layer from 2^9 to 20^13.
+We also matched the batch size of each network to the number of nodes in the hidden layer which produces square multiplications in the hidden layers.
+We ran these tests using 1, 6 and 12 threads.
+
+The single threaded results began to show speed up over the classical algorithm when the matrix dimensions were 2^10 and were able to show over a 20\% speed up when the dimensions were 2^13.
+It was more difficult to get performance improvements with more threads but both 6 and 12 thread results were able to show improvement in the larger dimensions.
+At 6 threads, the best algorithms showed small improvements when the dimensions were 2^11 and were able to get over a 10\% improvement at 2^13.
+The 12 thread algorithms were not able to see an improvement until the dimensions were 2^12 and were not able to get 10\% improvement at 2^13.
+
+\begin{figure*}[t!]
+%
+\subfloat[One thread]{
 \centering
-\renewcommand{\datafile}{data/NN_acc.dat}
-\renewcommand{\set}{train}
-\input{plots/NN_acc}
-\caption{MLP training accuracy for MNIST data set}
-\label{fig:NN_acc_train}
-\end{figure}
+\renewcommand{\datafile}{data/NN_seq.dat}
+\renewcommand{\threads}{1}
+\input{plots/NN_reltime}
+\label{fig:NNreltime_seq}}
+%
+\hfill
+\subfloat[Six threads]{
+\centering
+\renewcommand{\datafile}{data/NN_par_6.dat}
+\renewcommand{\threads}{6}
+\input{plots/NN_reltime}
+\label{fig:NNreltime_par_6}}
+%
+\hfill
+\subfloat[Twelve threads]{
+\centering
+\renewcommand{\datafile}{data/NN_par_12.dat}
+\renewcommand{\threads}{12}
+\input{plots/NN_reltime}
+\label{fig:NNreltime_par_12}}
+%
+\caption{Network training time relative to using classical matrix multiplication}
+\label{fig:NNreltime}
+\end{figure*}
 
@@
 	\centering
-\renewcommand{\datafile}{data/NN_acc.dat}
-\renewcommand{\set}{test}
-\input{plots/NN_acc}
-\caption{MLP test accuracy for MNIST data set}
-\label{fig:NN_acc_test}
-\end{figure}
-
-\section{Neural Network Performance}
-\label{sec:NNperf}
-
-\begin{enumerate}
-	\item describe synthetic network and implementation details
-	\item perf results for growing dim (maybe for serial and 12-threads separately)
-	\item parallel speedup results
-\end{enumerate}
-
-\begin{figure}
-\subfloat[One thread]{
-\centering
-\renewcommand{\datafile}{data/NN_seq.dat}
-\input{plots/NN_reltime}
-\label{fig:NNreltime_seq}}
-\\
-\subfloat[Six threads]{
-\centering
-\renewcommand{\datafile}{data/NN_par_6.dat}
-%\input{plots/NN_reltime}
-\label{fig:NNreltime_par_6}}
-\\
-\subfloat[Twelve threads]{
-\centering
-\renewcommand{\datafile}{data/NN_par_12.dat}
-%\input{plots/NN_reltime}
-\label{fig:NNreltime_par_12}}
-\caption{Network training time relative to using classical matrix multiplication}
-\label{fig:NNreltime}
+	\input{figs/neuralnet}
+	\caption{MNIST Neural Networks}
+	\label{fig:neuralnet}
 	\end{figure}

### 2021-03-06 latex/MPP21/paper.tex
reformatted a lot and made editing pass over new text, full draft complete, uploaded to easychair

--- 2021-03-05 latex/MPP21/paper.tex
@@
 \usepackage{ifthen}
+\usepackage{flushend}
 
@@
 Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no significant reduction in the training or testing error.
-Our performance results on a large MLP network show overall multithreaded performance improvements of up to \GB{W\%}.
+Our performance results on a large MLP network show overall multithreaded performance improvements of up to 13\%.
 \end{abstract}
@@
 	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications; and
-	\item present multithreaded performance improvements of a synthetic MLP of up to \GB{$Y\times$} over the use of classical matrix multiplication.
+	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication.
 \end{enumerate}
@@
 
-\begin{figure}
-\centering
-\renewcommand{\datafile}{data/matmul_seq.dat}
-\renewcommand{\threads}{1}
-\input{plots/matmul_perf}
-\caption{Single-threaded matrix multiplication performance}
-\label{fig:matmul_seq}
-\end{figure}
-
 \subsection{Parallel Performance}
+\label{sec:matmul_perf_par}
 
@@
 
-\begin{figure}
+\section{Application of APA Algorithms in Multi-Layer Perceptron Network Training}
+
+\begin{figure}[b]
 \centering
-\renewcommand{\datafile}{data/matmul_par_6.dat}
-\renewcommand{\threads}{6}
-\input{plots/matmul_perf}
-\caption{Multithreaded matrix multiplication performance (6 threads)}
-\label{fig:matmul_par_6}
+\input{figs/neuralnet}
+\caption{Multi-Layer Perceptron network structure}
+\label{fig:neuralnet}
 \end{figure}
 
-\begin{figure}
+\begin{figure*}[ht!]
+%
+\subfloat[Training accuracy]{
 \centering
-\renewcommand{\datafile}{data/matmul_par_12.dat}
-\renewcommand{\threads}{12}
-\input{plots/matmul_perf}
-\caption{Multithreaded matrix multiplication performance (12 threads)}
-\label{fig:matmul_par_12}
-\end{figure}
+\renewcommand{\datafile}{data/NN_acc.dat}
+\renewcommand{\set}{train}
+\input{plots/NN_acc}
+\label{fig:NN_acc_train}}
+%
+\hfill
+\subfloat[Test accuracy]{
+\centering
+\renewcommand{\datafile}{data/NN_acc.dat}
+\renewcommand{\set}{test}
+\input{plots/NN_acc}
+\label{fig:NN_acc_test}}
+%
+\caption{MLP network accuracy for MNIST data set}
+\label{fig:NN_acc}
+\end{figure*}
 
-\section{Neural Network Accuracy}
+\subsection{MLP Network Structure}
+
+To illustrate the effectiveness of APA algorithms in neural network training, we consider Multi-Layer Perceptron networks, as shown in \cref{fig:neuralnet}.
+We use TensorFlow 2.2.0 to build the networks and implement custom operators for matrix multiplication.
+For internal layers, we use the custom operator for both forward propagation and gradient calculation.
+For fair comparison against the classical algorithm, we use a custom classical operator that directly calls \texttt{sgemm}, which significantly outperformed TensorFlow's built-in matrix multiplication operator for fully connected layers.
+
+We show an MLP network with two hidden layers in \cref{fig:neuralnet}.
+The dimensions in the figure match those of the accuracy experiment described in \cref{sec:NNacc}.
+For the performance experiment of \cref{sec:NNperf}, we use an MLP with four hidden layers and varying numbers of nodes in the hidden layers.
+
+\subsection{Accuracy}
 \label{sec:NNacc}
 
-To measure the effect of the matrix multiplication error introduced from the APA algorithms on neural network accuracy, we trained a multi-layer perceptron (MLP) network on the MNIST dataset \cite{LBBH98} and measured its accuracy over the course of training.
-The network was created using Tensorflow version 2.2.0.
+To measure the effect of the matrix multiplication error introduced from the APA algorithms on neural network accuracy, we trained an MLP network on the MNIST dataset \cite{LBBH98} and measured its accuracy over the course of training.
 
@@
 The matrix multiplication error introduced from these algorithms had minimal effect on the training error; all algorithms yielded convergence to nearly full accuracy after 20 epochs or so.
-We see more variability in the test error due to the sensitivity inherent in generalizability of the model, though all algorithms achieve between 97\% and 99\% test accuracy.
+We see more variability in the test error due to the sensitivity inherent in the generalizability of the model, though all algorithms achieve between 97\% and 99\% test accuracy.
 
-\begin{figure}
-\centering
-\renewcommand{\datafile}{data/NN_acc.dat}
-\renewcommand{\set}{train}
-\input{plots/NN_acc}
-\caption{MLP training accuracy for MNIST data set}
-\label{fig:NN_acc_train}
-\end{figure}
-
-\begin{figure}
-\centering
-\renewcommand{\datafile}{data/NN_acc.dat}
-\renewcommand{\set}{test}
-\input{plots/NN_acc}
-\caption{MLP test accuracy for MNIST data set}
-\label{fig:NN_acc_test}
-\end{figure}
-
-\section{Neural Network Performance}
+\subsection{Performance}
 \label{sec:NNperf}
 
-\begin{enumerate}
-	\item describe synthetic network and implementation details
-	\item perf results for growing dim (maybe for serial and 12-threads separately)
-	\item parallel speedup results
-\end{enumerate}
-
-\begin{figure}
+\begin{figure*}[t!]
+%
 \subfloat[One thread]{
@@
 \renewcommand{\datafile}{data/NN_seq.dat}
+\renewcommand{\threads}{1}
 \input{plots/NN_reltime}
 \label{fig:NNreltime_seq}}
-\\
+%
+\hfill
 \subfloat[Six threads]{
@@
 \renewcommand{\datafile}{data/NN_par_6.dat}
-%\input{plots/NN_reltime}
+\renewcommand{\threads}{6}
+\input{plots/NN_reltime}
 \label{fig:NNreltime_par_6}}
-\\
+%
+\hfill
 \subfloat[Twelve threads]{
@@
 \renewcommand{\datafile}{data/NN_par_12.dat}
-%\input{plots/NN_reltime}
+\renewcommand{\threads}{12}
+\input{plots/NN_reltime}
 \label{fig:NNreltime_par_12}}
+%
 \caption{Network training time relative to using classical matrix multiplication}
 \label{fig:NNreltime}
-\end{figure}
+\end{figure*}
+
+The neural networks used for performance benchmarking were 6-layer MLPs (4 hidden layers), and were trained on the MNIST dataset. 
+We base the structure of the network on the fully connected canonical model of ParaDnn \cite{WWB20}.
+The purpose of these experiments was to measure the speed up in training time provided by the APA algorithms and not to measure the accuracy of the networks. 
+The APA algorithms are used only in the multiplications in the hidden layers during both forward and backward propagation; the standard operation was used in the input and output layers.
+
+Following the methodology of ParaDnn, we measure the performance of each algorithm with the number of nodes in each hidden layer varying from 512 to 8192. 
+We also match the batch size of each network to the number of nodes in the hidden layer to produce square matrix multiplications in the hidden layers. 
+\cref{fig:NNreltime} shows the empirical results using 1, 6, and 12 threads.
+
+As shown in \cref{fig:NNreltime_seq}, the single-threaded results begin to show speed up over the classical algorithm when the matrix dimensions are 1024.
+The highest performing algorithm is $\dims{4,4,4}$, which obtains a 25\% speed up at dimension 8192. 
+All algorithms outperform classical for dimensions 4096 and 8192.
+
+\Cref{fig:NNreltime_par_6} shows the results for six threads (one socket).
+As described in \cref{sec:matmul_perf_par}, the APA algorithms do not scale to higher numbers of cores as efficiently as \texttt{sgemm}, due to both the overhead of matrix additions (a memory bandwidth bottleneck) and reduced parallel performance of \texttt{sgemm} for smaller dimensions.
+In this case, not all APA algorithms outperform the classical algorithm.
+The highest performing algorithms are $\dims{4,4,2}$ and $\dims{4,4,4}$, which obtain an improvement of 13\% for dimension 8192.
+
+In the case of 12 threads (both sockets), as shown in \cref{fig:NNreltime_par_12}, most APA algorithms underperform the classical algorithm.
+These results match the behavior of standalone matrix multiplications using 12 threads (see \cref{sec:matmul_perf_par}); the lack of scaling is due in large part to the poor performance of multithreaded \texttt{sgemm} on remainder multiplications of small dimensions.
+The algorithm with a number of sub-multiplications that is a multiple of 12, $\dims{4,4,2}$, continues to perform well and is slightly faster than the classical algorithm for dimensions 4096 and 8192, attaining speedups of up to 5\%.
 
@@
 The results presented here show that neural network training, particularly for Multi-Layer Perceptron networks, is robust to such matrix multiplication error, and APA algorithms are a useful tool for accelerating the costly training computations.
-We observe that for APA algorithms offering a theoretical speedup of up to 39\%, we can achieve up to 28\% and 21\% improvements for single-threaded and multi-threaded implementations of individual matrix multiplications, respectively, which translate to up to \GB{W\%} speedup in overall training time on a multicore CPU.
+We observe that for APA algorithms offering a theoretical speedup of up to 39\%, we can achieve up to 28\% and 21\% improvements for single-threaded and multi-threaded implementations of individual matrix multiplications, respectively, which translate to up to 13\% speedup in overall training time on a multicore CPU.
 
@@
 
+\section*{Acknowledgments}
+
+This material is based upon work supported by the NSF Grant No. CCF1942892 and a Wake Forest University Undergraduate Research and Creative Activities (URECA) Research Fellowship.
+
 \bibliographystyle{IEEEtran}

### 2021-03-06 latex/MPP21/paper.tex
switching to new matmul perf data

@@
 In the performance results for 6 threads (\cref{fig:matmul_par_6}), we see that many of the algorithms start to outperform classical around dimension 2000, though some poor performance is observed for algorithms and particular matrix dimensions.
-The fastest algorithms (e.g., $\dims{4,4,4}$ and $\dims{4,4,2}$) achieve a speedup of up to 23\% over \texttt{gemm} and exceed the machine peak for classical algorithms.
+The fastest algorithms (e.g., $\dims{4,4,4}$ and $\dims{4,4,2}$) achieve a speedup of up to 25\% over \texttt{gemm} and exceed the machine peak for classical algorithms.
 We note that the $\dims{4,4,2}$ algorithm has 24 subproblems, which is a multiple of 6, so there are no remainder subproblems that require all threads.

### 2021-03-07 latex/MPP21/paper.tex
updated abstract with sequential NN speedup number
fix small typos

@@
 Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no significant reduction in the training or testing error.
-Our performance results on a large MLP network show overall multithreaded performance improvements of up to 13\%.
+Our performance results on a large MLP network show overall sequential and multithreaded performance improvements of up to 25\% and 13\%, respectively.
 \end{abstract}
@@
 Algorithms that offer more potential speedup tend to be less accurate.
-Two algorithms with smaller error than expected are $\dims{5,5,5}$ and ${7,2,2}$, which is explained by the coefficients of those algorithms including fraction pre-factors.
+Two algorithms with smaller error than expected are $\dims{5,5,5}$ and $\dims{7,2,2}$, which is explained by the coefficients of those algorithms including fraction pre-factors.
 While $\phi$ is computed using the exponents of the largest intermediate term of $\lambda$, the leading term of $\dims{5,5,5}$ has a constant of $1/4$, lessening its magnitude. 
@@
 All linear combinations are parallelized in a straightforward way, in order to maximize the memory bandwidth of the machine.
-The multiplications are parallelized as follows: given $r$ multiplications and $p$ threads, with $r=p\cdot q + \ell$ for integers $q$ and $\ell<r$, we assign each thread $q$ multiplications to be performed independently, and the remaining $\ell$ multiplications are performed by all threads using the multithreaded implementation of \texttt{gemm}.
+The multiplications are parallelized as follows: given $r$ multiplications and $p$ threads, with $r=p\cdot q + \ell$ for integers $q$ and $\ell<p$, we assign each thread $q$ multiplications to be performed independently, and the remaining $\ell$ multiplications are performed by all threads using the multithreaded implementation of \texttt{gemm}.
 \Cref{fig:hybrid} shows the hybrid strategy for $r=10$ (Bini's algorithm) and $p=4$.

### 2021-03-07 latex/MPP21/paper.tex
cleaned up new data

@@
 \label{eq:APAapprox}
-\hat C = A\cdot B + \lambda E + O(\lambda^2).
+\hat C = A\cdot B + \lambda E + O(\lambda^2),
 \end{equation}
+where $E$ is an error matrix that depends on $A$ and $B$ and corresponds to the leading term of the error polynomial.
 Thus, in exact arithmetic, letting $\lambda \rightarrow 0$ achieves arbitrarily small approximation error.
@@
 \begin{align*}
-C_{11} =& \lambda^{-1} (M_1 + M_2 - M_3 + M_4) \\
-C_{12} =& \lambda^{-1} (-M_3 + M_5) \\
-C_{21} =& M_4 + M_6 - M_{10}\\
-C_{22} =& M_1 - M_5 + M_9\\
-C_{31} =& \lambda^{-1} (-M_8 + M_{10}) \\
-C_{32} =& \lambda^{-1} (M_6 + M_7 - M_8 + M_9).\\
+\hat C_{11} =& \lambda^{-1} (M_1 + M_2 - M_3 + M_4) \\
+\hat C_{12} =& \lambda^{-1} (-M_3 + M_5) \\
+\hat C_{21} =& M_4 + M_6 - M_{10}\\
+\hat C_{22} =& M_1 - M_5 + M_9\\
+\hat C_{31} =& \lambda^{-1} (-M_8 + M_{10}) \\
+\hat C_{32} =& \lambda^{-1} (M_6 + M_7 - M_8 + M_9).\\
 \end{align*}
 
+\GB{give explicit expression for $\hat{C}_{11}$ and reference \cref{eq:APAapprox}}
 We highlight several properties of the rule which are common across all APA algorithms we consider.
@@
 \end{equation}
-The first two matrices specify the linear combinations taken of entries of $A$ and $B$, and the third matrix specifies the contributions of $M_1$ to the entries of $C$.
+The first two matrices specify the linear combinations taken of entries of $A$ and $B$, and the third matrix specifies the contributions of $M_1$ to the entries of $\hat C$.
 Ten such triplets completely specify Bini's algorithm.
@@
 These results match the behavior of standalone matrix multiplications using 12 threads (see \cref{sec:matmul_perf_par}); the lack of scaling is due in large part to the poor performance of multithreaded \texttt{sgemm} on remainder multiplications of small dimensions.
-The algorithm with a number of sub-multiplications that is a multiple of 12, $\dims{4,4,2}$, continues to perform well and is slightly faster than the classical algorithm for dimensions 4096 and 8192, attaining speedups of up to 5\%.
+The algorithm with a number of sub-multiplications that is a multiple of 12, $\dims{4,4,2}$, continues to perform well and is slightly faster than the classical algorithm for dimensions 4096 and 8192, attaining speedups of up to 7\%.
 

### 2021-03-07 latex/MPP21/paper.tex
forgot to add paper on previous commit

@@
 Training large neural networks is computationally expensive and the cost has spurred a surge of research into more efficient hardware, better algorithms, and techniques for trading off accuracy for performance.
+\JW{try to add example stat on how much time/money is spent on NN training}
 For instance, low-precision arithmetic has been shown to decrease running time with little to no effect on the ultimate learning task \cite{GAGN15,HCSEB17}, and new floating point formats have been developed and supported in hardware to implement highly efficient low- and mixed-precision computation \cite{KM+19,YWC20}.

### 2021-03-07 latex/MPP21/paper.tex
fixed et al configuration for bib items

--- 2021-03-07 latex/MPP21/paper.tex
@@
 \begin{document}
+
+\bstctlcite{IEEE:BSTcontrol}
 
@@
 Training large neural networks is computationally expensive and the cost has spurred a surge of research into more efficient hardware, better algorithms, and techniques for trading off accuracy for performance.
+\JW{try to add example stat on how much time/money is spent on NN training \cite{BM+20}}
 For instance, low-precision arithmetic has been shown to decrease running time with little to no effect on the ultimate learning task \cite{GAGN15,HCSEB17}, and new floating point formats have been developed and supported in hardware to implement highly efficient low- and mixed-precision computation \cite{KM+19,YWC20}.

### 2021-03-07 latex/MPP21/paper.tex
add GPT-3 comment
add example expression for bini

--- 2021-03-07 latex/MPP21/paper.tex
@@
 For instance, low-precision arithmetic has been shown to decrease running time with little to no effect on the ultimate learning task \cite{GAGN15,HCSEB17}, and new floating point formats have been developed and supported in hardware to implement highly efficient low- and mixed-precision computation \cite{KM+19,YWC20}.
+\JW{Open-AI's most recent language model, GPT-3, spent the equivalent of days training on a petaflop machine.
+Even a slight reduction in flops could save a large amount of time and reduce server costs.}\cite{BM+20}
 
@@
 
+We can also expressed for $\hat{C}_{11}$ as $\hat C_{11} = A_{11} \cdot B_{11} + A_{12} \cdot B_{21} - \lambda (A_{12} \cdot B_{11})$, where $E = A_{12} \cdot B_{11}$ is the error matrix.
 \GB{give explicit expression for $\hat{C}_{11}$ and reference \cref{eq:APAapprox}}

### 2021-03-07 latex/MPP21/paper.tex
incorporated OpenAI stat

@@
 
-\emph{Fast} matrix multiplication algorithms are those that perform fewer than the $2n^3+O(n^2)$ flops performed by the classical algorithm.
+\emph{Fast} matrix multiplication algorithms are those that perform fewer than the $2n^3+O(n^2)$ floating point operations (flops) performed by the classical algorithm.
 For example, Strassen's original fast matrix multiplication algorithm performs $O(n^{\log_2 7})$ flops, where $\log_27 \approx 2.81$ \cite{Strassen69}.
@@
 Our goal in this paper is to demonstrate that APA algorithms can offer practical performance improvements for applications that are tolerant to error in matrix multiplications, notably the training phase of neural networks.
-Training large neural networks is computationally expensive and the cost has spurred a surge of research into more efficient hardware, better algorithms, and techniques for trading off accuracy for performance.
+Training large neural networks can be incredibly computationally expensive.
+For example,  OpenAI's most recent language model, GPT-3, spent the equivalent of thousands of days training on a petaflop machine \cite{BM+20}. 
+Even a slight reduction in flops could save a large amount of time and money.
+This steep cost has spurred a surge of research into more efficient hardware, better algorithms, and techniques for trading off accuracy for performance.
 For instance, low-precision arithmetic has been shown to decrease running time with little to no effect on the ultimate learning task \cite{GAGN15,HCSEB17}, and new floating point formats have been developed and supported in hardware to implement highly efficient low- and mixed-precision computation \cite{KM+19,YWC20}.
-\JW{Open-AI's most recent language model, GPT-3, spent the equivalent of days training on a petaflop machine.
-Even a slight reduction in flops could save a large amount of time and reduce server costs.}\cite{BM+20}
 
@@
 
-We can also expressed for $\hat{C}_{11}$ as $\hat C_{11} = A_{11} \cdot B_{11} + A_{12} \cdot B_{21} - \lambda (A_{12} \cdot B_{11})$, where $E = A_{12} \cdot B_{11}$ is the error matrix.
 \GB{give explicit expression for $\hat{C}_{11}$ and reference \cref{eq:APAapprox}}

### 2021-03-07 latex/MPP21/paper.tex
incorporated C_11 example from Bini's alg

@@
 \end{align*}
+For example, this rule computes the first entry of the output matrix as $\hat C_{11} = A_{11}  B_{11} + A_{12}  B_{21} - \lambda A_{12}  B_{11}$.
+Thus, in this case, the first entry of the error matrix $E$ in \cref{eq:APAapprox} is $E_{11} = A_{12}  B_{11}$.
 
-\GB{give explicit expression for $\hat{C}_{11}$ and reference \cref{eq:APAapprox}}
 We highlight several properties of the rule which are common across all APA algorithms we consider.

### 2021-03-07 latex/MPP21/paper.tex
edits during final pass

@@
 Practical APA algorithms provide significant reduction in computation time and still provide enough accuracy for many applications like neural network training.
-We demonstrate that APA algorithms can be efficiently implemented and parallelized for shared-memory platforms to obtain up to 28\% and 21\% speedups over the fastest implementation of the classical algorithm using one core and 12 cores, respectively.
+We demonstrate that APA algorithms can be efficiently implemented and parallelized for multicore CPUs to obtain up to 28\% and 21\% speedups over the fastest implementation of the classical algorithm using one core and 12 cores, respectively.
 Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no significant reduction in the training or testing error.
@@
 \emph{Fast} matrix multiplication algorithms are those that perform fewer than the $2n^3+O(n^2)$ floating point operations (flops) performed by the classical algorithm.
-For example, Strassen's original fast matrix multiplication algorithm performs $O(n^{\log_2 7})$ flops, where $\log_27 \approx 2.81$ \cite{Strassen69}.
+For example, Strassen's original fast matrix multiplication algorithm performs $O(n^{2.81})$ flops \cite{Strassen69}.
 The true complexity of matrix multiplication, typically measured as the exponent $\omega$ for complexity $O(n^\omega)$ is an open question, but the current tightest upper bound is 2.37286 \cite{AW21}.
@@
 For example,  OpenAI's most recent language model, GPT-3, spent the equivalent of thousands of days training on a petaflop machine \cite{BM+20}. 
-Even a slight reduction in flops could save a large amount of time and money.
-This steep cost has spurred a surge of research into more efficient hardware, better algorithms, and techniques for trading off accuracy for performance.
+Even a slight reduction in flops could save a significant amount of time and money.
+These steep costs have spurred a surge of research into more efficient hardware, better algorithms, and innovative techniques for trading off accuracy for performance.
 For instance, low-precision arithmetic has been shown to decrease running time with little to no effect on the ultimate learning task \cite{GAGN15,HCSEB17}, and new floating point formats have been developed and supported in hardware to implement highly efficient low- and mixed-precision computation \cite{KM+19,YWC20}.
@@
 \begin{enumerate}
-	\item curate a collection of well-established and recently discovered practical APA algorithms;
+	\item curate a collection of both well-established and recently discovered practical APA algorithms;
 	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them, achieving up to 21\% performance improvement over the best parallel classical implementation,
@@
 
-A key characteristic of matrix multiplication, first demonstrated by Bini et al.~\cite{BCRL79}, is that it can be approximated to arbitrary accuracy more efficiently than being computed exactly.
+A key characteristic of matrix multiplication, first demonstrated by Bini et al.~\cite{BCRL79}, is that it can be approximated to arbitrary accuracy with less computation than required by exact algorithms.
 An APA algorithm is one that takes as input $A$, $B$, and a scalar parameter $0<\lambda<1$ and computes 
@@
 Note that the legend is ordered according to the error parameters, and the empirical error generally follows this ordering.
-The two most accurate APA algorithms are the least recently discovered: Bini's $\dims{3,2,2}$ and Schonhage's $\dims{3,3,3;21}$.
+The two most accurate APA algorithms are the least recently discovered: Bini's $\dims{3,2,2}$ and Schonhage's $\dims{3,3,3;21}$ (note that we indicate the rank of the algorithm here to distinguish from the alternative rank-20 algorithm with the same dimensions).
 Algorithms that offer more potential speedup tend to be less accurate.
-Two algorithms with smaller error than expected are $\dims{5,5,5}$ and $\dims{7,2,2}$, which is explained by the coefficients of those algorithms including fraction pre-factors.
+Two algorithms with smaller error than expected are $\dims{5,5,5}$ and $\dims{7,2,2}$, which is explained by the coefficients of those algorithms including fractional pre-factors.
 While $\phi$ is computed using the exponents of the largest intermediate term of $\lambda$, the leading term of $\dims{5,5,5}$ has a constant of $1/4$, lessening its magnitude. 
@@
 We are particularly interested in algorithms with rules for \emph{small} fixed sizes because they have more promise for practical performance.
-This is because larger fixed sizes result in multiplications of small submatrices, and matrix multiplication performance degrades for smaller dimensions.
+This is because larger fixed sizes result in multiplications of small submatrices, and classical matrix multiplication performance degrades for smaller dimensions.
 For instance, consider a rule for dimensions $\dims{4,4,4}$ applied to matrices of reasonable size, less than dimension $10{,}000$.
@@
 Instead of focusing on the exponent of the asymptotic complexity of fast algorithms, for practical algorithms we are more interested in the constant reduction in flops of a single recursive level (a single use of the rule of the algorithm).
-This is because in practice, for reasonable matrix dimensions (no more than 10,000, say), only 1 or 2 recursive levels will yield performance improvement \cite{BB15}.
+This is because in practice, for reasonable matrix dimensions, only 1 or 2 recursive levels will yield performance improvement \cite{BB15}.
 We also prefer algorithms with fewer nonzero coefficients in the linear combinations, because while less costly than multiplications, the matrix additions are less efficient (they are memory bandwidth bound) and prevent achieving the ideal speedup given by the reduction in multiplications.
@@
 
-We exploit the common algorithmic structure across APA (and exact) algorithms to develop a unified strategy of high performance multithreaded implementation using C++/OpenMP.
-We build upon the work of Benson and Ballard \cite{BB15}, using code generation to apply the strategy to each algorithm.
-
-\subsection{Experimental Platform}
-
-All experiments are performed on a dual-socket Intel Xeon E5-2620 (Sandy Bridge) server with 2 sockets each with 6 cores. 
-Each socket has a 15 MB L3 cache, and each core has a 256 KB L2 cache and 32 KB L1 data cache. 
-Each core has a clock rate of 2.00 GHz (with turbo boost disabled) and peak single precision flop rate of 32 GFLOPS. 
-Our code is compiled with GCC version 7.5.0, 
-and we use Intel's Math Kernel Library (MKL) version 2019.4.243.
-
-\subsection{Hybrid Parallelization Strategy}
-
-As described in \cref{sec:APAalgs}, each algorithm can be encoded by a set of triplets of coefficient matrices.
-From this representation, we generate recursive code that computes the linear combinations for the inputs to each multiplication and combines the outputs.
-The linear combinations are computed using a ``write-once'' strategy that was found to be most efficient in terms of memory bandwidth and performance.
-The multiplications are either computed recursively or via a call to \texttt{gemm}, the interface to a highly efficient BLAS implementation of the classical algorithm.
-Because we use only 1 recursive step in our experiments, there are no recursive calls, and every multiplication is performed by \texttt{gemm}.
-We note that the original code generation tool was designed primarily for exact algorithms, and we generalized it to apply to all of the APA algorithms we consider here, some of which have more complicated coefficients than previously considered.
-
-We adopt the ``hybrid'' parallelization strategy proposed by Benson and Ballard \cite{BB15}.
-All linear combinations are parallelized in a straightforward way, in order to maximize the memory bandwidth of the machine.
-The multiplications are parallelized as follows: given $r$ multiplications and $p$ threads, with $r=p\cdot q + \ell$ for integers $q$ and $\ell<p$, we assign each thread $q$ multiplications to be performed independently, and the remaining $\ell$ multiplications are performed by all threads using the multithreaded implementation of \texttt{gemm}.
-\Cref{fig:hybrid} shows the hybrid strategy for $r=10$ (Bini's algorithm) and $p=4$.
-
 \begin{figure}
@@
 \end{figure}
-
-The hybrid strategy is efficient because each thread can achieve close to the peak performance of a core when computing independent matrix multiplications, even for relatively small problems.
-The alternative strategy of employing multithreaded \texttt{gemm} for each of the $r$ multiplications (known as ``DFS'') suffers performance degradation for small problems, where the parallel implementation attains a smaller fraction of peak.
-The hybrid strategy also perfectly load balances the computation across threads, as opposed to the alternate strategy of assigning the $\ell$ remainder multiplications to $\ell$ different threads (known as ``BFS''), leaving the other $p-\ell$ threads idle.
 
@@
 
+We exploit the common algorithmic structure across APA (and exact) algorithms to develop a unified strategy of high performance multithreaded implementation using C++/OpenMP.
+We build upon the work of Benson and Ballard \cite{BB15}, using code generation to apply the strategy to each algorithm.
+
+\subsection{Experimental Platform}
+
+All experiments are performed on a dual-socket Intel Xeon E5-2620 (Sandy Bridge) server with 2 sockets each with 6 cores. 
+Each socket has a 15 MB L3 cache, and each core has a 256 KB L2 cache and 32 KB L1 data cache. 
+Each core has a clock rate of 2.00 GHz (with turbo boost disabled) and peak single precision flop rate of 32 GFLOPS. 
+Our code is compiled with GCC version 7.5.0, 
+and we use Intel's Math Kernel Library (MKL) version 2019.4.243.
+We use single precision in all experiments.
+
+\subsection{Hybrid Parallelization Strategy}
+
+As described in \cref{sec:APAalgs}, each algorithm can be encoded by a set of triplets of coefficient matrices.
+From this representation, we generate recursive code that computes the linear combinations for the inputs to each multiplication and combines the outputs.
+The linear combinations are computed using a ``write-once'' strategy that was found to be most efficient in terms of memory bandwidth and performance.
+The multiplications are either computed recursively or via a call to \texttt{gemm}, the interface to a highly efficient BLAS implementation of the classical algorithm.
+Because we use only 1 recursive step in our experiments, there are no recursive calls, and every multiplication is performed by \texttt{gemm}.
+We note that the original code generation tool was designed primarily for exact algorithms, and we generalized it to apply to all of the APA algorithms we consider here, some of which have more complicated coefficients than previously considered.
+
+We adopt the ``hybrid'' parallelization strategy proposed by Benson and Ballard \cite{BB15}.
+All linear combinations are parallelized in a straightforward way, in order to maximize the memory bandwidth of the machine.
+The multiplications are parallelized as follows: given $r$ multiplications and $p$ threads, with $r=p\cdot q + \ell$ for integers $q$ and $\ell<p$, we assign each thread $q$ multiplications to be performed independently, and the remaining $\ell$ multiplications are performed by all threads using the multithreaded implementation of \texttt{gemm}.
+\Cref{fig:hybrid} shows the hybrid strategy for $r=10$ (Bini's algorithm) and $p=4$.
+
+The hybrid strategy is efficient because each thread can achieve close to the peak performance of a core when computing independent matrix multiplications, even for relatively small problems.
+The alternative strategy of employing multithreaded \texttt{gemm} for each of the $r$ multiplications (known as ``DFS'') suffers performance degradation for small problems, where the parallel implementation attains a smaller fraction of peak.
+The hybrid strategy also perfectly load balances the computation across threads, as opposed to the alternate strategy of assigning the $\ell$ remainder multiplications to $\ell$ different threads (known as ``BFS''), leaving the other $p-\ell$ threads idle.
+
 \subsection{Sequential Performance}
 
-\Cref{fig:matmul_seq} reports the sequential performance of the APA algorithms in comparison to the most efficient implementation of the classical algorithm, MKL's \texttt{sgemm}.
+\Cref{fig:matmul_seq} reports the sequential performance of the APA algorithms in comparison to the most efficient implementation of the classical algorithm, MKL's single-precision \texttt{gemm}.
 The y-axis is the \emph{effective} GFLOPS, which is measured as $\num{1e-9} \cdot 2n^3 / \text{ time}$.
@@
 We vary the dimension from 512 up to 8192 and see that all algorithms outperform classical for dimensions larger than 2000 or so.
-The highest performing algorithm is $\dims{4,4,4}$, and at dimension 8192, it is 28\% faster than \texttt{sgemm}.
-Ignoring the cost of the matrix additions, $\dims{4,4,4}$ performs 39\% fewer flops than the classical algorithm; the drop to 28\% achieved improvement is because of the overhead of matrix additions and the reduced performance of \texttt{sgemm} on smaller matrices.
+The highest performing algorithm is $\dims{4,4,4}$, and at dimension 8192, it is 28\% faster than \texttt{gemm}.
+Ignoring the cost of the matrix additions, $\dims{4,4,4}$ performs 39\% fewer flops than the classical algorithm; the drop to 28\% achieved improvement is because of the overhead of matrix additions and the reduced performance of \texttt{gemm} on smaller matrices.
 We note that $\dims{4,4,2}$ is also high performing, achieving a 25\% observed improvement (out of a theoretical 33\%), along with $\dims{3,3,3;20}$ and $\dims{5,5,5}$.
@@
 While we can expect close to linear speedup with cores on the multiplications (which are perfectly load balanced), linear scaling is impossible to achieve with the additions.
-We also note that when the hybrid method uses all threads on remainder multiplications, the smaller dimensions make it harder for \texttt{sgemm} to scale as well as it can on the original matrix dimensions.
+We also note that when the hybrid method uses all threads on remainder multiplications, the smaller dimensions make it harder for multithreaded \texttt{gemm} to scale as well as it can on the original matrix dimensions.
 
@@
 Here we see a majority of the algorithms are slower than the classical algorithm, even for large matrices.
-We attribute the poor performance to the effect of lower \texttt{sgemm} performance for smaller matrices, as well as a lack of NUMA-aware optimization.
-The ``ramp-up'' range of \texttt{sgemm} performance is much shallower for 12 threads than for 6 threads, not achieving the plateau performance until dimension 4000 or so.
+We attribute the poor performance to the effect of lower \texttt{gemm} performance for smaller matrices, as well as a lack of NUMA-aware optimization.
+The ``ramp-up'' range of \texttt{gemm} performance is much shallower for 12 threads than for 6 threads, not achieving the plateau performance until dimension 4000 or so.
 This implies that the dimension of the submultiplications does not fall on the plateau for any APA algorithm, so the remainder multiplications suffer from poor parallel performance.
 The algorithm with no remainder multiplications, $\dims{4,2,2}$, does not suffer this problem and maintains higher performance.
-It exceeds \texttt{sgemm}'s effective performance at dimension 4000, exceeds the peak parallel performance of any classical algorithm for larger dimensions, and achieves a 21\% speedup over \texttt{sgemm} at dimension 8192 for an effective rate of 389 GFLOPS.
+It exceeds \texttt{gemm}'s effective performance at dimension 4000, exceeds the peak parallel performance of any classical algorithm for larger dimensions, and achieves a 21\% speedup over \texttt{gemm} at dimension 8192 for an effective rate of 389 GFLOPS.
 
 \section{Application of APA Algorithms in Multi-Layer Perceptron Network Training}
-
-\begin{figure}[b]
-\centering
-\input{figs/neuralnet}
-\caption{Multi-Layer Perceptron network structure}
-\label{fig:neuralnet}
-\end{figure}
 
@@
 
+\begin{figure}[hb!]
+\centering
+\input{figs/neuralnet}
+\caption{Multi-Layer Perceptron network structure}
+\label{fig:neuralnet}
+\end{figure}
+
 \subsection{MLP Network Structure}
@@
 For internal layers, we use the custom operator for both forward propagation and gradient calculation.
-For fair comparison against the classical algorithm, we use a custom classical operator that directly calls \texttt{sgemm}, which significantly outperformed TensorFlow's built-in matrix multiplication operator for fully connected layers.
+For fair comparison against the classical algorithm, we use a custom classical operator that directly calls \texttt{gemm}, which significantly outperformed TensorFlow's built-in matrix multiplication operator for fully connected layers.
 
@@
 Training is performed using batched stochastic gradient descent with a batch size of 300.
-APA algorithms were used for the middle multiplication of the network (yielding matrix multiplication dimensions $300\times 300\times300$), while the input and output layers use the classical matrix multiplication algorithm.
-The APA algorithms were also used during back propagation in the corresponding multiplications in the center of the network.
-One network was trained for 50 epochs for each of the APA algorithms and one was trained using the classical matrix multiplication algorithm.
+APA algorithms are used for the middle multiplication of the network (yielding matrix multiplication dimensions $300\times 300\times300$), while the input and output layers use the classical matrix multiplication algorithm.
+The APA algorithms are also used during back propagation in the corresponding multiplications in the center of the network.
+One network is trained for 50 epochs for each of the APA algorithms, and one is trained using the classical matrix multiplication algorithm.
 
 \Cref{fig:NN_acc_train} plots the accuracy on the training data over each epoch, and \cref{fig:NN_acc_test} shows the test accuracy over epochs. 
-The matrix multiplication error introduced from these algorithms had minimal effect on the training error; all algorithms yielded convergence to nearly full accuracy after 20 epochs or so.
+The matrix multiplication error introduced from these algorithms has minimal effect on the training error; all algorithms yield convergence to nearly full accuracy after 20 epochs or so.
 We see more variability in the test error due to the sensitivity inherent in the generalizability of the model, though all algorithms achieve between 97\% and 99\% test accuracy.
@@
 
-As shown in \cref{fig:NNreltime_seq}, the single-threaded results begin to show speed up over the classical algorithm when the matrix dimensions are 1024.
+As shown in \cref{fig:NNreltime_seq}, the single-threaded results begin to show speed up over the classical algorithm when the matrix dimensions are at least 1024.
 The highest performing algorithm is $\dims{4,4,4}$, which obtains a 25\% speed up at dimension 8192. 
@@
 \Cref{fig:NNreltime_par_6} shows the results for six threads (one socket).
-As described in \cref{sec:matmul_perf_par}, the APA algorithms do not scale to higher numbers of cores as efficiently as \texttt{sgemm}, due to both the overhead of matrix additions (a memory bandwidth bottleneck) and reduced parallel performance of \texttt{sgemm} for smaller dimensions.
+As described in \cref{sec:matmul_perf_par}, the APA algorithms do not scale to higher numbers of cores as efficiently as \texttt{gemm}, due to both the overhead of matrix additions (a memory bandwidth bottleneck) and reduced parallel performance of \texttt{gemm} for smaller dimensions.
 In this case, not all APA algorithms outperform the classical algorithm.
@@
 In the case of 12 threads (both sockets), as shown in \cref{fig:NNreltime_par_12}, most APA algorithms underperform the classical algorithm.
-These results match the behavior of standalone matrix multiplications using 12 threads (see \cref{sec:matmul_perf_par}); the lack of scaling is due in large part to the poor performance of multithreaded \texttt{sgemm} on remainder multiplications of small dimensions.
+These results match the behavior of standalone matrix multiplications using 12 threads (see \cref{sec:matmul_perf_par}); the lack of scaling is due in large part to the poor performance of multithreaded \texttt{gemm} on remainder multiplications of small dimensions.
 The algorithm with a number of sub-multiplications that is a multiple of 12, $\dims{4,4,2}$, continues to perform well and is slightly faster than the classical algorithm for dimensions 4096 and 8192, attaining speedups of up to 7\%.
@@
 The results presented here show that neural network training, particularly for Multi-Layer Perceptron networks, is robust to such matrix multiplication error, and APA algorithms are a useful tool for accelerating the costly training computations.
-We observe that for APA algorithms offering a theoretical speedup of up to 39\%, we can achieve up to 28\% and 21\% improvements for single-threaded and multi-threaded implementations of individual matrix multiplications, respectively, which translate to up to 13\% speedup in overall training time on a multicore CPU.
+We observe that for APA algorithms offering a theoretical speedup of up to 39\%, we can achieve up to 28\% and 21\% improvements for single- and multi-threaded implementations of individual matrix multiplications, respectively, which translate to up to 13\% speedup in overall training time on a multicore CPU.
 
@@
 We note that an algorithm for dimensions $\dims{m,n,k}$ can be translated into an algorithm for $\dims{n,m,k}$ and any other reordering of the dimensions \cite{BB15}.
-For very large matrices, it may also be beneficial to consider more than one recursive step of the same algorithm or to use a combination of two or three different algorithms across recursive steps (i.e., uniform, non-stationary algorithms \cite{BBDLS16}).
+For very large matrices, it may also be beneficial to consider more than one recursive step of the same algorithm or a combination of two or three different algorithms across recursive steps (uniform, non-stationary algorithms \cite{BBDLS16}).
 
@@
 Techniques used to accelerate Strassen's algorithm on the GPU \cite{HYvdG20,LAES13} can be generalized to other fast algorithms as done in work for multicore CPUs \cite{BB15,HRMvdG17}.
+We believe improved performance benefits can be achieved by APA algorithms on GPU architectures due to the relatively higher memory bandwidth.
 
-\section*{Acknowledgments}
+\paragraph*{Acknowledgments}
 
 This material is based upon work supported by the NSF Grant No. CCF1942892 and a Wake Forest University Undergraduate Research and Creative Activities (URECA) Research Fellowship.
+
+\vspace{-.175cm}
 

### 2021-04-19 TF_op/VGG11.py
add vgg11

@@ -1,0 +1,310 @@
+"""VGG19 model for Keras.
+
+Code copied from: https://github.com/tensorflow/tensorflow/blob/v2.4.1/tensorflow/python/keras/applications/vgg19.py
+
+# tensorflow 2.4.1
+
+Reference:
+  - [Very Deep Convolutional Networks for Large-Scale Image Recognition](
+      https://arxiv.org/abs/1409.1556) (ICLR 2015)
+"""
+
+import time
+import os
+
+import numpy as np
+
+from tensorflow.python.keras import backend
+from tensorflow.python.keras.applications import imagenet_utils
+from tensorflow.python.keras.engine import training
+from tensorflow.python.keras.layers import VersionAwareLayers
+from tensorflow.python.keras.utils import data_utils
+from tensorflow.python.keras.utils import layer_utils
+from tensorflow.python.lib.io import file_io
+from tensorflow.python.util.tf_export import keras_export
+from tensorflow.python.ops import array_ops, gen_math_ops, math_ops
+
+import tensorflow as tf
+from tensorflow import keras
+
+
+WEIGHTS_PATH = ('https://storage.googleapis.com/tensorflow/keras-applications/'
+                'vgg19/vgg19_weights_tf_dim_ordering_tf_kernels.h5')
+WEIGHTS_PATH_NO_TOP = ('https://storage.googleapis.com/tensorflow/'
+                       'keras-applications/vgg19/'
+                       'vgg19_weights_tf_dim_ordering_tf_kernels_notop.h5')
+
+layers = VersionAwareLayers()
+
+
+num_threads = 12
+mm_algo = 'smirnov444'
+epsilon_values = {
+    'bini322': 2**-11,
+    'schonhage333': 2**-5,
+    'smirnov224': 2**-7,
+    'smirnov225': 2**-5,
+    'smirnov272': 2**-3,
+    'smirnov323': 2**-5,
+    'smirnov333': 2**-3,
+    'smirnov334': 2**-5,
+    'smirnov442': 2**-5,
+    'smirnov444': 2**-5,
+    'smirnov552': 2**-5,
+    'smirnov555': 2**-5
+}
+
+
+
+@tf.RegisterGradient("FastMatMul")
+def _Fast_MatMul_grad(op, grad):
+    bt = array_ops.transpose(op.inputs[1])
+    at = array_ops.transpose(op.inputs[0])
[…]

### 2021-04-19 TF_op/VGG16.py

@@ -1,32 +1,0 @@
-import tensorflow as tf
-from tensorflow import keras
-
-if __name__ == '__main__':
-    model = tf.keras.applications.VGG16(include_top=True, weights=None, classes=1000, classifier_activation='softmax')
-
-    model.compile(
-        optimizer=keras.optimizers.RMSprop(),  # Optimizer
-        # Loss function to minimize
-        loss=keras.losses.SparseCategoricalCrossentropy(),
-        # List of metrics to monitor
-        metrics=[keras.metrics.SparseCategoricalAccuracy()],
-    )
-
-    image = tf.image.decode_jpeg(tf.io.read_file('test_image.JPEG'))
-
-    image = tf.keras.applications.vgg19.preprocess_input(image)
-
-    image = tf.image.resize(image, (224, 224))
-    image = tf.reshape(image, [1, 224, 224, 3])
-
-    epochs = 2
-    num_examples = 150
-
-    y_train = tf.ones([num_examples])
-    x_train = []
-    for i in range(num_examples):
-        x_train.append(image)
-
-    x_train = tf.concat(x_train, 0)
-
-    model.fit(x_train, y_train, batch_size=num_examples, epochs=epochs)

### 2021-04-19 TF_op/VGG11.py
cleann up vgg11

@@ -27,13 +27,6 @@ from tensorflow.python.ops import array_ops, gen_math_ops, math_ops
 import tensorflow as tf
 from tensorflow import keras
 
-
-WEIGHTS_PATH = ('https://storage.googleapis.com/tensorflow/keras-applications/'
-                'vgg19/vgg19_weights_tf_dim_ordering_tf_kernels.h5')
-WEIGHTS_PATH_NO_TOP = ('https://storage.googleapis.com/tensorflow/'
-                       'keras-applications/vgg19/'
-                       'vgg19_weights_tf_dim_ordering_tf_kernels_notop.h5')
-
 layers = VersionAwareLayers()
 
 
@@ -106,7 +99,7 @@ def VGG11(
     pooling=None,
     classes=1000,
     classifier_activation='softmax'):
-  """Instantiates the VGG19 architecture.
+  """Instantiates the VGG11 architecture.
   Reference:
   - [Very Deep Convolutional Networks for Large-Scale Image Recognition](
       https://arxiv.org/abs/1409.1556) (ICLR 2015)
@@ -249,25 +242,7 @@ def VGG11(
   else:
     inputs = img_input
   # Create model.
-  model = training.Model(inputs, x, name='vgg19')
-
-  # Load weights.
-  if weights == 'imagenet':
-    if include_top:
-      weights_path = data_utils.get_file(
-          'vgg19_weights_tf_dim_ordering_tf_kernels.h5',
-          WEIGHTS_PATH,
-          cache_subdir='models',
-          file_hash='cbe5617147190e668d6c5d5026f83318')
-    else:
-      weights_path = data_utils.get_file(
-          'vgg19_weights_tf_dim_ordering_tf_kernels_notop.h5',
-          WEIGHTS_PATH_NO_TOP,
-          cache_subdir='models',
-          file_hash='253f8cb515780f3b799900260a226db6')
-    model.load_weights(weights_path)
-  elif weights is not None:
-    model.load_weights(weights)
+  model = training.Model(inputs, x, name='vgg11')
 
   return model
 
@@ -283,7 +258,7 @@ if __name__ == '__main__':
     image = tf.image.resize(image, (224, 224))
     image = tf.reshape(image, [1, 224, 224, 3])
 
-    model = VGG19(include_top=True, weights=None, input_tensor=None, pooling=None)
+    model = VGG11(include_top=True, weights=None, input_tensor=None, pooling=None)
 
     model.compile(
[…]

### 2021-04-19 TF_op/VGG11.py
update batch size

@@ -269,11 +269,11 @@ if __name__ == '__main__':
     )
 
     epochs = 2
-    num_examples = 500
+    batch_size = 500
 
-    y_train = tf.ones([num_examples])
+    y_train = tf.ones([batch_size])
     x_train = []
-    for i in range(num_examples):
+    for i in range(batch_size):
         x_train.append(image)
 
     x_train = tf.concat(x_train, 0)
@@ -281,7 +281,7 @@ if __name__ == '__main__':
     print('\n\nAll images put in memory\n\n')
 
     a = time.time()
-    model.fit(x_train, y_train, batch_size=1024, epochs=epochs)
+    model.fit(x_train, y_train, batch_size=batch_size, epochs=epochs)
     b = time.time()
 
     print(f"Total train time: {b-a} seconds\nTime per epoch: {(b-a)/epochs}")

### 2021-04-19 TF_op/VGG11.py
fix vgg 11

@@ -80,9 +80,6 @@ class Fast_Linear(keras.layers.Layer):
         self.activation = activation
 
     def call(self, inputs):
-        print('\ninputs * weights')
-        print('inputs shape:', inputs.shape)
-        print('weights shape:', self.w.shape)
         output = fast_mm_module.FastMatMul(a_matrix=inputs, b_matrix=self.w, epsilon=self.epsilon, steps=1,
                                                     numthreads=num_threads) + self.b
         if self.activation == 'softmax':
@@ -179,55 +176,55 @@ def VGG11(
       img_input = layers.Input(tensor=input_tensor, shape=input_shape)
     else:
       img_input = input_tensor
-  # Block 1
-  x = layers.Conv2D(
-      64, (3, 3), activation='relu', padding='same', name='block1_conv1')(
-          img_input)
+    # Block 1
+    x = layers.Conv2D(
+        64, (3, 3), activation='relu', padding='same', name='block1_conv1')(
+        img_input)
+    x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block1_pool')(x)
 
-  # Block 2
-  x = layers.Conv2D(
-      128, (3, 3), activation='relu', padding='same', name='block2_conv1')(x)
+    # Block 2
+    x = layers.Conv2D(
+        128, (3, 3), activation='relu', padding='same', name='block2_conv1')(x)
+    x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block2_pool')(x)
 
-  # Block 3
-  x = layers.Conv2D(
-      256, (3, 3), activation='relu', padding='same', name='block3_conv1')(x)
-  x = layers.Conv2D(
-      256, (3, 3), activation='relu', padding='same', name='block3_conv2')(x)
+    # Block 3
+    x = layers.Conv2D(
+        256, (3, 3), activation='relu', padding='same', name='block3_conv1')(x)
+    x = layers.Conv2D(
+        256, (3, 3), activation='relu', padding='same', name='block3_conv2')(x)
+    x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block3_pool')(x)
 
-  # Block 4
-  x = layers.Conv2D(
-      512, (3, 3), activation='relu', padding='same', name='block4_conv1')(x)
-  x = layers.Conv2D(
-      512, (3, 3), activation='relu', padding='same', name='block4_conv2')(x)
+    # Block 4
[…]

### 2021-04-19 TF_op/VGG11.py
change to tabs

@@ -94,88 +94,89 @@ def VGG11(
     input_tensor=None,
     input_shape=None,
     pooling=None,
-    classes=1000,
-    classifier_activation='softmax'):
-  """Instantiates the VGG11 architecture.
-  Reference:
-  - [Very Deep Convolutional Networks for Large-Scale Image Recognition](
-      https://arxiv.org/abs/1409.1556) (ICLR 2015)
-  By default, it loads weights pre-trained on ImageNet. Check 'weights' for
-  other options.
-  This model can be built both with 'channels_first' data format
-  (channels, height, width) or 'channels_last' data format
-  (height, width, channels).
-  The default input size for this model is 224x224.
-  Note: each Keras Application expects a specific kind of input preprocessing.
-  For VGG19, call `tf.keras.applications.vgg19.preprocess_input` on your
-  inputs before passing them to the model.
-  Arguments:
-    include_top: whether to include the 3 fully-connected
-      layers at the top of the network.
-    weights: one of `None` (random initialization),
-        'imagenet' (pre-training on ImageNet),
-        or the path to the weights file to be loaded.
-    input_tensor: optional Keras tensor
-      (i.e. output of `layers.Input()`)
-      to use as image input for the model.
-    input_shape: optional shape tuple, only to be specified
-      if `include_top` is False (otherwise the input shape
-      has to be `(224, 224, 3)`
-      (with `channels_last` data format)
-      or `(3, 224, 224)` (with `channels_first` data format).
-      It should have exactly 3 inputs channels,
-      and width and height should be no smaller than 32.
-      E.g. `(200, 200, 3)` would be one valid value.
-    pooling: Optional pooling mode for feature extraction
-      when `include_top` is `False`.
-      - `None` means that the output of the model will be
-          the 4D tensor output of the
-          last convolutional block.
-      - `avg` means that global average pooling
-          will be applied to the output of the
-          last convolutional block, and thus
-          the output of the model will be a 2D tensor.
-      - `max` means that global max pooling will
-          be applied.
-    classes: optional number of classes to classify images
-      into, only to be specified if `include_top` is True, and
-      if no `weights` argument is specified.
-    classifier_activation: A `str` or callable. The activation function to use
-      on the "top" layer. Ignored unless `include_top=True`. Set
-      `classifier_activation=None` to return the logits of the "top" layer.
-  Returns:
[…]

### 2021-04-20 TF_op/shape_vgg_test.sh
add new matmul benchmark data for small batch size on vgg, and fix dimension problem in python script

@@ -4,16 +4,18 @@ do
 for shape in bs,25,4 bs,4,25 25,bs,4 bs,4,4 4,bs,4 bs,4,1 bs,1,4 4,bs,1
 do
 
+mkdir matmalTest/par_${sq}/${shape}/2
+
 for name in classic dgemm bini322 bini232 bini223 schonhage333 smirnov224 smirnov242 smirnov422 smirnov225 smirnov252 smirnov522 smirnov272 smirnov227 smirnov722 smirnov323 smirnov332 smirnov233 smirnov333 smirnov343 smirnov433 smirnov334 smirnov442 smirnov424 smirnov244 smirnov444 smirnov552 smirnov525 smirnov255 smirnov555
 do
 
-python shape_vgg_time_test.py --tf ${sq} --omp ${sq} --mm ${name} --shape ${shape} --num 1
+python shape_vgg_time_test.py --tf ${sq} --omp ${sq} --mm ${name} --shape ${shape} --num 2
 
-#done
+done
 
 cd matmalTest
 python convert.py --shape ${shape}
-cd ..
+cd ../
 
 done
 

### 2021-04-20 TF_op/shape_vgg_time_test.py

@@ -72,23 +72,27 @@ loops = 3
 times = np.zeros((loops, 4))
 k=0
 
-for dim in range(1024, 4096+1024, 1024):
+for dim in range(256, 1024+256, 256):
     
+    total_dim = 0
     if bs_dim == 0:
         a = tf.Variable(tf.random.uniform(shape=(dim, reg_dim[1] * 1024)), dtype=tf.float32)
         b = tf.Variable(tf.random.uniform(shape=(reg_dim[1] * 1024, reg_dim[2] * 1024)), dtype=tf.float32)
+        total_dim = dim* reg_dim[1] * 1024 * reg_dim[2] * 1024
     elif bs_dim == 1:
         a = tf.Variable(tf.random.uniform(shape=(reg_dim[0] * 1024, dim)), dtype=tf.float32)
         b = tf.Variable(tf.random.uniform(shape=(dim, reg_dim[2] * 1024)), dtype=tf.float32)
+        total_dim = dim* reg_dim[0] * 1024 * reg_dim[2] * 1024
     elif bs_dim == 2:
         a = tf.Variable(tf.random.uniform(shape=(reg_dim[0] * 1024, reg_dim[1] * 1024)), dtype=tf.float32)
         b = tf.Variable(tf.random.uniform(shape=(reg_dim[1] * 1024, dim)), dtype=tf.float32)
+        total_dim = dim* reg_dim[1] * 1024 * reg_dim[0] * 1024
 
     if algo_name == 'classic':
         t1 = time.time()
         regular = tf.matmul(a, b)
         t2 = time.time()
-        times[0][k] = 1e-9 * 2 * (dim**3) / (t2-t1)
+        times[0][k] = 1e-9 * 2 * total_dim / (t2-t1)
     
     else:
         for i in range(loops):
@@ -97,7 +101,7 @@ for dim in range(1024, 4096+1024, 1024):
             op = fast_mm_module.FastMatMul(a_matrix=a, b_matrix=b, epsilon=2**epsilon_, steps=step_, numthreads=args.omp)
             t2 = time.time()
 
-            times[i][k]=  1e-9 * 2 * (dim**3) / (t2-t1)
+            times[i][k]=  1e-9 * 2 * total_dim / (t2-t1)
         
     #print(f'Average relative error: {diff/loops}')
     if bs_dim == 0:

### 2021-04-21 TF_op/VGG11.py
update vgg11 to measure each layer

@@ -23,6 +23,7 @@ from tensorflow.python.keras.utils import layer_utils
 from tensorflow.python.lib.io import file_io
 from tensorflow.python.util.tf_export import keras_export
 from tensorflow.python.ops import array_ops, gen_math_ops, math_ops
+from tensorflow.keras.callbacks import Callback
 
 import tensorflow as tf
 from tensorflow import keras
@@ -212,7 +213,7 @@ def VGG11(
 
     # Classification block
     x = layers.Flatten(name='flatten')(x)
-    print(x.shape)
+
     #x = layers.Dense(4096, activation='relu', name='fc1')(x)
     fast_layer1 = Fast_Linear(units=4096, input_dim=25088, activation='relu')
     x = fast_layer1(x)
@@ -242,6 +243,18 @@ def VGG11(
     return model
 
 
+
+class TimeHistory(Callback):
+    def on_train_begin(self, logs={}):
+        self.times = []
+
+    def on_epoch_begin(self, batch, logs={}):
+        self.epoch_time_start = time.time()
+
+    def on_epoch_end(self, batch, logs={}):
+        self.times.append(time.time() - self.epoch_time_start)
+
+
 if __name__ == '__main__':
 
     fast_mm_module = tf.load_op_library(f'obj/{mm_algo}_mat_mul.so')
@@ -255,16 +268,10 @@ if __name__ == '__main__':
 
     model = VGG11(include_top=True, input_tensor=None, pooling=None)
 
-    model.compile(
-        optimizer=keras.optimizers.RMSprop(),  # Optimizer
-        # Loss function to minimize
-        loss=keras.losses.SparseCategoricalCrossentropy(),
-        # List of metrics to monitor
-        metrics=[keras.metrics.SparseCategoricalAccuracy()],
-    )
+
 
     epochs = 2
-    batch_size = 500
+    batch_size = 250
 
     y_train = tf.ones([batch_size])
     x_train = []
@@ -275,8 +282,44 @@ if __name__ == '__main__':
 
     print('\n\nAll images put in memory\n\n')
 
-    a = time.time()
-    model.fit(x_train, y_train, batch_size=batch_size, epochs=epochs)
-    b = time.time()
+    print(model.layers)
 
-    print(f"Total train time: {b-a} seconds\nTime per epoch: {(b-a)/epochs}")
+
+
+
+    time_callback = TimeHistory()
+
+
+
+    # Loop through each layer setting it Trainable and others as non trainable
+    results = []
[…]

### 2021-04-21 TF_op/VGG11.py
updage to dgemm

@@ -32,7 +32,7 @@ layers = VersionAwareLayers()
 
 
 num_threads = 12
-mm_algo = 'smirnov444'
+mm_algo = 'dgemm'
 epsilon_values = {
     'bini322': 2**-11,
     'schonhage333': 2**-5,
@@ -282,7 +282,8 @@ if __name__ == '__main__':
 
     print('\n\nAll images put in memory\n\n')
 
-    print(model.layers)
+    for l in model.layers:
+        print(l.name)
 
 
 

### 2021-04-22 latex/ICPP21/paper.tex
updated formatting for ICPP

--- 2021-03-07 latex/MPP21/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-%\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
+\documentclass[sigconf]{acmart}
+
+\usepackage{amsmath,amsfonts}
 \usepackage{algorithmic}
@@
 \usepackage{ifthen}
-\usepackage{flushend}
+\usepackage{balance}
 
@@
 
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} } 
 \newcommand{\dims}[1]{\langle #1 \rangle}
@@
 
-\begin{document}
-
-\bstctlcite{IEEE:BSTcontrol}
+\acmConference[ICPP'21]{International Conference on Parallel Processing}{August 9--12}{Chicago, IL, USA}
 
@@
 
-\author{\IEEEauthorblockN{Grey Ballard}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Wake Forest University}\\
-Winston-Salem, NC, USA \\
-\email{ballard@wfu.edu}}
-\and
-\IEEEauthorblockN{Jack Weissenberger}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Wake Forest University}\\
-Winston-Salem, NC, USA \\
-\email{jack.weissenberger@gmail.com}}
-\and
-\IEEEauthorblockN{Luoping Zhang}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Wake Forest University}\\
-Winston-Salem, NC, USA \\
-\email{zhanl317@wfu.edu}}
-}
-
-\maketitle
+%\author{\IEEEauthorblockN{Grey Ballard}
+%\IEEEauthorblockA{\textit{Department of Computer Science} \\
+%\textit{Wake Forest University}\\
+%Winston-Salem, NC, USA \\
+%\email{ballard@wfu.edu}}
+%\and
+%\IEEEauthorblockN{Jack Weissenberger}
+%\IEEEauthorblockA{\textit{Department of Computer Science} \\
+%\textit{Wake Forest University}\\
+%Winston-Salem, NC, USA \\
+%\email{jack.weissenberger@gmail.com}}
+%\and
+%\IEEEauthorblockN{Luoping Zhang}
+%\IEEEauthorblockA{\textit{Department of Computer Science} \\
+%\textit{Wake Forest University}\\
+%Winston-Salem, NC, USA \\
+%\email{zhanl317@wfu.edu}}
+%}
 
@@
 \end{abstract}
+
+\begin{document}
+\maketitle
 
@@
 
-\paragraph*{Acknowledgments}
+%\paragraph*{Acknowledgments}
+%
+%This material is based upon work supported by the NSF Grant No. CCF1942892 and a Wake Forest University Undergraduate Research and Creative Activities (URECA) Research Fellowship.
 
-This material is based upon work supported by the NSF Grant No. CCF1942892 and a Wake Forest University Undergraduate Research and Creative Activities (URECA) Research Fellowship.
-
-\vspace{-.175cm}
-
-\bibliographystyle{IEEEtran}
+\newpage
+\balance
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{refs}

### 2021-04-22 TF_op/VGG11.py
update vgg11

@@ -64,16 +64,16 @@ def _Fast_MatMul_grad(op, grad):
 
 
 class Fast_Linear(keras.layers.Layer):
-    def __init__(self, units=32, input_dim=32, activation='relu'):
+    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True):
         super(Fast_Linear, self).__init__()
         w_init = tf.random_normal_initializer()
         self.w = tf.Variable(
             initial_value=w_init(shape=(input_dim, units), dtype="float32"),
-            trainable=True,
+            trainable=trainable,
         )
         b_init = tf.zeros_initializer()
         self.b = tf.Variable(
-            initial_value=b_init(shape=(units,), dtype="float32"), trainable=True
+            initial_value=b_init(shape=(units,), dtype="float32"), trainable=trainable
         )
 
         self.epsilon = epsilon_values.get(mm_algo, 1e-2)
@@ -90,12 +90,14 @@ class Fast_Linear(keras.layers.Layer):
 
 
 def VGG11(
-    include_top=True,
-    weights='imagenet',
-    input_tensor=None,
-    input_shape=None,
-    pooling=None,
-    classes=1000,):
+        include_top=True,
+        weights='imagenet',
+        input_tensor=None,
+        input_shape=None,
+        pooling=None,
+        classes=1000,
+        fast_trainable=True
+        ):
     """Instantiates the VGG19 architecture.
       Reference:
       - [Very Deep Convolutional Networks for Large-Scale Image Recognition](
@@ -215,17 +217,17 @@ def VGG11(
     x = layers.Flatten(name='flatten')(x)
 
     #x = layers.Dense(4096, activation='relu', name='fc1')(x)
-    fast_layer1 = Fast_Linear(units=4096, input_dim=25088, activation='relu')
+    fast_layer1 = Fast_Linear(units=4096, input_dim=25088, activation='relu', trainable=fast_trainable)
     x = fast_layer1(x)
 
     #x = layers.Dense(4096, activation='relu', name='fc2')(x)
-    fast_layer2 = Fast_Linear(units=4096, input_dim=4096, activation='relu')
+    fast_layer2 = Fast_Linear(units=4096, input_dim=4096, activation='relu', trainable=fast_trainable)
     x = fast_layer2(x)
 
-    #imagenet_utils.validate_activation(classifier_activation, weights)
+    #imagenet_utils.validate_activation('softmax', weights)
 
-    #x = layers.Dense(classes, activation=classifier_activation, name='predictions')(x)
[…]

### 2021-04-22 latex/ICPP21/paper.tex

@@
-\documentclass[sigconf]{acmart}
+\documentclass[sigconf,review,anonymous]{acmart}
 
@@
 
-%\author{\IEEEauthorblockN{Grey Ballard}
-%\IEEEauthorblockA{\textit{Department of Computer Science} \\
-%\textit{Wake Forest University}\\
-%Winston-Salem, NC, USA \\
-%\email{ballard@wfu.edu}}
-%\and
-%\IEEEauthorblockN{Jack Weissenberger}
-%\IEEEauthorblockA{\textit{Department of Computer Science} \\
-%\textit{Wake Forest University}\\
-%Winston-Salem, NC, USA \\
-%\email{jack.weissenberger@gmail.com}}
-%\and
-%\IEEEauthorblockN{Luoping Zhang}
-%\IEEEauthorblockA{\textit{Department of Computer Science} \\
-%\textit{Wake Forest University}\\
-%Winston-Salem, NC, USA \\
-%\email{zhanl317@wfu.edu}}
-%}
+\author{\IEEEauthorblockN{Grey Ballard}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Wake Forest University}\\
+Winston-Salem, NC, USA \\
+\email{ballard@wfu.edu}}
+\and
+\IEEEauthorblockN{Jack Weissenberger}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Wake Forest University}\\
+Winston-Salem, NC, USA \\
+\email{jack.weissenberger@gmail.com}}
+\and
+\IEEEauthorblockN{Luoping Zhang}
+\IEEEauthorblockA{\textit{Department of Computer Science} \\
+\textit{Wake Forest University}\\
+Winston-Salem, NC, USA \\
+\email{zhanl317@wfu.edu}}
+}
 
@@
 In floating point arithmetic, however, there is a lower bound on the approximation error that depends on the working precision and properties of the algorithm.
-Thus, APA algorithms are often considered to have insufficient accuracy for most applications and have largely been overlooked as practical tools despite their performance potential and ability to outperform their exact counterparts \cite{BB15}.
+Thus, APA algorithms are often considered to have insufficient accuracy for most applications and have largely been overlooked as practical tools despite their performance potential and ability to outperform exact algorithms \cite{BB15}.
 
@@
 
-Our contribution is the use of APA matrix multiplication algorithms to address this problem and accelerate training.
+Our contribution is the use of APA matrix multiplication algorithms to address this problem.
 In particular we
@@
 \centering
-\caption{Properties of Arbitrary Precision Approximating Algorithms  \\ Speedup and Error are computed assuming 1 recursive step}
+\caption{Properties of APA algorithms. Speedup and error are computed assuming 1 recursive step}
 \label{tab:algs}
@@
 %
-\caption{Single- and multi-threaded square matrix multiplication performance, relative to $2n^3$ operations}
+\caption{Single- and multi-threaded square matrix multiplication performance, relative to $2n^3$ operations.  The machine peak with respect to classical algorithms is given as a dotted line.}
 \label{fig:matmul_perf}

### 2021-04-22 TF_op/vgg_fc.py
add vgg_fc

@@ -1,0 +1,112 @@
+import time
+import os
+
+import numpy as np
+
+from tensorflow.python.keras import backend
+from tensorflow.python.keras.applications import imagenet_utils
+from tensorflow.python.keras.engine import training
+from tensorflow.python.keras.layers import VersionAwareLayers
+from tensorflow.python.keras.utils import data_utils
+from tensorflow.python.keras.utils import layer_utils
+from tensorflow.python.lib.io import file_io
+from tensorflow.python.util.tf_export import keras_export
+from tensorflow.python.ops import array_ops, gen_math_ops, math_ops
+
+import tensorflow as tf
+from tensorflow import keras
+
+layers = VersionAwareLayers()
+
+
+@tf.RegisterGradient("FastMatMul")
+def _Fast_MatMul_grad(op, grad):
+    bt = array_ops.transpose(op.inputs[1])
+    at = array_ops.transpose(op.inputs[0])
+    grad_a = fast_mm_module.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_module.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    # a = math_ops.conj(op.inputs[0])
+    # b = math_ops.conj(op.inputs[1])
+    # grad_a = gen_math_ops.mat_mul(grad, b, transpose_b=True)
+    # grad_b = gen_math_ops.mat_mul(a, grad, transpose_a=True)
+    return grad_a, grad_b
+
+
+class Fast_Linear(keras.layers.Layer):
+    def __init__(self, units=32, input_dim=32, activation='relu'):
+        super(Fast_Linear, self).__init__()
+        w_init = tf.random_normal_initializer()
+        self.w = tf.Variable(
+            initial_value=w_init(shape=(input_dim, units), dtype="float32"),
+            trainable=True,
+        )
+        b_init = tf.zeros_initializer()
+        self.b = tf.Variable(
+            initial_value=b_init(shape=(units,), dtype="float32"), trainable=True
+        )
+
+        self.epsilon = epsilon_values.get(mm_algo, 1e-2)
+
+        self.activation = activation
+
+num_threads = 12
+mm_algo = 'smirnov444'
+epsilon_values = {
+    'bini322': 2**-11,
[…]

### 2021-04-22 TF_op/vgg_fc.py
update vgg

@@ -78,7 +78,7 @@ if __name__ == '__main__':
     model_input = layers.Input(shape=25088)
     # x = layers.Dense(4096, activation='relu', name='fc1')(x)
     fast_layer1 = Fast_Linear(units=4096, input_dim=25088, activation='relu')
-    x = fast_layer1()
+    x = fast_layer1(model_input)
 
     # x = layers.Dense(4096, activation='relu', name='fc2')(x)
     fast_layer2 = Fast_Linear(units=4096, input_dim=4096, activation='relu')

### 2021-04-22 TF_op/vgg_fc.py
update vgg_fc

@@ -33,21 +33,29 @@ def _Fast_MatMul_grad(op, grad):
 
 
 class Fast_Linear(keras.layers.Layer):
-    def __init__(self, units=32, input_dim=32, activation='relu'):
+    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True):
         super(Fast_Linear, self).__init__()
         w_init = tf.random_normal_initializer()
         self.w = tf.Variable(
             initial_value=w_init(shape=(input_dim, units), dtype="float32"),
-            trainable=True,
+            trainable=trainable,
         )
         b_init = tf.zeros_initializer()
         self.b = tf.Variable(
-            initial_value=b_init(shape=(units,), dtype="float32"), trainable=True
+            initial_value=b_init(shape=(units,), dtype="float32"), trainable=trainable
         )
 
         self.epsilon = epsilon_values.get(mm_algo, 1e-2)
 
         self.activation = activation
+
+    def call(self, inputs):
+        output = fast_mm_module.FastMatMul(a_matrix=inputs, b_matrix=self.w, epsilon=self.epsilon, steps=1,
+                                                    numthreads=num_threads) + self.b
+        if self.activation == 'softmax':
+            return tf.nn.softmax(output)
+        else:
+            return tf.nn.relu(output)
 
 num_threads = 12
 mm_algo = 'smirnov444'

### 2021-04-22 TF_op/VGG11.py
try random generator

@@ -245,7 +245,6 @@ def VGG11(
     return model
 
 
-
 class TimeHistory(Callback):
     def on_train_begin(self, logs={}):
         self.times = []
@@ -273,7 +272,9 @@ if __name__ == '__main__':
     epochs = 2
     batch_size = 256
 
-    y_train = tf.ones([batch_size])
+    g = tf.random.Generator()
+    y_train = g.normal(shape=(None, batch_size))
+    #y_train = tf.ones([batch_size])
     x_train = []
     for i in range(batch_size):
         x_train.append(image)
@@ -301,5 +302,8 @@ if __name__ == '__main__':
     b = time.time()
 
     print(f"Total time: {b-a} seconds, Time per epoch ({epochs}): {(b-a)/epochs}")
+
+
+    # Total time: 149.12401008605957 seconds, Time per epoch (2): 74.56200504302979
 
 

### 2021-04-22 TF_op/vgg_fc.py

@@ -118,3 +118,5 @@ if __name__ == '__main__':
     b = time.time()
 
     print(f"Total time: {b - a} seconds, Time per epoch ({epochs}): {(b - a) / epochs}")
+
+    # Total time: 6.033765554428101 seconds, Time per epoch (2): 3.0168827772140503

### 2021-04-22 TF_op/VGG11.py
update to random

@@ -272,8 +272,7 @@ if __name__ == '__main__':
     epochs = 2
     batch_size = 256
 
-    g = tf.random.Generator()
-    y_train = g.normal(shape=(None, batch_size))
+    y_train = tf.random.uniform(shape=(None, batch_size))
     #y_train = tf.ones([batch_size])
     x_train = []
     for i in range(batch_size):

### 2021-04-22 TF_op/vgg_fc.py

@@ -80,8 +80,8 @@ if __name__ == '__main__':
     epochs = 2
     batch_size = 256
 
-    y_train = tf.ones([batch_size])
-    x_train = tf.ones([batch_size, 25088])
+    y_train = tf.random.uniform(shape=[batch_size])
+    x_train = tf.random.uniform(shape=[batch_size, 25088])
 
     model_input = layers.Input(shape=25088)
     # x = layers.Dense(4096, activation='relu', name='fc1')(x)

### 2021-04-22 TF_op/VGG11.py
test

@@ -184,37 +184,38 @@ def VGG11(
     x = layers.Conv2D(
         64, (3, 3), activation='relu', padding='same', name='block1_conv1')(
         img_input)
-    x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block1_pool')(x)
-
-    # Block 2
-    x = layers.Conv2D(
-        128, (3, 3), activation='relu', padding='same', name='block2_conv1')(x)
-    x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block2_pool')(x)
-
-    # Block 3
-    x = layers.Conv2D(
-        256, (3, 3), activation='relu', padding='same', name='block3_conv1')(x)
-    x = layers.Conv2D(
-        256, (3, 3), activation='relu', padding='same', name='block3_conv2')(x)
-    x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block3_pool')(x)
-
-    # Block 4
-    x = layers.Conv2D(
-        512, (3, 3), activation='relu', padding='same', name='block4_conv1')(x)
-    x = layers.Conv2D(
-        512, (3, 3), activation='relu', padding='same', name='block4_conv2')(x)
-    x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block4_pool')(x)
-
-    # Block 5
-    x = layers.Conv2D(
-        512, (3, 3), activation='relu', padding='same', name='block5_conv1')(x)
-    x = layers.Conv2D(
-        512, (3, 3), activation='relu', padding='same', name='block5_conv2')(x)
-    x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block5_pool')(x)
+    # x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block1_pool')(x)
+    #
+    # # Block 2
+    # x = layers.Conv2D(
+    #     128, (3, 3), activation='relu', padding='same', name='block2_conv1')(x)
+    # x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block2_pool')(x)
+    #
+    # # Block 3
+    # x = layers.Conv2D(
+    #     256, (3, 3), activation='relu', padding='same', name='block3_conv1')(x)
+    # x = layers.Conv2D(
+    #     256, (3, 3), activation='relu', padding='same', name='block3_conv2')(x)
+    # x = layers.MaxPooling2D((2, 2), strides=(2, 2), name='block3_pool')(x)
+    #
+    # # Block 4
[…]

### 2021-04-22 TF_op/vgg_fc.py

@@ -113,7 +113,7 @@ if __name__ == '__main__':
     model.fit(x_train, y_train,
               epochs=epochs,
               batch_size=batch_size,
-              verbose=0)
+              )
 
     b = time.time()
 

### 2021-04-22 TF_op/VGG11.py
test

@@ -273,7 +273,7 @@ if __name__ == '__main__':
     epochs = 2
     batch_size = 256
 
-    y_train = tf.random.uniform(shape=(None, batch_size))
+    y_train = tf.random.uniform(shape=[batch_size])
     #y_train = tf.ones([batch_size])
     x_train = []
     for i in range(batch_size):

### 2021-04-22 TF_op/vgg_fc.py
442

@@ -58,7 +58,7 @@ class Fast_Linear(keras.layers.Layer):
             return tf.nn.relu(output)
 
 num_threads = 12
-mm_algo = 'smirnov444'
+mm_algo = 'smirnov442'
 epsilon_values = {
     'bini322': 2**-11,
     'schonhage333': 2**-5,
@@ -77,8 +77,8 @@ epsilon_values = {
 if __name__ == '__main__':
 
     fast_mm_module = tf.load_op_library(f'obj/{mm_algo}_mat_mul.so')
-    epochs = 2
-    batch_size = 256
+    epochs = 3
+    batch_size = 1000
 
     y_train = tf.random.uniform(shape=[batch_size])
     x_train = tf.random.uniform(shape=[batch_size, 25088])
@@ -117,6 +117,7 @@ if __name__ == '__main__':
 
     b = time.time()
 
-    print(f"Total time: {b - a} seconds, Time per epoch ({epochs}): {(b - a) / epochs}")
+    print(f"\n\nTotal time: {b - a} seconds, Time per epoch ({epochs}): {(b - a) / epochs}")
+    print(f"Batch size: {batch_size}, Algorithm: {mm_algo}")
 
-    # Total time: 6.033765554428101 seconds, Time per epoch (2): 3.0168827772140503
+    #Total time: 6.033765554428101 seconds, Time per epoch (2): 3.0168827772140503

### 2021-04-23 latex/ICPP21/paper.tex
add comments and new sections

@@
 Our performance results on a large MLP network show overall sequential and multithreaded performance improvements of up to 25\% and 13\%, respectively.
+\RZ{We also demonstrate that APA algorithms are efficiently implemented on a more complicated network VGG-11?}
 \end{abstract}
@@
 	\item curate a collection of both well-established and recently discovered practical APA algorithms;
-	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them, achieving up to 21\% performance improvement over the best parallel classical implementation,
-	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications; and
-	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication.
+	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them, achieving up to 21\% performance improvement over the best parallel classical implementation;
+	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications;
+	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication; and
+	\item \RZ{present performance improvements of VGG-11? network of ?\% over the classical matrix multiplication.}
 \end{enumerate}
@@
 
+\subsection{Skewed Matrix Multiplication}
+When computing square matrix multiplications, we have demonstrated that $\dims{4,4,2}$ and $\dims{4,4,4}$ algorithms have the best performance.
+However, algorithms with more skewed aspect ratios will likely perform better for problems with matching skewed matrix multiplication dimensions.
+We note that an algorithm for dimensions $\dims{m,n,k}$ can be translated into an algorithm for $\dims{n,m,k}$ and any other reordering of the dimensions \cite{BB15}.
+\RZ{transposing row major matrix problem?}
+\RZ{Best algorithm for each shape of matrices and its figure}
+
 \section{Application of APA Algorithms in Multi-Layer Perceptron Network Training}
@@
 
+\section{Application of APA algorithms in VGG-11}
+\subsection{VGG-11 Network}
+\RZ{description of the model of VGG-11 and VGG-19?}
+
+% \subsection{Accuracy}
+
+\subsection{Performance}
+\RZ{VGG-11 fully connected layer: BS x 25K x 4K / BS x 4K x 4K / BS x 4K x 1K? with plot}
+
 \section{Conclusion}

### 2021-04-23 TF_op/fast_vgg_fc.py
add fast vgg

@@ -1,0 +1,126 @@
+import time
+import os
+
+import numpy as np
+
+from tensorflow.python.keras import backend
+from tensorflow.python.keras.applications import imagenet_utils
+from tensorflow.python.keras.engine import training
+from tensorflow.python.keras.layers import VersionAwareLayers
+from tensorflow.python.keras.utils import data_utils
+from tensorflow.python.keras.utils import layer_utils
+from tensorflow.python.lib.io import file_io
+from tensorflow.python.util.tf_export import keras_export
+from tensorflow.python.ops import array_ops, gen_math_ops, math_ops
+
+import tensorflow as tf
+from tensorflow import keras
+
+layers = VersionAwareLayers()
+
+
+@tf.RegisterGradient("FastMatMul")
+def _Fast_MatMul_grad(op, grad):
+    bt = array_ops.transpose(op.inputs[1])
+    at = array_ops.transpose(op.inputs[0])
+    grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_424.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    # a = math_ops.conj(op.inputs[0])
+    # b = math_ops.conj(op.inputs[1])
+    # grad_a = gen_math_ops.mat_mul(grad, b, transpose_b=True)
+    # grad_b = gen_math_ops.mat_mul(a, grad, transpose_a=True)
+    return grad_a, grad_b
+
+
+class Fast_Linear(keras.layers.Layer):
+    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True, mm_module=object):
+        super(Fast_Linear, self).__init__()
+        w_init = tf.random_normal_initializer()
+        self.w = tf.Variable(
+            initial_value=w_init(shape=(input_dim, units), dtype="float32"),
+            trainable=trainable,
+        )
+        b_init = tf.zeros_initializer()
+        self.b = tf.Variable(
+            initial_value=b_init(shape=(units,), dtype="float32"), trainable=trainable
+        )
+
+        self.epsilon = epsilon_values.get(mm_algo, 1e-2)
+
+        self.activation = activation
+
+        self.fast_mm_module = mm_module
+
+    def call(self, inputs):
[…]

### 2021-04-23 TF_op/fast_vgg_fc.py
update fast vgg fc

@@ -33,7 +33,7 @@ def _Fast_MatMul_grad(op, grad):
 
 
 class Fast_Linear(keras.layers.Layer):
-    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True, mm_module=object):
+    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True, mm_module=object, mm_algo='smirnov444'):
         super(Fast_Linear, self).__init__()
         w_init = tf.random_normal_initializer()
         self.w = tf.Variable(
@@ -88,17 +88,17 @@ if __name__ == '__main__':
 
     model_input = layers.Input(shape=25088)
     # x = layers.Dense(4096, activation='relu', name='fc1')(x)
-    fast_layer1 = Fast_Linear(units=4096, input_dim=25088, activation='relu', mm_module=fast_mm_442)
+    fast_layer1 = Fast_Linear(units=4096, input_dim=25088, activation='relu', mm_module=fast_mm_442, mm_algo='smirnov442')
     x = fast_layer1(model_input)
 
     # x = layers.Dense(4096, activation='relu', name='fc2')(x)
-    fast_layer2 = Fast_Linear(units=4096, input_dim=4096, activation='relu', mm_module=fast_mm_442)
+    fast_layer2 = Fast_Linear(units=4096, input_dim=4096, activation='relu', mm_module=fast_mm_442, mm_algo='smirnov442')
     x = fast_layer2(x)
 
     # imagenet_utils.validate_activation('softmax', weights)
 
     # x = layers.Dense(classes, activation='softmax', name='predictions')(x)
-    fast_output_layer = Fast_Linear(units=1000, input_dim=4096, activation='softmax', mm_module=fast_mm_242)
+    fast_output_layer = Fast_Linear(units=1000, input_dim=4096, activation='softmax', mm_module=fast_mm_242, mm_algo='smirnov242')
     x = fast_output_layer(x)
 
     model = training.Model(model_input, x, name='FC')
@@ -121,6 +121,6 @@ if __name__ == '__main__':
     b = time.time()
 
     print(f"\n\nTotal time: {b - a} seconds, Time per epoch ({epochs}): {(b - a) / epochs}")
-    print(f"Batch size: {batch_size}, Algorithm: {mm_algo}")
+    print(f"Batch size: {batch_size}")
 
     #Total time: 6.033765554428101 seconds, Time per epoch (2): 3.0168827772140503

### 2021-04-23 TF_op/fast_vgg_fc.py
update fastvgg

@@ -19,7 +19,19 @@ from tensorflow import keras
 layers = VersionAwareLayers()
 
 
-@tf.RegisterGradient("FastMatMul")
+@tf.RegisterGradient("fast_mm_442")
+def _Fast_MatMul_grad(op, grad):
+    bt = array_ops.transpose(op.inputs[1])
+    at = array_ops.transpose(op.inputs[0])
+    grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_424.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    # a = math_ops.conj(op.inputs[0])
+    # b = math_ops.conj(op.inputs[1])
+    # grad_a = gen_math_ops.mat_mul(grad, b, transpose_b=True)
+    # grad_b = gen_math_ops.mat_mul(a, grad, transpose_a=True)
+    return grad_a, grad_b
+
+@tf.RegisterGradient("fast_mm_424")
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
@@ -33,7 +45,7 @@ def _Fast_MatMul_grad(op, grad):
 
 
 class Fast_Linear(keras.layers.Layer):
-    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True, mm_module=object, mm_algo='smirnov444'):
+    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True, mm_module=fast_mm_442, mm_algo='smirnov442'):
         super(Fast_Linear, self).__init__()
         w_init = tf.random_normal_initializer()
         self.w = tf.Variable(
@@ -52,7 +64,7 @@ class Fast_Linear(keras.layers.Layer):
         self.fast_mm_module = mm_module
 
     def call(self, inputs):
-        output = self.fast_mm_module.FastMatMul(a_matrix=inputs, b_matrix=self.w, epsilon=self.epsilon, steps=1,
+        output = self.fast_mm_module(a_matrix=inputs, b_matrix=self.w, epsilon=self.epsilon, steps=1,
                                                     numthreads=num_threads) + self.b
         if self.activation == 'softmax':
             return tf.nn.softmax(output)
[…]

### 2021-04-23 TF_op/fast_vgg_fc.py
update fast_vgg

@@ -24,19 +24,19 @@ def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
     grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_424.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_442.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
     # a = math_ops.conj(op.inputs[0])
     # b = math_ops.conj(op.inputs[1])
     # grad_a = gen_math_ops.mat_mul(grad, b, transpose_b=True)
     # grad_b = gen_math_ops.mat_mul(a, grad, transpose_a=True)
     return grad_a, grad_b
 
-@tf.RegisterGradient("fast_mm_424")
+@tf.RegisterGradient("fast_mm_242")
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
     grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_424.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_442.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
     # a = math_ops.conj(op.inputs[0])
     # b = math_ops.conj(op.inputs[1])
     # grad_a = gen_math_ops.mat_mul(grad, b, transpose_b=True)
@@ -45,7 +45,7 @@ def _Fast_MatMul_grad(op, grad):
 
 
 class Fast_Linear(keras.layers.Layer):
-    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True, mm_module=fast_mm_442, mm_algo='smirnov442'):
+    def __init__(self, units=32, input_dim=32, activation='relu', trainable=True, mm_module=object, mm_algo='smirnov442'):
[…]

### 2021-04-23 TF_op/fast_vgg_fc.py
test

@@ -19,19 +19,7 @@ from tensorflow import keras
 layers = VersionAwareLayers()
 
 
-@tf.RegisterGradient("fast_mm_442")
-def _Fast_MatMul_grad(op, grad):
-    bt = array_ops.transpose(op.inputs[1])
-    at = array_ops.transpose(op.inputs[0])
-    grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_442.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
-    # a = math_ops.conj(op.inputs[0])
-    # b = math_ops.conj(op.inputs[1])
-    # grad_a = gen_math_ops.mat_mul(grad, b, transpose_b=True)
-    # grad_b = gen_math_ops.mat_mul(a, grad, transpose_a=True)
-    return grad_a, grad_b
-
-@tf.RegisterGradient("fast_mm_242")
+@tf.RegisterGradient("FastMatMul")
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
@@ -89,10 +77,12 @@ epsilon_values = {
 
 if __name__ == '__main__':
 
-    fast_mm_442 = tf.load_op_library(f'obj/smirnov442_mat_mul.so')
-    fast_mm_442 = fast_mm_442.FastMatMul
     fast_mm_242 = tf.load_op_library(f'obj/smirnov242_mat_mul.so')
-    fast_mm_242 = fast_mm_242.FastMatMul
+    print(dir(fast_mm_242))
+
+    fast_mm_442 = tf.load_op_library(f'obj/smirnov442_mat_mul.so')
+    print(dir(fast_mm_442))
+
     #fast_mm_424 = tf.load_op_library(f'obj/smirnov424_mat_mul.so')
 
 

### 2021-04-23 TF_op/fast_vgg_fc.py
update op names

@@ -78,10 +78,8 @@ epsilon_values = {
 if __name__ == '__main__':
 
     fast_mm_242 = tf.load_op_library(f'obj/smirnov242_mat_mul.so')
-    print(dir(fast_mm_242))
 
     fast_mm_442 = tf.load_op_library(f'obj/smirnov442_mat_mul.so')
-    print(dir(fast_mm_442))
 
     #fast_mm_424 = tf.load_op_library(f'obj/smirnov424_mat_mul.so')
 

### 2021-04-23 TF_op/src/smirnov242_mat_mul.cc

@@ -11,7 +11,7 @@
 
 using namespace tensorflow;
 
-REGISTER_OP("FastMatMul")
+REGISTER_OP("FastMatMul242")
     .Attr("epsilon: float")
     .Attr("steps: int")
     .Attr("numthreads: int")

### 2021-04-23 TF_op/src/smirnov244_mat_mul.cc

@@ -11,7 +11,7 @@
 
 using namespace tensorflow;
 
-REGISTER_OP("FastMatMul")
+REGISTER_OP("FastMatMul244")
     .Attr("epsilon: float")
     .Attr("steps: int")
     .Attr("numthreads: int")

### 2021-04-23 TF_op/src/smirnov424_mat_mul.cc

@@ -11,7 +11,7 @@
 
 using namespace tensorflow;
 
-REGISTER_OP("FastMatMul")
+REGISTER_OP("FastMatMul424")
     .Attr("epsilon: float")
     .Attr("steps: int")
     .Attr("numthreads: int")

### 2021-04-23 TF_op/src/smirnov442_mat_mul.cc

@@ -11,7 +11,7 @@
 
 using namespace tensorflow;
 
-REGISTER_OP("FastMatMul")
+REGISTER_OP("FastMatMul442")
     .Attr("epsilon: float")
     .Attr("steps: int")
     .Attr("numthreads: int")

### 2021-04-23 TF_op/src/smirnov444_mat_mul.cc

@@ -11,7 +11,7 @@
 
 using namespace tensorflow;
 
-REGISTER_OP("FastMatMul")
+REGISTER_OP("FastMatMul444")
     .Attr("epsilon: float")
     .Attr("steps: int")
     .Attr("numthreads: int")

### 2021-04-23 TF_op/src/smirnov525_mat_mul.cc

@@ -11,7 +11,7 @@
 
 using namespace tensorflow;
 
-REGISTER_OP("FastMatMul")
+REGISTER_OP("FastMatMul525")
     .Attr("epsilon: float")
     .Attr("steps: int")
     .Attr("numthreads: int")

### 2021-04-23 TF_op/fast_vgg_fc.py
update fast_vgg with different opps

@@ -19,16 +19,33 @@ from tensorflow import keras
 layers = VersionAwareLayers()
 
 
-@tf.RegisterGradient("FastMatMul")
+# output to second FC
+@tf.RegisterGradient("FastMatMul242")
+def _Fast_MatMul_grad(op, grad):
+    bt = array_ops.transpose(op.inputs[1])
+    at = array_ops.transpose(op.inputs[0])
+    grad_a = fast_mm_444.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_244.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    return grad_a, grad_b
+
+
+# second FC to first
+@tf.RegisterGradient("FastMatMul442")
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
     grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_442.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
-    # a = math_ops.conj(op.inputs[0])
-    # b = math_ops.conj(op.inputs[1])
-    # grad_a = gen_math_ops.mat_mul(grad, b, transpose_b=True)
-    # grad_b = gen_math_ops.mat_mul(a, grad, transpose_a=True)
+    grad_b = fast_mm_424.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    return grad_a, grad_b
+
+
+# second FC to first
+@tf.RegisterGradient("FastMatMul525")
+def _Fast_MatMul_grad(op, grad):
+    bt = array_ops.transpose(op.inputs[1])
+    at = array_ops.transpose(op.inputs[0])
+    grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_525.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
[…]

### 2021-04-23 TF_op/fast_vgg_fc.py
update

@@ -95,11 +95,17 @@ epsilon_values = {
 if __name__ == '__main__':
 
     fast_mm_242 = tf.load_op_library('obj/smirnov242_mat_mul.so')
+    fast_mm_242 = fast_mm_242.FastMatMul242
     fast_mm_442 = tf.load_op_library('obj/smirnov442_mat_mul.so')
+    fast_mm_442 = fast_mm_442.FastMatMul442
     fast_mm_525 = tf.load_op_library('obj/smirnov525_mat_mul.so')
+    fast_mm_525 = fast_mm_525.FastMatMul525
     fast_mm_424 = tf.load_op_library('obj/smirnov424_mat_mul.so')
+    fast_mm_424 = fast_mm_424.FastMatMul424
     fast_mm_444 = tf.load_op_library('obj/smirnov444_mat_mul.so')
+    fast_mm_444 = fast_mm_444.FastMatMul444
     fast_mm_244 = tf.load_op_library('obj/smirnov244_mat_mul.so')
+    fast_mm_244 = fast_mm_244.FastMatMul244
 
 
 

### 2021-04-23 TF_op/fast_vgg_fc.py
update gradients

@@ -24,8 +24,8 @@ layers = VersionAwareLayers()
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
-    grad_a = fast_mm_444.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_244.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_a = fast_mm_444.FastMatMul444(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_244.FastMatMul244(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
     return grad_a, grad_b
 
 
@@ -34,8 +34,8 @@ def _Fast_MatMul_grad(op, grad):
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
-    grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_424.FastMatMul(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_a = fast_mm_442.FastMatMul442(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_424.FastMatMul424(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
     return grad_a, grad_b
 
 
@@ -44,8 +44,8 @@ def _Fast_MatMul_grad(op, grad):
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
-    grad_a = fast_mm_442.FastMatMul(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
[…]

### 2021-04-23 TF_op/fast_vgg_fc.py
update grad

@@ -24,8 +24,8 @@ layers = VersionAwareLayers()
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
-    grad_a = fast_mm_444.FastMatMul444(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_244.FastMatMul244(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_a = fast_mm_444(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_244(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
     return grad_a, grad_b
 
 
@@ -34,8 +34,8 @@ def _Fast_MatMul_grad(op, grad):
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
-    grad_a = fast_mm_442.FastMatMul442(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_424.FastMatMul424(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_a = fast_mm_442(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
+    grad_b = fast_mm_424(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
     return grad_a, grad_b
 
 
@@ -44,8 +44,8 @@ def _Fast_MatMul_grad(op, grad):
 def _Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
-    grad_a = fast_mm_442.FastMatMul442(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
-    grad_b = fast_mm_525.FastMatMul525(a_matrix=at, b_matrix=grad, epsilon=1e-2, steps=1, numthreads=num_threads)
[…]

### 2021-04-23 TF_op/fast_vgg_fc.py
update function names

@@ -21,7 +21,7 @@ layers = VersionAwareLayers()
 
 # output to second FC
 @tf.RegisterGradient("FastMatMul242")
-def _Fast_MatMul_grad(op, grad):
+def first_Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
     grad_a = fast_mm_444(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
@@ -31,7 +31,7 @@ def _Fast_MatMul_grad(op, grad):
 
 # second FC to first
 @tf.RegisterGradient("FastMatMul442")
-def _Fast_MatMul_grad(op, grad):
+def second_Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
     grad_a = fast_mm_442(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)
@@ -41,7 +41,7 @@ def _Fast_MatMul_grad(op, grad):
 
 # second FC to first
 @tf.RegisterGradient("FastMatMul525")
-def _Fast_MatMul_grad(op, grad):
+def thrid_Fast_MatMul_grad(op, grad):
     bt = array_ops.transpose(op.inputs[1])
     at = array_ops.transpose(op.inputs[0])
     grad_a = fast_mm_442(a_matrix=grad, b_matrix=bt, epsilon=1e-2, steps=1, numthreads=num_threads)

### 2021-04-23 TF_op/src/smirnov242_mat_mul.cc
update register opps

@@ -99,4 +99,4 @@ public:
 
     };
 
-REGISTER_KERNEL_BUILDER(Name("FastMatMul").Device(DEVICE_CPU), FastMatMulOp);
+REGISTER_KERNEL_BUILDER(Name("FastMatMul242").Device(DEVICE_CPU), FastMatMulOp);

### 2021-04-23 TF_op/src/smirnov244_mat_mul.cc

@@ -99,4 +99,4 @@ public:
 
     };
 
-REGISTER_KERNEL_BUILDER(Name("FastMatMul").Device(DEVICE_CPU), FastMatMulOp);
+REGISTER_KERNEL_BUILDER(Name("FastMatMul244").Device(DEVICE_CPU), FastMatMulOp);

### 2021-04-23 TF_op/src/smirnov424_mat_mul.cc

@@ -99,4 +99,4 @@ public:
 
     };
 
-REGISTER_KERNEL_BUILDER(Name("FastMatMul").Device(DEVICE_CPU), FastMatMulOp);
+REGISTER_KERNEL_BUILDER(Name("FastMatMul424").Device(DEVICE_CPU), FastMatMulOp);

### 2021-04-23 TF_op/src/smirnov442_mat_mul.cc

@@ -99,4 +99,4 @@ public:
 
     };
 
-REGISTER_KERNEL_BUILDER(Name("FastMatMul").Device(DEVICE_CPU), FastMatMulOp);
+REGISTER_KERNEL_BUILDER(Name("FastMatMul442").Device(DEVICE_CPU), FastMatMulOp);

### 2021-04-23 TF_op/src/smirnov444_mat_mul.cc

@@ -99,4 +99,4 @@ public:
 
     };
 
-REGISTER_KERNEL_BUILDER(Name("FastMatMul").Device(DEVICE_CPU), FastMatMulOp);
+REGISTER_KERNEL_BUILDER(Name("FastMatMul444").Device(DEVICE_CPU), FastMatMulOp);

### 2021-04-23 TF_op/src/smirnov525_mat_mul.cc

@@ -99,4 +99,4 @@ public:
 
     };
 
-REGISTER_KERNEL_BUILDER(Name("FastMatMul").Device(DEVICE_CPU), FastMatMulOp);
+REGISTER_KERNEL_BUILDER(Name("FastMatMul525").Device(DEVICE_CPU), FastMatMulOp);

### 2021-04-24 latex/ICPP21/paper.tex
add plot for VGG fast combo time data
update paper and vgg plot

@@
 \RZ{Best algorithm for each shape of matrices and its figure}
+In order to achieve the best performance on a neural network model VGG-11, which we will apply these APA algorithms on, we decided to measure the performance all 28 algorithms on the shapes of multiplied matrices involved in the fully connected layers of VGG-11.
+Each fully connected layers of VGG-11 contains the number of weights of 4096, 4096, and 1000, and the input dimension of the first layer is 25088.
+Therefore, there are 9 in total multiplications involved: $\dims{25088, 4096, bs}$.
 
@@
 \section{Application of APA algorithms in VGG-11}
+
+\begin{figure}[hb!]
+\centering
+\input{plots/VGG_time}
+\caption{Total time test on VGG-11 with 6 threads}
+\label{fig:vgg}
+\end{figure}
+
 \subsection{VGG-11 Network}
@@
 \RZ{VGG-11 fully connected layer: BS x 25K x 4K / BS x 4K x 4K / BS x 4K x 1K? with plot}
+\cref{fig:vgg}
+\RZ{Fast Combo: 525, 442, 242}
 

### 2021-04-24 latex/ICPP21/paper.tex
submitted to ICPP (version without new data)
update paper

@@
 Each fully connected layers of VGG-11 contains the number of weights of 4096, 4096, and 1000, and the input dimension of the first layer is 25088.
-Therefore, there are 9 in total multiplications involved: $\dims{25088, 4096, bs}$.
+Therefore, there are 9 in total multiplications involved: $\dims{bs, 25088, 4096}$, $\dims{bs, 4096, 25088}$, $\dims{25088, bs, 4096}$, $\dims{bs, 4096, 4096}$, $\dims{bs, 4096, 4096}$, $\dims{4096, dims, 4096}$, $\dims{bs, 4096, 1000}$, $\dims{bs, 1000, 4096}$, and $\dims{4096, bs, 1000}$.
+($bs$ stands for the batch size of the input.)
+
+The performance results for 6 threads shows that algorithm $\dims{4,4,2}$ and its variances are still the dominant one among all 12 APA algorithms and their variances.
 

### 2021-04-24 latex/ICPP21/paper.tex
update paper on vgg figures

@@
 	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication; and
-	\item \RZ{present performance improvements of VGG-11? network of ?\% over the classical matrix multiplication.}
+	\item \RZ{present performance improvements of VGG-19 network of 15\% over the classical matrix multiplication.}
 \end{enumerate}
@@
 \centering
-\caption{Properties of APA algorithms. Speedup and error are computed assuming 1 recursive step}
-\label{tab:algs}
 \begin{tabular}{| c | c c c | c c c |} 
@@
 \end{tabular}
+\caption{Properties of APA algorithms. Speedup and error are computed assuming 1 recursive step}
+\label{tab:algs}
 \end{table}
@@
 
-\section{Application of APA algorithms in VGG-11}
+\section{Application of APA algorithms in VGG-19}
 
-\begin{figure}[hb!]
+\begin{figure*}
+%
+\subfloat[One thread]{
+\centering
+\input{plots/VGG_time_seq}
+\label{fig:vgg_seq}}
+	%
+\hfill	
+\subfloat[Six threads]{
 \centering
 \input{plots/VGG_time}
-\caption{Total time test on VGG-11 with 6 threads}
+\label{fig:vgg_par6}}
+%
+\caption{Total time test on the fully connected layes on VGG-19 using $\dims{4,4,2}$ and dgemm}
 \label{fig:vgg}
-\end{figure}
+\end{figure*}
 
-\subsection{VGG-11 Network}
-\RZ{description of the model of VGG-11 and VGG-19?}
+\subsection{VGG-19 Network}
+VGG19 is a convolutional neural network typically trained on the imagenet dataset as a classification task. The model is one of the highest performing algorithms for this task. 
+The network is composed of 19 layers, 16 convolutional and 3 fully connected. 
+The 3 fully connected layers consist of 25,088, 4096 and 1000 nodes and are placed at the end of the network. 
+These create very large multiplications in forwards and backwards propagation. 
+Using faster APA algorithms for the multiplication we are able to speed up the training of the network.
 
@@
 \subsection{Performance}
-\RZ{VGG-11 fully connected layer: BS x 25K x 4K / BS x 4K x 4K / BS x 4K x 1K? with plot}
-\cref{fig:vgg}
-\RZ{Fast Combo: 525, 442, 242}
+
+Replacing the classic matrix multiplication algorithm in these layers with the Smirnov 442 algorithm provides up to a 10\% speed up of the fully connected layers during training with 6 threads and a 15\% speed up using a single thread.
+\cref{fig:vgg} plot the training time of each network with increasing batch sizes
 
@@
 While the accuracy and performance results are promising for MLP networks and the MNIST dataset, we believe that APA algorithms can also be effective in state-of-the-art networks and more difficult machine learning problems.
-For example, the VGG-19 network for large-scale visual recognition \cite{SZ15} is based on a deep convolutional network and is bottlenecked by large fully connected layers, including a layer with dimensions $25088\times 4096$, where APA algorithms can offer significant performance improvement.
+For example, the VGG-19 network for large-scale visual recognition \cite{SZ15} is based on a deep convolutional network and is bottlenecked by large fully connected layers.
+Our result are able to show a 15\% performance improvement using the sequential algorithm, and a 10\% improvement using parallel algorithm.
+This highlights the practical applications on training large networks.
 
@@
 
-\newpage
 \balance

### 2021-04-24 latex/ICPP21/paper.tex
update figures in paper
submitted version to ICPP with VGG 442 data

@@
 
+\setcopyright{rightsretained}
+
 \acmConference[ICPP'21]{International Conference on Parallel Processing}{August 9--12}{Chicago, IL, USA}
@@
 Our performance results on a large MLP network show overall sequential and multithreaded performance improvements of up to 25\% and 13\%, respectively.
-\RZ{We also demonstrate that APA algorithms are efficiently implemented on a more complicated network VGG-11?}
+We also demonstrate up to 15\% improvement when training the fully connected layers of the VGG-19 image classification network.
 \end{abstract}
@@
 	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications;
-	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication; and
-	\item \RZ{present performance improvements of VGG-19 network of 15\% over the classical matrix multiplication.}
+	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the classical algorithm; and
+	\item show performance improvements for training the fully connected layers of the VGG-19 network of up to 15\%.
 \end{enumerate}
@@
 
+\begin{comment}
 \subsection{Skewed Matrix Multiplication}
@@
 The performance results for 6 threads shows that algorithm $\dims{4,4,2}$ and its variances are still the dominant one among all 12 APA algorithms and their variances.
+\end{comment}
 
@@
 
-\begin{figure*}
+\begin{figure}
 %
@@
 \input{plots/VGG_time_seq}
-\label{fig:vgg_seq}}
-	%
+\label{fig:vgg_seq}}%
 \hfill	
@@
 %
-\caption{Total time test on the fully connected layes on VGG-19 using $\dims{4,4,2}$ and dgemm}
+\caption{Per-batch training time on the fully connected layers of VGG-19 using $\dims{4,4,2}$ and Classical}
 \label{fig:vgg}
-\end{figure*}
+\end{figure}
 
-\subsection{VGG-19 Network}
-VGG19 is a convolutional neural network typically trained on the imagenet dataset as a classification task. The model is one of the highest performing algorithms for this task. 
-The network is composed of 19 layers, 16 convolutional and 3 fully connected. 
-The 3 fully connected layers consist of 25,088, 4096 and 1000 nodes and are placed at the end of the network. 
-These create very large multiplications in forwards and backwards propagation. 
-Using faster APA algorithms for the multiplication we are able to speed up the training of the network.
+%\subsection{VGG-19 Network}
+VGG-19 \cite{SZ15} is a convolutional neural network typically trained on the ImageNet dataset as a classification task. 
+The model is one of the highest performing algorithms for this task, and the network is composed of 19 layers, 16 convolutional and 3 fully connected. 
+The 3 fully connected layers consist of 25{,}088, 4096, and 1000 nodes, creating very large matrix multiplications in forward and backward propagation. 
+By using faster APA algorithms for the multiplication, we are able to speed up the training of the network.
 
-% \subsection{Accuracy}
+%\subsection{Performance}
 
-\subsection{Performance}
-
-Replacing the classic matrix multiplication algorithm in these layers with the Smirnov 442 algorithm provides up to a 10\% speed up of the fully connected layers during training with 6 threads and a 15\% speed up using a single thread.
-\cref{fig:vgg} plot the training time of each network with increasing batch sizes
+Replacing the classic matrix multiplication algorithm in these layers with the $\dims{4,4,2}$ algorithm provides up to a 10\% speed up of the fully connected layers during training with 6 threads and a 15\% speed up using a single thread.
+\Cref{fig:vgg} shows the per-batch training time of the fully connected layers across batch sizes.
 
@@
 We note that an algorithm for dimensions $\dims{m,n,k}$ can be translated into an algorithm for $\dims{n,m,k}$ and any other reordering of the dimensions \cite{BB15}.
-For very large matrices, it may also be beneficial to consider more than one recursive step of the same algorithm or a combination of two or three different algorithms across recursive steps (uniform, non-stationary algorithms \cite{BBDLS16}).
+For very large matrices, it may be worth considering more than one recursive step of the same algorithm or a combination of two or three different algorithms across recursive steps (uniform, non-stationary algorithms \cite{BBDLS16}).
 

### 2021-07-02 latex/ICPP21/paper.tex
some small edits in preparation for camera ready

@@
-\documentclass[sigconf,review,anonymous]{acmart}
+\documentclass[sigconf]{acmart}
 
@@
 \usetikzlibrary{positioning}
+\usetikzlibrary{external} 
+\tikzexternalize[optimize=false, prefix=tikz/]
 
@@
 
-\author{\IEEEauthorblockN{Grey Ballard}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Wake Forest University}\\
-Winston-Salem, NC, USA \\
-\email{ballard@wfu.edu}}
-\and
-\IEEEauthorblockN{Jack Weissenberger}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Wake Forest University}\\
-Winston-Salem, NC, USA \\
-\email{jack.weissenberger@gmail.com}}
-\and
-\IEEEauthorblockN{Luoping Zhang}
-\IEEEauthorblockA{\textit{Department of Computer Science} \\
-\textit{Wake Forest University}\\
-Winston-Salem, NC, USA \\
-\email{zhanl317@wfu.edu}}
-}
+\author{Grey Ballard}
+\affiliation{%
+   \institution{Wake Forest University}
+   \city{Winston-Salem, NC}
+   \country{USA}}
+\email{ballard@wfu.edu}
+
+\author{Jack Weissenberger}
+\affiliation{%
+   \institution{Wake Forest University}
+   \city{Winston-Salem, NC}
+   \country{USA}}
+\email{jack.weissenberger@gmail.com}
+
+\author{Luoping Zhang}
+\affiliation{%
+   \institution{Wake Forest University}
+   \city{Winston-Salem, NC}
+   \country{USA}}
+\email{zhanl317@wfu.edu}
 
@@
 
-Nearly all fast matrix multiplication algorithms are based a rule for multiplying matrices of fixed size, and the reduction in asymptotic complexity stems from using the rule recursively on general matrices.
+Nearly all fast matrix multiplication algorithms are based on a rule for multiplying matrices of fixed size, and the reduction in asymptotic complexity stems from using the rule recursively on general matrices.
 For example, Strassen's algorithm is specified by a rule for multiplying two $2\times2$ matrices (denoted $\dims{2,2,2}$) using 7 multiplications instead of the classical algorithm's 8 multiplications.
@@
 \end{tabular}
-\caption{Properties of APA algorithms. Speedup and error are computed assuming 1 recursive step}
+\caption{Properties of APA algorithms. Speedup and error are computed assuming 1 recursive step.}
 \label{tab:algs}
@@
 We show empirical error results for uniform random inputs of varying dimension in \cref{fig:accuracy}, as compared to the classical algorithm.
-We measure the relative Frobenius norm error, or $\|C-\hat C\|_F/\|C\|$, where $\hat C$ is computed by each algorithm and $C$ is computed using the classical algorithm in double precision.
+We measure the relative Frobenius norm error, or $\|C-\hat C\|_F/\|C\|_F$, where $\hat C$ is computed by each algorithm and $C$ is computed using the classical algorithm in double precision.
 In order to choose the optimal $\lambda$ value for each algorithm, we tested the 5 powers of 2 closest to the theoretical optimal value and chose the best.
@@
 
-%\paragraph*{Acknowledgments}
-%
-%This material is based upon work supported by the NSF Grant No. CCF1942892 and a Wake Forest University Undergraduate Research and Creative Activities (URECA) Research Fellowship.
+\paragraph*{Acknowledgments}
+
+This material is based upon work supported by the NSF Grant No. 1942892 and a Wake Forest University Undergraduate Research and Creative Activities (URECA) Research Fellowship.
 

### 2021-07-05 latex/ICPP21/paper.tex
add ACM copyright info

@@
 
-\acmConference[ICPP'21]{International Conference on Parallel Processing}{August 9--12}{Chicago, IL, USA}
+\copyrightyear{2021} 
+\acmYear{2021} 
+\setcopyright{acmcopyright}\acmConference[ICPP Workshops '21]{50th International Conference on Parallel Processing Workshop}{August 9--12, 2021}{Lemont, IL, USA}
+\acmBooktitle{50th International Conference on Parallel Processing Workshop (ICPP Workshops '21), August 9--12, 2021, Lemont, IL, USA}
+\acmPrice{15.00}
+\acmDOI{10.1145/3458744.3474050}
+\acmISBN{978-1-4503-8441-4/21/08}
 
-\title{Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms
-%\thanks{Identify applicable funding agency here. If none, delete this.}
-}
+\title{Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms}
 
@@
 
-This material is based upon work supported by the NSF Grant No. 1942892 and a Wake Forest University Undergraduate Research and Creative Activities (URECA) Research Fellowship.
+This material is based upon work supported by the National Science Foundation Grant No. 1942892 and a Wake Forest University Undergraduate Research and Creative Activities (URECA) Research Fellowship.
 

### 2021-07-06 latex/ICPP21/paper.tex
converted input of tikz files to including pdf versions
add CCS concepts and keywords, converted away from cleveref

@@
 \usepackage{hyperref}
-\usepackage[nameinlink]{cleveref}
-\crefname{section}{\S}{Sections}
-\crefname{figure}{Fig.}{Figs.}
-\Crefname{figure}{Figure}{Figures}
-\crefname{table}{Table}{Tables}
+%\usepackage[nameinlink]{cleveref}
+%\crefname{section}{\S}{Sections}
+%\crefname{figure}{Fig.}{Figs.}
+%\Crefname{figure}{Figure}{Figures}
+%\crefname{table}{Table}{Tables}
 
 % tikz
-\usepackage{tikz,pgfplots}
-\usetikzlibrary{positioning}
-\usetikzlibrary{external} 
-\tikzexternalize[optimize=false, prefix=tikz/]
+%\usepackage{tikz,pgfplots}
+%\usetikzlibrary{positioning}
+%\usetikzlibrary{external} 
+%\tikzexternalize[optimize=false, prefix=tikz/]
 
@@
 \acmISBN{978-1-4503-8441-4/21/08}
+
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10002950.10003705.10011686</concept_id>
+       <concept_desc>Mathematics of computing~Mathematical software performance</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+  <concept>
+       <concept_id>10010147.10010257.10010293.10010294</concept_id>
+       <concept_desc>Computing methodologies~Neural networks</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Mathematics of computing~Mathematical software performance}
+\ccsdesc[500]{Computing methodologies~Neural networks}
+
+\keywords{neural networks, multilayer perceptrons, APA matrix multiplication}
 
@@
 
-One of the reasons such algorithms are impractical is that they are based on so-called \emph{Arbitrary Precision Approximating} (APA) algorithms, which we describe in detail in \cref{sec:APA}.
+One of the reasons such algorithms are impractical is that they are based on so-called \emph{Arbitrary Precision Approximating} (APA) algorithms, which we describe in detail in \S~\ref{sec:APA}.
 In exact arithmetic, these algorithms compute an approximation of the correct result, where the error is polynomial in a nonzero parameter of the algorithm \cite{BLR80}.
@@
 In floating point arithmetic, choosing too small a value for $\lambda$ leads to accumulation of roundoff error that exceeds the approximation error.
-We discuss optimizing $\lambda$ in \cref{sec:APAerr}.
+We discuss optimizing $\lambda$ in \S~\ref{sec:APAerr}.
 
@@
 For example, this rule computes the first entry of the output matrix as $\hat C_{11} = A_{11}  B_{11} + A_{12}  B_{21} - \lambda A_{12}  B_{11}$.
-Thus, in this case, the first entry of the error matrix $E$ in \cref{eq:APAapprox} is $E_{11} = A_{12}  B_{11}$.
+Thus, in this case, the first entry of the error matrix $E$ in eq.~\eqref{eq:APAapprox} is $E_{11} = A_{12}  B_{11}$.
 
@@
 The first parameter, $\sigma$, is the smallest positive exponent of the error polynomial and represents the approximation error.
-\Cref{eq:APAapprox} shows the error as a polynomial of $\lambda$ whose leading term is linear in $\lambda$.
-If an algorithm satisfies \cref{eq:APAapprox} with $E\neq 0$, then $\sigma=1$.
+Equation~\eqref{eq:APAapprox} shows the error as a polynomial of $\lambda$ whose leading term is linear in $\lambda$.
+If an algorithm satisfies eq.~\eqref{eq:APAapprox} with $E\neq 0$, then $\sigma=1$.
 However, if $E=0$, then $\sigma>1$ is the degree of the leading monomial.
@@
 This parameter represents the effect of roundoff error caused by floating point arithmetic involving the largest intermediate values computed by the algorithm.
-For example, the triplet given in \cref{eq:triplet} yields a sum of negative exponents of $0+0+1=1$, and in the case of Bini's algorithm, no other triplet has a larger sum, so $\varphi=1$ for that algorithm.
+For example, the triplet given in eq.~\eqref{eq:triplet} yields a sum of negative exponents of $0+0+1=1$, and in the case of Bini's algorithm, no other triplet has a larger sum, so $\varphi=1$ for that algorithm.
 Smaller $\varphi$ implies smaller error due to roundoff, and the APA algorithms we consider exhibit a range of values.
@@
 Note that if multiple recursive steps are used, then $\varphi$ increases proportional to the number of steps, so straightforward optimization of $\lambda$ results in error of $O(2^{-d\sigma/(\sigma+s\varphi)})$ for $s$ recursive steps.
-The parameters and minimum error for the algorithms we consider are presented in \cref{sec:APAprops}.
+The parameters and minimum error for the algorithms we consider are presented in \S~\ref{sec:APAprops}.
 
-We show empirical error results for uniform random inputs of varying dimension in \cref{fig:accuracy}, as compared to the classical algorithm.
+We show empirical error results for uniform random inputs of varying dimension in Fig.~\ref{fig:accuracy}, as compared to the classical algorithm.
 We measure the relative Frobenius norm error, or $\|C-\hat C\|_F/\|C\|_F$, where $\hat C$ is computed by each algorithm and $C$ is computed using the classical algorithm in double precision.
@@
 \renewcommand{\datafile}{data/matmul_acc.dat}
-\input{plots/accuracy}
+%\input{plots/accuracy}
+\includegraphics{tikz-taps/paper-figure0}
 \caption{Relative Frobenius norm error for APA algorithms on random inputs.}
@@
 While $\phi$ is computed using the exponents of the largest intermediate term of $\lambda$, the leading term of $\dims{5,5,5}$ has a constant of $1/4$, lessening its magnitude. 
-We study the effects of the matrix multiplication error on neural network training and test accuracy in \cref{sec:NNacc}.
+We study the effects of the matrix multiplication error on neural network training and test accuracy in \S~\ref{sec:NNacc}.
 
@@
 
-\Cref{tab:algs} shows the key performance and accuracy properties of the APA algorithms we consider.
+Table~\ref{tab:algs} shows the key performance and accuracy properties of the APA algorithms we consider.
 Each row corresponds to an algorithm, and the first row includes the classical algorithm for comparison.
@@
 \centering
-\input{figs/hybrid}
+%\input{figs/hybrid}
+\includegraphics{tikz-taps/paper-figure1.pdf}
 \caption{Illustration of hybrid parallelization strategy for $r=10$ and 4 threads.
@@
 \renewcommand{\threads}{1}
-\input{plots/matmul_perf}
+%\input{plots/matmul_perf}
+\includegraphics{tikz-taps/paper-figure2.pdf}
 %\caption{Single-threaded matrix multiplication performance}
@@
 \renewcommand{\threads}{6}
-\input{plots/matmul_perf}
+%\input{plots/matmul_perf}
+\includegraphics{tikz-taps/paper-figure3.pdf}
 %\caption{Multithreaded matrix multiplication performance (6 threads)}
@@
 \renewcommand{\threads}{12}
-\input{plots/matmul_perf}
+%\input{plots/matmul_perf}
+\includegraphics{tikz-taps/paper-figure4.pdf}
 %\caption{Multithreaded matrix multiplication performance (6 threads)}
@@
 
-As described in \cref{sec:APAalgs}, each algorithm can be encoded by a set of triplets of coefficient matrices.
+As described in \S~\ref{sec:APAalgs}, each algorithm can be encoded by a set of triplets of coefficient matrices.
 From this representation, we generate recursive code that computes the linear combinations for the inputs to each multiplication and combines the outputs.
@@
 The multiplications are parallelized as follows: given $r$ multiplications and $p$ threads, with $r=p\cdot q + \ell$ for integers $q$ and $\ell<p$, we assign each thread $q$ multiplications to be performed independently, and the remaining $\ell$ multiplications are performed by all threads using the multithreaded implementation of \texttt{gemm}.
-\Cref{fig:hybrid} shows the hybrid strategy for $r=10$ (Bini's algorithm) and $p=4$.
+Figure~\ref{fig:hybrid} shows the hybrid strategy for $r=10$ (Bini's algorithm) and $p=4$.
 
@@
 
-\Cref{fig:matmul_seq} reports the sequential performance of the APA algorithms in comparison to the most efficient implementation of the classical algorithm, MKL's single-precision \texttt{gemm}.
+Figure~\ref{fig:matmul_seq} reports the sequential performance of the APA algorithms in comparison to the most efficient implementation of the classical algorithm, MKL's single-precision \texttt{gemm}.
 The y-axis is the \emph{effective} GFLOPS, which is measured as $\num{1e-9} \cdot 2n^3 / \text{ time}$.
@@
 
-We report parallel performance in \cref{fig:matmul_par_6,fig:matmul_par_12} for running with 6 threads (one socket) and 12 threads (both sockets).
+We report parallel performance in Figs.~\ref{fig:matmul_par_6} and \ref{fig:matmul_par_12} for running with 6 threads (one socket) and 12 threads (both sockets).
 Overall, speedups of APA algorithms over classical are reduced in the parallel case.
@@
 
-In the performance results for 6 threads (\cref{fig:matmul_par_6}), we see that many of the algorithms start to outperform classical around dimension 2000, though some poor performance is observed for algorithms and particular matrix dimensions.
+In the performance results for 6 threads (Fig.~\ref{fig:matmul_par_6}), we see that many of the algorithms start to outperform classical around dimension 2000, though some poor performance is observed for algorithms and particular matrix dimensions.
 The fastest algorithms (e.g., $\dims{4,4,4}$ and $\dims{4,4,2}$) achieve a speedup of up to 25\% over \texttt{gemm} and exceed the machine peak for classical algorithms.
@@
 
-\Cref{fig:matmul_par_12} shows the results for 12 threads.
+Figure~\ref{fig:matmul_par_12} shows the results for 12 threads.
 Here we see a majority of the algorithms are slower than the classical algorithm, even for large matrices.
@@
 \renewcommand{\set}{train}
-\input{plots/NN_acc}
+%\input{plots/NN_acc}
+\includegraphics{tikz-taps/paper-figure5.pdf}
 \label{fig:NN_acc_train}}
@@
 \renewcommand{\set}{test}
-\input{plots/NN_acc}
+%\input{plots/NN_acc}
+\includegraphics{tikz-taps/paper-figure6.pdf}
 \label{fig:NN_acc_test}}
@@
 \centering
-\input{figs/neuralnet}
+%\input{figs/neuralnet}
+\includegraphics{tikz-taps/paper-figure7.pdf}
 \caption{Multi-Layer Perceptron network structure}
@@
 
-To illustrate the effectiveness of APA algorithms in neural network training, we consider Multi-Layer Perceptron networks, as shown in \cref{fig:neuralnet}.
+To illustrate the effectiveness of APA algorithms in neural network training, we consider Multi-Layer Perceptron networks, as shown in Fig.~\ref{fig:neuralnet}.
 We use TensorFlow 2.2.0 to build the networks and implement custom operators for matrix multiplication.
@@
 
-We show an MLP network with two hidden layers in \cref{fig:neuralnet}.
-The dimensions in the figure match those of the accuracy experiment described in \cref{sec:NNacc}.
-For the performance experiment of \cref{sec:NNperf}, we use an MLP with four hidden layers and varying numbers of nodes in the hidden layers.
+We show an MLP network with two hidden layers in Fig.~\ref{fig:neuralnet}.
+The dimensions in the figure match those of the accuracy experiment described in \S~\ref{sec:NNacc}.
+For the performance experiment of \S~\ref{sec:NNperf}, we use an MLP with four hidden layers and varying numbers of nodes in the hidden layers.
 
@@
 
-\Cref{fig:NN_acc_train} plots the accuracy on the training data over each epoch, and \cref{fig:NN_acc_test} shows the test accuracy over epochs. 
+Figure~\ref{fig:NN_acc_train} plots the accuracy on the training data over each epoch, and Fig.~\ref{fig:NN_acc_test} shows the test accuracy over epochs. 
 The matrix multiplication error introduced from these algorithms has minimal effect on the training error; all algorithms yield convergence to nearly full accuracy after 20 epochs or so.
@@
 \renewcommand{\threads}{1}
-\input{plots/NN_reltime}
+%\input{plots/NN_reltime}
+\includegraphics{tikz-taps/paper-figure8.pdf}
 \label{fig:NNreltime_seq}}
@@
 \renewcommand{\threads}{6}
-\input{plots/NN_reltime}
+%\input{plots/NN_reltime}
+\includegraphics{tikz-taps/paper-figure9.pdf}
 \label{fig:NNreltime_par_6}}
@@
 \renewcommand{\threads}{12}
-\input{plots/NN_reltime}
+%\input{plots/NN_reltime}
+\includegraphics{tikz-taps/paper-figure10.pdf}
 \label{fig:NNreltime_par_12}}
@@
 We also match the batch size of each network to the number of nodes in the hidden layer to produce square matrix multiplications in the hidden layers. 
-\cref{fig:NNreltime} shows the empirical results using 1, 6, and 12 threads.
+Fig.~\ref{fig:NNreltime} shows the empirical results using 1, 6, and 12 threads.
 
-As shown in \cref{fig:NNreltime_seq}, the single-threaded results begin to show speed up over the classical algorithm when the matrix dimensions are at least 1024.
+As shown in Fig.~\ref{fig:NNreltime_seq}, the single-threaded results begin to show speed up over the classical algorithm when the matrix dimensions are at least 1024.
 The highest performing algorithm is $\dims{4,4,4}$, which obtains a 25\% speed up at dimension 8192. 
@@
 
-\Cref{fig:NNreltime_par_6} shows the results for six threads (one socket).
-As described in \cref{sec:matmul_perf_par}, the APA algorithms do not scale to higher numbers of cores as efficiently as \texttt{gemm}, due to both the overhead of matrix additions (a memory bandwidth bottleneck) and reduced parallel performance of \texttt{gemm} for smaller dimensions.
+Figure~\ref{fig:NNreltime_par_6} shows the results for six threads (one socket).
+As described in \S~\ref{sec:matmul_perf_par}, the APA algorithms do not scale to higher numbers of cores as efficiently as \texttt{gemm}, due to both the overhead of matrix additions (a memory bandwidth bottleneck) and reduced parallel performance of \texttt{gemm} for smaller dimensions.
 In this case, not all APA algorithms outperform the classical algorithm.
@@
 
-In the case of 12 threads (both sockets), as shown in \cref{fig:NNreltime_par_12}, most APA algorithms underperform the classical algorithm.
-These results match the behavior of standalone matrix multiplications using 12 threads (see \cref{sec:matmul_perf_par}); the lack of scaling is due in large part to the poor performance of multithreaded \texttt{gemm} on remainder multiplications of small dimensions.
+In the case of 12 threads (both sockets), as shown in Fig.~\ref{fig:NNreltime_par_12}, most APA algorithms underperform the classical algorithm.
+These results match the behavior of standalone matrix multiplications using 12 threads (see \S~\ref{sec:matmul_perf_par}); the lack of scaling is due in large part to the poor performance of multithreaded \texttt{gemm} on remainder multiplications of small dimensions.
 The algorithm with a number of sub-multiplications that is a multiple of 12, $\dims{4,4,2}$, continues to perform well and is slightly faster than the classical algorithm for dimensions 4096 and 8192, attaining speedups of up to 7\%.
@@
 \centering
-\input{plots/VGG_time_seq}
+%\input{plots/VGG_time_seq}
+\includegraphics{tikz-taps/paper-figure11.pdf}
 \label{fig:vgg_seq}}%
@@
 \centering
-\input{plots/VGG_time}
+%\input{plots/VGG_time}
+\includegraphics{tikz-taps/paper-figure12.pdf}
 \label{fig:vgg_par6}}
@@
 Replacing the classic matrix multiplication algorithm in these layers with the $\dims{4,4,2}$ algorithm provides up to a 10\% speed up of the fully connected layers during training with 6 threads and a 15\% speed up using a single thread.
-\Cref{fig:vgg} shows the per-batch training time of the fully connected layers across batch sizes.
+Figure~\ref{fig:vgg} shows the per-batch training time of the fully connected layers across batch sizes.
 

### 2021-07-06 latex/ICPP21/paper.tex
converted to single column format and fixed up some issues

@@
-\documentclass[sigconf]{acmart}
+\documentclass[manuscript]{acmart}
 
@@
 
-\title{Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms}
+\title[Accelerating NN Training using APA Matrix Multiplication Algorithms]{Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms}
 
@@
 %\input{plots/matmul_perf}
-\includegraphics{tikz-taps/paper-figure2.pdf}
+\includegraphics[scale=.85]{tikz-taps/paper-figure2.pdf}
 %\caption{Single-threaded matrix multiplication performance}
@@
 %\input{plots/matmul_perf}
-\includegraphics{tikz-taps/paper-figure3.pdf}
+\includegraphics[scale=.85]{tikz-taps/paper-figure3.pdf}
 %\caption{Multithreaded matrix multiplication performance (6 threads)}
@@
 %\input{plots/matmul_perf}
-\includegraphics{tikz-taps/paper-figure4.pdf}
+\includegraphics[scale=.85]{tikz-taps/paper-figure4.pdf}
 %\caption{Multithreaded matrix multiplication performance (6 threads)}
@@
 %\input{plots/NN_acc}
-\includegraphics{tikz-taps/paper-figure5.pdf}
+\includegraphics[scale=.9]{tikz-taps/paper-figure5.pdf}
 \label{fig:NN_acc_train}}
@@
 %\input{plots/NN_acc}
-\includegraphics{tikz-taps/paper-figure6.pdf}
+\includegraphics[scale=.9]{tikz-taps/paper-figure6.pdf}
 \label{fig:NN_acc_test}}
@@
 
-\begin{figure}[hb!]
+\subsection{MLP Network Structure}
+
+\begin{figure}[ht!]
 \centering
@@
 \end{figure}
-
-\subsection{MLP Network Structure}
 
@@
 %\input{plots/NN_reltime}
-\includegraphics{tikz-taps/paper-figure8.pdf}
+\includegraphics[scale=.85]{tikz-taps/paper-figure8.pdf}
 \label{fig:NNreltime_seq}}
@@
 %\input{plots/NN_reltime}
-\includegraphics{tikz-taps/paper-figure9.pdf}
+\includegraphics[scale=.85]{tikz-taps/paper-figure9.pdf}
 \label{fig:NNreltime_par_6}}
@@
 %\input{plots/NN_reltime}
-\includegraphics{tikz-taps/paper-figure10.pdf}
+\includegraphics[scale=.85]{tikz-taps/paper-figure10.pdf}
 \label{fig:NNreltime_par_12}}
@@
 %\input{plots/VGG_time_seq}
-\includegraphics{tikz-taps/paper-figure11.pdf}
+\includegraphics[scale=1.75]{tikz-taps/paper-figure11.pdf}
 \label{fig:vgg_seq}}%
@@
 %\input{plots/VGG_time}
-\includegraphics{tikz-taps/paper-figure12.pdf}
+\includegraphics[scale=1.75]{tikz-taps/paper-figure12.pdf}
 \label{fig:vgg_par6}}
